---
title: "The Analysis Behind Movies"
author: YES Group
date: "May 16, 2018"
output:
    xaringan::moon_reader:
    css: ["default","myremark.css"]
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  echo=FALSE,
  comment = "",
  fig.height = 4,
  fig.width = 8,
  fig.align = "center",
  cache = FALSE
)

```

```{r}

## Loading the libraries
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(tibble)
library(visdat)
library(naniar)
library(rworldmap)
library(stringr)
library(tidyr)
library(rvest)
library(broom)
library(UpSetR)
library(plotly)

```


# Overview

- Introduction
- Data Processing
- Plot perspectives (ROI, actors, genre, movie studios, and runtime effects)
- Conclusion

---

# Introduction

The main focus of our project was to understand what variables could affect a movie's return on investment (ROI). 

About the movie data set:
- The movie data set was sourced from Kaggle with 40,000+ movies 
- It does not incorporate every movie that was producted
- Movies ranges from 1910s to 2017


---

# Data Processing

The data set variables includes budget, genres, vote average, actors, production companies, production companies, revenue etc.

Some challenges were:
- The majority of the numerical data was missing (budget and revenue)
- The string data was messy

Solutions:
- Implemented web scrapping for numerical data
- Performed significant amount of string manipulation to retrieve the required string data

---

# Summary of plots

- 1) Understanding the distirbution of movie productions across the world for this data set
- 2) Top 10 grossing films vs. top 10 ROI
- 3) The effect of a popular actor against ROI
- 4) Effect of genre and its ROI
- 5) Influence of movie studios and run-times on ROI

---

## 1. The distribution of movie productions

```{r results='hide'}

Mixed_Data_V2 <- read.csv("Cleaned_Data.csv")


Mixed_Data_V2 <- Mixed_Data_V2 %>%
  mutate(production_countries = as.character(production_countries))
country_codes <- read_csv("country-codes.csv")

country_codes$amount <- 0
  

for (i in seq(1,nrow(country_codes),1)){
temp_name = country_codes$official_name_en[i]
temp_alpha_code = country_codes$`ISO3166-1-Alpha-3`[i]

test <- filter(Mixed_Data_V2, str_detect(Mixed_Data_V2$production_countries, paste(temp_name, collapse="|")))
country_codes$amount[i] = count(test)

}


country_codes <- country_codes %>%
  mutate(`ISO3166-1-Alpha-3` = as.factor(`ISO3166-1-Alpha-3`)) %>%
  mutate(amount = as.numeric(amount)) %>%
  mutate(amount = ifelse(amount == 0, NA, amount))
prod_map <- joinCountryData2Map(country_codes, joinCode = "ISO3",
  nameJoinColumn = "ISO3166-1-Alpha-3")

 mapCountryData(prod_map, nameColumnToPlot="amount", catMethod = "logFixedWidth", missingCountryCol = gray(.8), mapTitle = 'Number of Movie Productions In Each Country - Log Scale')


```

- North America as well as certain areas of Europe (France and Italy) have been home to a majority of the worlds film-productions 
- Countries such as Australia, China and India have also commonly been chosen as areas for film production
- Factors affecting location: unique filming environments, home grown filmmaking culture development, and overall globalisation of filmmaking 

## 2. Top 10 grossing films vs. top 10 ROI
```{r}
##Question (8) What are the top 10 grossing films?
Mixed_Data_V2 <- Mixed_Data_V2 %>%
  mutate(budget = ifelse(budget < 5000, NA, budget)) %>%  # Agreed value that $5000 should be null
  mutate(revenue = ifelse(revenue == 0, NA, revenue)) %>%    # $0 indicating the value was not
  mutate(ROI = revenue/budget)%>%
  mutate(vote_average2 = round(vote_average)) %>%
  mutate(ROI_log10 = log10(ROI)) 

Top10 <- Mixed_Data_V2 %>%
  arrange(desc(WorldwideBox.Office)) %>%
    slice(1:10)

Top10$Title <- factor(Top10$Title, level = Top10$Title[order(-Top10$WorldwideBox.Office)])


## Top 10 ROI with Top 10 boxoffice
TopROI <- Mixed_Data_V2 %>%
  filter(!is.na(ROI)) %>%
  arrange(desc(ROI)) %>%
  slice(1:10)

TopROI$class = "Top 10 ROI"
Top10$class = "Top 10 box office"  

Top20 <- rbind(TopROI, Top10)


Top20$Title <- factor(Top20$Title, level = Top20$Title[order(-Top20$ROI)])


p <- Top20 %>%
  ggplot() + 
  geom_point(aes(x=Title, y=ROI, size=vote_average2, colour = class ))+
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) + 
  ggtitle("Top 10 ROIs with 10 top box office") 

ggplotly(p)  %>% config(displayModeBar = F) 


```

- High world-wide box office performance does not always constitute great ROI
- The highest earning films in terms of ROI are observed to return ROI orders of magnitude larger than those of the top high-grossing films


---
# 3. The effect of a popular actor against ROI

```{r}
#Question
#Does having a world class actor, such as the beloved Tom Hanks, increase the ROI of the genre of movie.
#Tom Hanks ROI vs Avg for that category
castList <- c("Tom Hanks")
cast_tom <- filter(Mixed_Data_V2, str_detect(Actors, paste(castList, collapse="|")))
cast_tom <- cast_tom %>%
  arrange(desc(WorldwideBox.Office)) %>%
  filter(!is.na(ROI))##%>%
  ##slice(1:10)

Mixed_Data_v3<-Mixed_Data_V2 %>%
  filter(!is.na(ROI)) %>%
  group_by(All_Genres)%>%
  summarise(mean(ROI))

# Joining the tables
cast_tom <- left_join(cast_tom, Mixed_Data_v3,
                            by = c("All_Genres"="All_Genres"))
colnames(cast_tom)[ncol(cast_tom)]<- "GenAvg"

p <- cast_tom %>%
  ggplot() + geom_point(aes(x=All_Genres, y=ROI, color='Top Tom Hanks',label = Title))+
geom_point(aes(x=All_Genres, y=GenAvg,color='Genre Average'))+
theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
ggtitle("Tom Hanks movies ROI by Genre vs the average for the Genre")
ggplotly(p,tooltip=c("Title","All_Genres","ROI","GenAvg")) %>% config(displayModeBar = F) 

```

- Tom Hank's movies generally performs better than the average in the same movie genres
- Popular actors may have significant marketing value 
- It could be worth it for studios spending more money to hire A-list actors to boost the overall ROI

---
# 4. Effect of genre and its ROI
```{r}
Genres_Votes <- as.data.frame(matrix(ncol = 0, nrow = 16))
Genres_Votes$Genre <- names(Mixed_Data_V2[,c(4:19)])
Genres_Votes$Average <- 0
  

for (i in seq(1,nrow(Genres_Votes),1)){

  temp_genre = Genres_Votes$Genre[i]
  

test <- filter(Mixed_Data_V2, str_detect(Mixed_Data_V2$All_Genres, paste(temp_genre, collapse="|")))
Genres_Votes$Average[i] = mean(test$vote_average ,na.rm = T)
Genres_Votes$ROI[i]=mean(test$ROI,na.rm=T)


}

x <- Genres_Votes %>%
  ggplot() + geom_point(aes(x=Genre, y=ROI, size=Average, colour = Genre))+
theme(axis.text.x = element_text(angle = -45, hjust = 0), legend.position = 'none') +
ggtitle("Average ROI per category")
ggplotly(x) %>% config(displayModeBar = F) 

##upset(Mixed_Data_V2[,c(4:19,33)], set.metadata = list(data = Genres_Votes, plots = list(list(type = "hist", column = "Average", assign = 20))))
```

- On average documentaries, horror and mystery movies have the highest returns of investment
- Although ROI can be high, the average vote for that genre may be low (size of circle)
- Generally the same vote average across all genres

---

# 5. Influence of movie studios and run-times on ROI
```{r}
#Question: Discussion of the influence of movie studios and general film run-times on ROI.

metadata_full <- read.csv("movies_metadata.csv", na.strings=c("","NA"))
metadata <- metadata_full[,c("id", "production_companies", "runtime")]
rm(metadata_full)

metadata <- metadata %>%
  mutate(production_companies = as.character(production_companies)) %>%
  mutate(production_companies = sub("\\[]",NA, production_companies)) %>%
  mutate(runtime = as.numeric(runtime)) %>%
  mutate(production_companies = sub(".*name': '*(.*?) *',.*", "\\1", production_companies)) %>%
  mutate(id = as.numeric(as.character(id))) %>%
  mutate(production_companies = ifelse(production_companies == "Columbia Pictures Corporation","Columbia Pictures", production_companies)) %>%
  mutate(production_companies = ifelse(production_companies == "Paramount Pictures Corporation" ,"Paramount Pictures", production_companies)) %>%
  mutate(production_companies = ifelse(production_companies == "Paramount" ,"Paramount Pictures", production_companies)) %>%
  filter(!is.na(runtime))

top_companies <- metadata %>%
  count(production_companies) %>%
  na.omit(metadata, cols = 'production_companies') %>%
  mutate(n = as.numeric(n))

runtime_average <- metadata %>%
  group_by(production_companies) %>%
  summarise(runtime = mean(runtime)) #%>%
  #mutate(runtime = ifelse(runtime == 0, NA, runtime))

runtime_average <- na.omit(runtime_average, cols = 'runtime')

Mixed_Data_V3 <- left_join(Mixed_Data_V2, metadata,
                            by = c("id"="id"))

Mixed_Data_V3 <- na.omit(Mixed_Data_V3, cols = 'production_companies.y')

prod_ROI <- Mixed_Data_V3 %>%
  group_by(production_companies.y) %>%
  summarise(ROI = mean(ROI))

prod_ROI <- left_join(prod_ROI, top_companies, by = c("production_companies.y"="production_companies"))

prod_ROI <- left_join(prod_ROI, runtime_average, by = c("production_companies.y"="production_companies"))

prod_10_n <- prod_ROI %>%
  arrange(desc(n)) %>%
  slice(1:10)

prod_10_ROI <- prod_ROI %>%
  arrange(desc(ROI)) %>%
  slice(1:10)

p <- prod_10_n %>%
  ggplot(aes(x=runtime, y = ROI, size = n, fill = production_companies.y)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name = "Production Companies") +
  theme(legend.title=element_blank()) +
  geom_point(aes(label = production_companies.y)) + ggtitle("Top 10 Prod. Companies by # of movies made") 

ggplotly(p)  %>% config(displayModeBar = F) 
```

- Film studios that produce the most movies tend to have similiar run times
- These particular studio's tend to have within 2% to 10% ROI on each film

---

# 5. Influence of movie studios and run-times on ROI (Continued)
```{r}
l <- prod_10_ROI %>%
  ggplot(aes(x=runtime, y = n, size = ROI, fill = production_companies.y)) +
  ylab("Number of movies produced") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name = "Production Companies") +
  theme(legend.title=element_blank()) +
  geom_point(aes(label = production_companies.y)) + ggtitle("Top 10 Prod. Companies by ROI") 

ggplotly(l) %>% config(displayModeBar = F) 
```

- Smaller studios produce lower budget films
- Smaller studios generally produce films for more niche demographics 
- More original content is developed

---

# CONCLUSION

- Mainly exploring movies produced in America and certain parts of Europe
- Top box office does not indicate top ROI
- Popular actors could have a positive effect on ROI
- Top ROI by genres are mystery and genre. All genres generally have the same average vote
- The number of movies made does not dictate a higher ROI. Original genre-type content is more significant.

The trends in the various plots highlights that production companies are encouraged to mix certain factors in order to maximise ROI. We hope to encourage production companies to become more innovative in the films they produce by unveiling trends that link profitibility to factors beyond what is known and trusted thus providing insight into the nature of filmmaking as a business and art.    
