---
title: "Assignment2"
author: "Group DQ"
date: "2018Äê3ÔÂ24ÈÕ"
output: html_document
---


#First input the data
```{r}
library(tidyverse)
library(stringr)
survey <- read_csv("survey.csv", na = c("n/a", "NA"))%>% filter(id != 64)
```

#check the data

```{r}
summary(survey)
glimpse(survey)
```


#change date in Q2 to numerical
```{r}

#Q2
survey <- survey %>%
  mutate(Q2 = sub("1/2", "0.5", Q2)) %>%
  mutate(Q2 = sub(" hours", "", Q2)) %>%
  mutate(Q2 = sub("\n", "", Q2)) %>%
  mutate(Q2 = sub("hrs", "", Q2)) %>%
  mutate(Q2 = sub("HOUR", "", Q2)) %>%
  mutate(Q2 = sub("4/7", "0.57", Q2)) %>%
  mutate(Q2 = sub("three times", "", Q2))
survey$Q2[27] <- "2.0"
survey <- survey %>% mutate(Q2 = as.numeric(Q2))


```


#only get the numbers form Q5,10,13,21,22,23,24
```{r}
#Q5

library(stringr)
numextract <- function(string){ 
  str_extract(string, "\\-*\\d+\\.*\\d*")
} 

survey$Q5 <- numextract(survey$Q5)

#Q10

survey$Q10 <- numextract(survey$Q10)

#Q13
survey$Q13 <- numextract(survey$Q13)

#Q21
survey$Q21 <- numextract(survey$Q21)

#Q22
survey$Q22 <- numextract(survey$Q22)

#Q23  

survey <- survey %>%
  mutate(Q23 = sub("1997", "21", Q23))%>%
  mutate(Q23 = sub("XXI", "21", Q23))
survey$Q23 <- numextract(survey$Q23)

#24
survey$Q24 <- numextract(survey$Q24)



```

#change all letter to CAPITAL
```{r}
survey$Q4 <- toupper(survey$Q4)
survey$Q6 <- toupper(survey$Q6)
survey$Q9 <- toupper(survey$Q9)
survey$Q18 <- toupper(survey$Q18)
survey$Q25 <- toupper(survey$Q25)
survey$Q32 <- toupper(survey$Q32)
```

# unify the data
```{r}
library(lettercase)

#Q4
survey$Q4 <-as.factor(str_replace(survey$Q4, "-", ""))
survey <- survey %>%
  mutate(Q4 = sub("ALT ROCK", "ALTERNATIVE ROCK", Q4))%>%
  mutate(Q4 = sub("HIPHOP", "HIP HOP", Q4))%>%
  mutate(Q4 = sub("POP MUSIC", "POP", Q4))%>%
  mutate(Q4 = sub("POPULAR", "POP", Q4))%>%
  mutate(Q4 = sub("HIPPOP", "POP", Q4))


#Q6
survey <- survey %>%
  mutate(Q6 = sub("FLYING", "FLY", Q6))%>%
  mutate(Q6 = sub("FLIGHT", "FLY", Q6))%>%
  mutate(Q6 = sub("FORECAST THE FUTURE", "FORECASTING", Q6))%>%
  mutate(Q6 = sub("INVISIBLE", "INVISIBILITY", Q6))%>%
  mutate(Q6 = sub("SUPER-SPEED", "SPEED", Q6))%>%
  mutate(Q6 = sub("TIME CONTROL", "TIME", Q6))%>%
  mutate(Q6 = sub("TIME TRAVEL", "TIME", Q6))%>%
  mutate(Q6 = sub("TIME-REVERSE", "TIME", Q6))%>%
  mutate(Q6 = sub("TELEPORTING", "TELEPORTATION", Q6))



#Q9
survey <- survey %>%
  mutate(Q9 = sub("NETLFIX", "NETFLIX", Q9))%>%
  mutate(Q9 = sub("NON", "NONE", Q9))%>%
  mutate(Q9 = sub("NOTHING", "NONE", Q9))%>%
  mutate(Q9 = sub("I DON'T WATCH TV", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("I DON'T WATCH TELEVISION.", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("I DO NOT WATCH TELEVISION.", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("DONT WATCH TV", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("DONT WATCH", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("DON'T WATCH TELEVISION", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("NONEE", "NONE", Q9))%>%
  mutate(Q9 = sub("I I DON'T WATCH TELEVISION", "I DON'T WATCH TELEVISION", Q9))%>%
  mutate(Q9 = sub("I DON'T WATCH TELEVISION", "I DON'T WATCH TELEVISION", Q9))

#Q18
survey <- survey %>%
  mutate(Q18 = sub("GAME", "GAMING", Q18))%>%
  mutate(Q18 = sub("GAMES", "GAMING", Q18))%>%
  mutate(Q18 = sub("LISTEN TO MUSIC", "LISTEN MUSIC", Q18))%>%
  mutate(Q18 = sub("PLAY GAMES", "GAMING", Q18))%>%
  mutate(Q18 = sub("VIDEO GAME", "GAMING", Q18))%>%
  mutate(Q18 = sub("WATCH TV", "TV", Q18))%>%
  mutate(Q18 = sub("WATCH YOUTUBE", "YOUTUBE", Q18))%>%
  mutate(Q18 = sub("PLAY VIDEO GAMINGS", "GAMING", Q18))%>%
  mutate(Q18 = sub("GAMINGS", "GAMING", Q18))%>%
  mutate(Q18 = sub("PLAY GAMINGS", "GAMING", Q18))%>%
  mutate(Q18 = sub("PLAY GAMING", "GAMING", Q18))


#Q32
survey <- survey %>%
  mutate(Q32 = sub("CHOPSTICK(S)", "CHOPSTICKS", Q32))


```

#change the character to number and assume they are numeric
#this is easiler to get the correlation between the character and other variable.
#however the shape of them only represent the count of each level inside this variable, not the continously relate to other levels

```{r}
#how many level in each col
sapply(survey, function(col) length(unique(col)))

table(survey$Q4)
unique(survey$Q4)


#survey2 data is only used in ggpair

survey2 <- survey %>%
  mutate(Q4 = rank(survey$Q4))%>%
  mutate(Q6 = rank(survey$Q6)) %>%
  mutate(Q9 = rank(survey$Q9)) %>%
  mutate(Q18 = rank(survey$Q18)) %>%
  mutate(Q25 = rank(survey$Q25)) %>%
  mutate(Q32 = rank(survey$Q32))

survey$Q2 <- as.numeric(survey$Q2)
survey2$Q4 <- as.numeric(survey2$Q4)
survey2$Q5 <- as.numeric(survey2$Q5)     
survey2$Q6 <- as.numeric(survey2$Q6)
survey2$Q9 <- as.numeric(survey2$Q9)
survey2$Q13 <- as.numeric(survey2$Q13)
survey2$Q18 <- as.numeric(survey2$Q18)
survey2$Q24 <- as.numeric(survey2$Q24)
survey2$Q25 <- as.numeric(survey2$Q25)
survey2$Q32 <- as.numeric(survey2$Q32)




```

#check the correlation and shape between each variables
##this will cost long time, delete "#" if runing the program

```{r}
library(GGally)

ggpairs(survey2[,2:12])
#ggpairs(survey2[,13:23])
#ggpairs(survey2[,24:34])

#this code will cost alot of time!!

#ggpairs(survey2[,2:34],cardinality_threshold = NULL)

```

#base on the ggpairs, select the interesting or stranging shape of graph
#check the question if really relate if they have correlation ratio 
#show the single graph with the original data without ranking

```{r}

library(ggplot2)


ggplot(survey,aes(x=Q2)) + geom_histogram(na.rm = TRUE)

ggplot(survey,aes(x=Q20))+ geom_bar()

ggplot(survey,aes(x=Q33))+ geom_bar()

ggplot(survey,aes(x=Q3,fill=Q1)) + geom_bar() 

ggplot(survey,aes(x=Q12,fill=Q17)) + geom_bar() 

ggplot(survey,aes(x=Q31,fill=Q30)) + geom_bar()
table(survey2$Q30,survey$Q31)

ggplot(survey, aes(x=Q30, fill=Q31))  + 
  geom_bar() + facet_wrap(~ Q17)
```

#none of questions are highly correlated
#in Q2, someone commuting 120 each week
#in Q13, there is one guys spend 300 everyday to eat
#in Q20 most of people aggre same-sex marriage
#in Q33, most of people like cheese
#base on Q3 and Q1, the students attended Ed Sheeran's concert in 2018 are most in year 3
#base on Q12 and Q17, double degree have more chance have prior coding experience
#base on Q30 and Q31, student who not stuggling with this unit have higher chance like this unit
#in addition they most don't have pior coding experience


