---
title: "assignment_3"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(visdat)
library(naniar)
library(impute)
library(data.table)
library(broom)
```

```{r reading in data}
houses <- read_csv("Melbourne_housing_FULL.csv")
houses <- houses %>%
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom)
```

```{r missing report}
vis_dat(houses, palette = "cb_safe")
vis_miss(houses, sort_miss=TRUE) + theme(aspect.ratio=1)
s_miss <- miss_summary(houses)
s_miss$miss_case_table

s_miss_group <- houses %>% 
  group_by(Type) %>% miss_summary()
s_miss_group$miss_case_table

houses_shadow <- bind_shadow(houses)
ggplot(data = houses_shadow,
       aes(x = Bedroom2, Bathroom, colour=Bathroom_NA)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point(alpha=0.7) + theme(aspect.ratio=1) 
```
##Missing Values Analysis

######From the first graph we can see that the types of values missing are only from integer variables: Price, Bedroom2, Bathroom. At first glance it appears that Bedroom2 and Bathroom have the same missing values consistently and that there are a similar amount of missing values from Price as there is from these two columns. 

######The second graph indicates that 11.5% of values are missing in total.That the Bathroom and Bedroom2 data is missing more variables than price by 2.2%. It is also important to note that many of the price values that are NA do not match up with NA values from the Bedroom2/Bathroom NA values.

######Looking at the number of observations missing values, its interesting to note that more observations miss two values than one. This is because values missing from Bedroom2 tend also to miss from bathroom. There is a very small number of cases in comparison to the number of data sets that have all three variables (Price, Bedroom2 and Bathroom) missing values in an observations.


```{r fixing na values for price}
pricedata <- houses %>%
  select(Price, Rooms, Type, Bedroom2, Bathroom, Distance) %>%
  subset(!is.na(Bedroom2)) %>%
  subset(!is.na(Bathroom))
pricedata <-  pricedata %>%
  mutate(idnumb = as.numeric(row.names(pricedata)))


pricedatafixed <- pricedata%>%
  select(Price, Rooms, Type, Bathroom, Bedroom2, idnumb) %>%
  arrange(Price) %>%
  subset(is.na(Price)) %>%
  select(Rooms, Type, Bedroom2, Bathroom, idnumb)

pricedataexist <- pricedata%>%
  select(Price, Rooms, Type, Bathroom, Bedroom2, idnumb) %>%
  arrange(Price) %>%
  subset(!is.na(Price))

pricefit <- lm(Price~Rooms+Type+Bedroom2+Bathroom, data=pricedataexist)
tidy(pricefit)
pricedatafixed <- pricedatafixed %>%
  mutate( 'predicted_Price' = round(predict(pricefit, pricedatafixed))) %>%
  complete(idnumb = 1:26631)

datafitted <- pricedata %>%
  mutate( 'PredPrice' = pricedatafixed$predicted_Price) %>%
  transform(Price = pmax(Price, PredPrice, na.rm = TRUE)) %>%
  select(Price, Rooms, Type, Bedroom2, Bathroom, Distance)
vis_miss(datafitted, sort_miss=TRUE) + theme(aspect.ratio=1)
```
##Fixing NA Values

######Because of the relationship between missing values in Bedroom2 and Bathroom and the fact that proportionally very few values only had all three variables missing these observations were removed. A linear model was then created using only observations that have all variables. This was then used to predict the price values that were missing from the data set. 

```{r variables decisions}
ggplot(datafitted,
       aes(x = Distance,
           y = Price)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

ggplot(datafitted,
       aes(x = Rooms,
           y = Price)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

ggplot(datafitted,
       aes(x = Type,
           y = Price)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

ggplot(datafitted,
       aes(x = Type,
           y = Rooms)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

ggplot(datafitted,
       aes(x = Bedroom2,
           y = Price)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

ggplot(datafitted,
       aes(x = Bathroom,
           y = Price)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

ggplot(datafitted,
       aes(x = Bathroom,
           y = Bedroom2)) +
  scale_colour_brewer(palette="Dark2") +
  geom_point() +
  theme(aspect.ratio=1)

```
##Variable Decisions 

######Deciding that price would be the dependent variable as it would be determined by the factors surrounding the house. All of the other variables were graphed against price to see what relationships could be seen. 

######From Graph 1 we can see that there doesn't appear to be any kind of distinct relationship between distance and price. Therefore distance was not used in the model creation.

######From Graph 2 we can see that, in general the more rooms that existed the higher the price will be. 

######From Graph 3 we can see that the houses have the opportunity to have a greater price, followed by the townhouses and then units.

######Graph 4 looks at how the number of rooms is effected by the type of house. Intuitively it feels like the bigger the place the more rooms it would have. (Aka. Houses have more rooms, then townhouses then apartments). This indicates an interaction between Type and Rooms.

######Graph 5 and 6 look at the relationships between Bedroom2 and Price or Bathroom and Price. These both appear to be related to Price. The final graph looks at how they interact with each other which also appears to be very linear.
```{r model creation}
modelfitted <- lm(Price ~ Type*Rooms + Bedroom2*Bathroom, data=datafitted)
tidy(modelfitted)

model <- augment(modelfitted, datafitted)
  
ggplot(model, aes(x=.fitted, y=Price)) + 
  geom_point() +
  geom_line(aes(y=.fitted), colour="blue")

ggplot(model, aes(x=.fitted, y=Price)) + 
  geom_point() +
  facet_grid(~Type) +
  geom_line(aes(y=.fitted), colour="blue")

glance(modelfitted)

```
#Model

######The interactions between Rooms and Type, Bedroom2 and Bathroom were included in the model. Looking at the two graphs available show that the  overall model fits quite well, given that there is an abundance of values close together. The second graph shows that the model is fitted best for units. Glimpsing the model shows however that the Rsquared term is quite low at 0.3267. This is due to the fact that there are many other variables that dictate how expensive a house/apartment/townhouse actually costs. These include location (such as suburb), newness of the house, and facilities (eg. pool, garage, tennis court). Given the fact that these variables were not taken into account a low Rsquared term is expected. 
