---
title: "Assignment_2"
author: "Bilbies"
date: "27 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading the data:
This was provided as part of the assignment properties. 
Warnings have been put to false so that the graphs presented can be seen clearly and the visual representations are clear.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
survey <- read_csv("data/survey.csv")
```


## Tidying the data: 
#### Questions that needed to be tidied:
1,2,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,21,22,23,24,25,26,27,28,29,30,31,32

Only rows 1 to 63 are used because the 64th row was an observation containing no answers. This means it would not impact of the general plots that will be created. 

mutate(): mutate allows to change each value in conjunction with case_when().

case_when(): this takes the particular question (variable) and then convert certain values within column to what you require.

else_if(): here changes are only made if certain conditions are satisfied. 
```{r}
survey1 <- survey[1:63,] %>% 
  mutate(
    Q1 = case_when(
      Q1 == "A" ~ "1", 
      Q1 == "B" ~ "2", 
      Q1 == "C" ~ "3", 
      Q1 == "D" ~ "4", 
      TRUE ~ "Other"),
    
    Q2 = case_when(
      Q2 == "three times\n" ~ NA_character_, 
      Q2 == "1/2" ~ "0.5", 
      Q2 == "4\n" ~ "4", 
      Q2 == "5hrs" ~ "5", 
      Q2 == "2 hours" ~ "2", 
      Q2 == "2 hours ( to uni)"~ "2", 
      Q2 == "3 hours" ~ "3", 
      Q2 == "0.5 hours" ~ "0.5",
      Q2 == "20HOUR" ~ NA_character_, 
      Q2 == "120" ~ NA_character_, 
      Q2 == "30" ~ NA_character_, 
      Q2 == "10" ~ NA_character_, 
      Q2 == "14" ~ NA_character_, 
      Q2 == "9" ~ NA_character_, 
      Q2 == "8" ~ NA_character_, 
      Q2 == "25" ~ NA_character_, 
      Q2 == "4/7" ~ NA_character_, 
      Q2 == "20" ~ NA_character_, 
      Q2 == "12" ~ NA_character_, 
      TRUE ~ Q2),
    
    Q4 = case_when(
      Q4 == "Anything" ~ NA_character_,
      Q4 == "Depression" ~ NA_character_,
      Q4 == "light\n" ~ NA_character_,
      Q4 == "music" ~ NA_character_,
      Q4 == "-" ~ NA_character_,
      Q4 == "Pop music" ~ "Pop",
      Q4 == "K-pop" ~ "Asian-Pop",
      Q4 == "K-Pop" ~ "Asian-Pop",
      Q4 == "alt rock" ~ "Alternative Rock",
      Q4 == "classical" ~ "Classical",
      Q4 == "country music" ~ "Country Music",
      Q4 == "hip-hop" ~ "Hip-Hop",
      Q4 == "Hiphop" ~ "Hip-Hop",
      Q4 == "hip-pop" ~ "Hip-Hop", 
      Q4 == "hip hop" ~ "Hip-Hop", 
      Q4 == "jazz" ~ "Jazz",
      Q4 == "mandopop" ~ "Asian-Pop", 
      Q4 == "pop" ~ "Pop",
      Q4 == "pop music" ~ "Pop",
      Q4 == "popular" ~ "Pop",
      Q4 == "rnb" ~ "R&B",
      Q4 == "RnB" ~ "R&B", 
      Q4 == "RNB" ~ "R&B",
      Q4 == "Rock and/or Roll" ~ "Rock",
      TRUE ~ Q4),
    
    Q5 = case_when(
      Q5 == "40hrs" ~ "40", 
      Q5 == "24 Hours" ~ "24",
      Q5 == "40 hours" ~ "40",
      Q5 == "35 hours" ~ "35",
      Q5 == "Choosing not to answer" ~ NA_character_,
      Q5 == "-" ~ NA_character_,
      Q5 == "24HOUR" ~ "24",
      Q5 == "6 hours" ~ "6",
      Q5 == "24 Hours\n" ~ "24",
      Q5 =="72\n" ~ "72",
      TRUE ~ Q5),
    
    Q6 = case_when(
      Q6 == "speed"~"Super-speed", 
      Q6 == "forecasting\n" ~ "Predict the future", 
      Q6 == "teleporting" ~ "Teleportation", 
      Q6 == "teleportation" ~ "Teleportation", 
      Q6 == "go back to a past time" ~ "Time travel", 
      Q6 == "predict the future" ~ "Predict the future", 
      Q6 == "flying" ~ "Flying",
      Q6 == "Flight"~ "Flying",
      Q6 == "have another super power" ~ "Have another super power",
      Q6 == "fly"~ "Flying",
      Q6 == "I don't need any super power." ~ "None",
      Q6 == "time control" ~ "Freeze time",
      Q6 == "to never have to sleep" ~ "To never have to sleep",
      Q6 == "forecast the future" ~ "Predict the future",
      Q6 == "remember everything I saw" ~ "Perfect memory",
      Q6 == "time travel" ~ "Time travel",
      Q6 == "Time-Reverse" ~ "Time travel",
      Q6 == "invisibility" ~ "Invisibility",
      Q6 == "freeze time" ~ "Freeze time",
      Q6 == "Control Time" ~ "Freeze time",
      Q6 == "fire" ~ "Fire powers",
      Q6 == "Never forget" ~ "Perfect memory",
      Q6 == "invisible" ~ "Invisibility",
      Q6 == "Invisible" ~ "Invisibility",
      Q6 == "Magic like Dr Strange" ~ "Magic",
      Q6 == "Fly" ~ "Flying",
      Q6 == "Not needing sleep" ~ "To never have to sleep",
      Q6 == "time" ~ "Time travel",
      Q6 == "can go to wherever i want in any time" ~ "Teleportation",
      Q6 == "imortality" ~ "Imortality",
      Q6 == "Nothing" ~ "None",
      TRUE ~ Q6),
    
    Q7 = case_when(
      Q7 == "A" ~ "two sizes too big", 
      Q7 == "B" ~ "one size too small", 
      Q7 == "C" ~ "neither of these"),
    
    Q8 = case_when(
      Q8 == "A" ~ "[0,3)", 
      Q8 == "B" ~ "[3,6)", 
      Q8 == "C" ~ "[6,9)", 
      Q8 == "D" ~ "[9,12)", 
      Q8 == "E" ~ "More than 12"),
    
    Q9 = case_when(
      Q9 == "n/a" ~ "None",
      Q9 == "no idea"~"None",
      Q9=="None"~"None",
      Q9=="No television"~"None",
      Q9=="I do not watch television."~"None",
      Q9=="I don't watch television."~"None",
      Q9=="Nothing"~"None",
      Q9=="none"~"None",
      Q9=="i don't watch tv"~"None",
      Q9=="non"~"None",
      Q9=="dont watch tv" ~ "None", 
      Q9 == "no idea\n" ~"None" , 
      Q9 =="stream.tv\n" ~ "Uncertified online service", 
      Q9 == "reality show" ~ "None", 
      Q9=="phone" ~ "Uncertified online service" , 
      Q9 == "watchseries" ~ "Uncertified online service", 
      Q9 == "Netlfix" ~ "Netflix", 
      Q9 == "netflix" ~"Netflix", 
      Q9 == "Don't watch television" ~"None",
      Q9=="internet"~"Uncertified online service",
      Q9=="pc"~"Uncertified online service",
      Q9=="I don't" ~"None", 
      Q9=="Netflix & Stan"~"Netflix", 
      Q9=="dont watch"~"None", 
      Q9=="Foxtel..."~"Uncertified online service", 
      Q9=="youtube"~"Youtube", 
      Q9=="tivo"~"Uncertified online service",
      Q9 =="optus" ~ "Optus", 
      Q9=="Kodi"~"Uncertified online service", 
      Q9=="Crunchyroll"~"Uncertified online service",
      Q9=="hdonline.is"~"Uncertified online service", 
      Q9=="I don't watch television\n"~"None",
      Q9=="stan" ~ "Stan",
      Q9=="youtube, netflix" ~ "Youtube",
      TRUE ~ Q9),
    
    Q10 = case_when(
      Q10 == "?" ~ NA_character_,
      Q10 == "1\n" ~ "1",
      TRUE ~ Q10),
    
    Q11 = case_when(
      Q11 == "A" ~ "elective", 
      Q11 == "B" ~ "core"),
    
    Q12 = case_when(
      Q12 == "A" ~ "single", 
      Q12 == "B" ~ "double"),
    
    Q13 = if_else(Q13 == "$30", "30", Q13),
    Q13 = if_else(Q13 == "$5", "5", Q13),        
    Q13 = if_else(Q13 == "$15", "15", Q13),
    Q13 = if_else(Q13 == "35 dollars", "35", Q13),
    Q13 = if_else(Q13 == "$35", "35", Q13),

    Q14 = case_when(
      Q14 == "A" ~ "0", 
      Q14 == "B" ~ "1", 
      Q14 == "C" ~ "2", 
      Q14 == "D" ~ "3", 
      Q14 == "E" ~ "4", 
      Q14 == "F" ~ "5", 
      Q14 == "G" ~ "more than 5"),
    
    Q15 = case_when(
      Q15 == "A" ~ "cucumbers", 
      Q15 == "B" ~ "carrots", 
      Q15 == "C" ~ "neither"),
    
    Q16 = case_when(
      Q16 == "F" ~"Other",
      Q16 == "A" ~ "Dog",
      Q16 == "B" ~ "Cat",
      Q16 == "C" ~ "Rabbit",
      Q16 == "D" ~ "Bird",
      Q16 == "E" ~ "Reptile",
      Q16 == "A, B" ~ "Multiple",
      Q16 == "A, C" ~ "Multiple",
      Q16 == "B, D, E" ~ "Multiple",
      Q16 == "A, B, D" ~ "Multiple",
      Q16 == "A, E" ~ "Multiple",
      Q16 == "A, D, F" ~ "Multiple",
      Q16 == "E, F" ~ "Multiple",
      TRUE ~ Q16),
    
    Q17 = case_when(
      Q17 == "A" ~ "yes", 
      Q17 == "B" ~ "no"),
    
    Q18 = case_when(
      Q18 == "listen to music" ~ "Music",
      Q18 == "sport" ~ "Sport",
      Q18 == "Basketball" ~ "Sport",
      Q18 == "watch youtube\n" ~ "Youtube",
      Q18 == "watch vedio" ~ "Youtube",
      Q18 == "internet, stretch, water" ~ "Social media",
      Q18 == "cook" ~ "Cook",
      Q18 == "sleep" ~ "Sleep",
      Q18 == "youtube" ~ "Youtube",
      Q18 == "gaming" ~ "Gaming",
      Q18 == "To listen music" ~ "Music",
      Q18 == "training my aim in different FPS games." ~ "Gaming",
      Q18 == "Play games" ~ "Gaming",
      Q18 == "social media" ~ "Social media",
      Q18 == "video game" ~ "Gaming",
      Q18 == "watch tv" ~ "TV",
      Q18 == "Games" ~ "Gaming",
      Q18 == "gym\ninternet surfing" ~ "Gym",
      Q18 == "computer games" ~ "Gaming",
      Q18 == "play games"~"Gaming",
      Q18 == "Instagram" ~ "Social media",
      Q18 == "Game" ~ "Gaming",
      Q18 == "league of legends" ~ "Gaming",
      Q18 == "the office" ~ "TV",
      Q18 == "Read, Play Guitar, Electronics, Sport" ~ "Read",
      Q18 == "This is us and cry" ~ "TV",
      Q18 == "Swim, cook" ~ "Sport",
      Q18 == "tv" ~ "TV",
      Q18 == "Have a trip\n" ~ "Have a trip",
      Q18 == "piano" ~ "Piano",
      Q18 == "Play video games" ~ "Gaming",
      Q18 == "Study more" ~ "Study",
      Q18 == "listen music" ~ "Music",
      Q18 == "Watch Something/ Light Exercise" ~ "TV",
      Q18 == "play my phone" ~ "Social media",
      Q18 == "work" ~ "Work",
      Q18 == "Your mum" ~ NA_character_,
      TRUE ~ Q18),
    
    Q19 = case_when(
      Q19 == "A" ~ "Windows", 
      Q19 == "B" ~ "Mac", 
      Q19 == "C" ~ "Linux", 
      Q19 == "D" ~ "Other"),
    
    Q21 = if_else(Q21 == "7;30", "7.5", Q21),

    Q22 = if_else(Q22 == "3days", "3", Q22),
    
    Q23 = if_else(Q23 == "1997", "20", Q23),
    Q23 = if_else(Q23 == "XXI", "21", Q23),
    
    Q24 = case_when(
      Q24 == "I don't know" ~ NA_character_, TRUE ~ Q24),
    
    Q25 = case_when(
      Q25 == "afl" ~ "AFL", 
      Q25 == "badminton" ~ "Badminton", 
      Q25 == "basketball" ~ "Basketball", 
      Q25 == "Electronic games and running" ~ "Running", 
      Q25 == "equestrian" ~ "Equestrian", 
      Q25 == "European Handball" ~ "European handball", 
      Q25 == "Football" ~ "Soccer", 
      Q25 == "soccer" ~ "Soccer", 
      Q25 == "motorsport" ~ "Motorsport", 
      Q25 == "netball" ~ "Netball", 
      Q25 == "rowing" ~ "Rowing", 
      Q25 == "running" ~ "Running", 
      Q25 == "skiing" ~ "Skiing",
      Q25 == "sleep" ~ "Sleep", 
      Q25 == "swim" ~ "Swimming", 
      Q25 == "swimming" ~ "Swimming", 
      Q25 == "swimming\n" ~ "Swimming", 
      Q25 == "tennis" ~ "Tennis", 
      Q25 == "Table Tennis" ~ "Table tennis", 
      Q25 == "mma" ~ "MMA", 
      Q25 == "Choosing not to answer" ~ NA_character_, 
      Q25 == "Yes" ~ NA_character_, TRUE ~ Q25 ),
    
    Q26 = case_when(
      Q26 == "A" ~ "Female", 
      Q26 == "B" ~ "Male", 
      Q26 == "C" ~ NA_character_),
    
    Q27 = case_when(
      Q27 == "A" ~ "Fridge", 
      Q27 == "B" ~ "Pantry"),
    
    Q28 = case_when(
      Q28 == "A" ~ "0", 
      Q28 == "B" ~ "1-5", 
      Q28 == "C" ~ "6-10", 
      Q28 == "D" ~ "10-20", 
      Q28 == "E" ~ "More than 20"),
  
    Q29 = case_when(
      Q29 == "A" ~ "Yes", 
      Q29 == "B" ~ "No"),  

    Q30 = case_when(
      Q30 == "A" ~ "No", 
      Q30 == "B" ~ "Yes"),  
  
    Q31 = case_when(
      Q31 == "A" ~ "Yes", 
      Q31 == "B" ~ "No"),
    
    Q32 = case_when(
      Q32 == "wok" ~ "Wok", 
      Q32 == "whisk" ~ "Whisk", 
      Q32 == "tongs" ~ "Tongs", 
      Q32 == "That thing that cuts eggs" ~ "Egg cutter", 
      Q32 == "spoon" ~ "Spoon", 
      Q32 == "a spoon" ~ "Spoon", 
      Q32 == "spork" ~ "Spork", 
      Q32 == "spatula" ~ "Spatula", 
      Q32 == "sharper" ~ "Sharper", 
      Q32 == "refrigerator" ~ "Refrigerator",
      Q32 == "Fridge" ~ "Refrigerator", 
      Q32 == "pot" ~ "Pot", 
      Q32 == "pan" ~ "Pan", 
      Q32 == "oven" ~ "Oven", 
      Q32 == "IKEA soup ladle" ~ "Ladle", 
      Q32 == "knife" ~ "Knife",
      Q32 == "fork" ~ "Fork", 
      Q32 == "dish washer" ~ "Dish washer", 
      Q32 == "colander" ~ "Colander", 
      Q32 == "Coffee machine" ~ "Coffee Machine", 
      Q32 == "coffee cup" ~ "Coffee cup", 
      Q32 == "chopstick(s)" ~ "Chopsticks", 
      Q32 == "None" ~ NA_character_, 
      Q32 == "someone can eat" ~ NA_character_, TRUE ~ Q32)
    )
```


#### Converting each column within the data frame to be the right data type:
Questions Converted:
2,3,5,10,13,20,21,22,23,24,33

The data type of the values had to be converted depending on the what they were representing. At the beginning, all the variables where character types. However, some observations were numbers (double or integer types) and there were also logical types (true/false). Changing to the correct data types allowed the group to make correct visual representations.
```{r}
survey2 <- survey1 %>%
  mutate(
    Q2 = as.numeric(Q2),             # travel time was presented as both whole and fraction numbers
    Q3 = as.logical(Q3),             # it was a true/false question
    Q5 = as.numeric(Q5),             # because the option for fractions was given
    Q10 = as.numeric(Q10),           # litres of water drunk was presented as both whole and fraction numbers
    Q13 = as.numeric(Q13),           # money spent on food was both whole dollars and fractions
    Q20 = as.logical(Q20),           # it was a true/false question
    Q21 = as.numeric(Q21),           # people sleep for hours and also fraction of hours
    Q22 = as.integer(Q22),           # people cannot catch up with friends a fraction number of times 
    Q23 = as.integer(Q23),           # values were present to the nearest year, not fraction of years
    Q24 = as.numeric(Q24),           # because height is a continuous variable
    Q33 = as.logical(Q33)            # it was a true/false question
    )
```


##Graphical Representations

#### Box Plot of Heights
The boxplot only reveals the range of heights present within our class, and the median height, however it doesn't inform us about other aspects within the class.
```{r warning=FALSE}
ggplot(data = survey2, aes(x = "", y = Q24)) + 
  geom_boxplot() + 
  labs(x = "ETC1010 Class Height", y = "Height (cms)")
```


#### Box Plot of Heights filtered by Gender
On the other hand, the filtered box plot with gender reveals a lot more.Heights can be compared by the people and is filted by their gender. From this representation we found that most males in the class are taller than females. The tallest person is a male (over 200 cms), and the shortest person is a female (~ 154 cms). This graph also cofirms the belief that generally males are taller than females in the overall population. In addition, it also reveals the median heights for each gender. There is however, 1 observation missing, because it was a missing value, the respondent decided not to answer the question.
```{r warning=FALSE}
survey3 <- filter(survey2, Q26 != is.na(Q26))
ggplot(data = survey3, aes(x = Q26, y = Q24)) + 
  geom_boxplot(aes(fill = Q26)) + 
  labs(x = "Gender", y = "Height (cms)", fill = "Gender")
```


#### BarChart Plot of Operating Systems facetted by Gender
This bar chart underneath isfacetted by gender reveals the difference in preference of operating system based on gender. As a team, we wanted to explore a relationship between the collective mindset of the genders and the type of OS used by them. As such, we could identify that there was a 50/50 split between Mac and Windows for females, and no Females using Linux in this class. Windows OS heavily outweighed Macs and Linux in Males, with 25 males using Windows, 11 using Macs and 2 using Linux. From this graph, we can conclude that within the population of ETC1010 lectures, more males use Windows than the other OS, and there is a 50/50 split between Mac and Windows for females.
```{r warning=FALSE}
survey3 <- filter(survey2, Q26 != is.na(Q26))
ggplot(data = survey3) + 
  geom_bar(mapping = aes(x = Q19, fill = Q19)) + 
  labs(x = "Operating System", y = "Number of people") + 
  facet_grid(~ Q26)
```


#### BarChart Plot of Enjoyment of ETC1010 facetted by Prior Coding Experience
As a team, we hypothesised that people who have prior coding experience would enjoy this subject a lot more than those students who have no prior coding experience. This bar chart, facetted by “no prior coding experience” and “prior coding experience”, demonstrates that there is a large difference between the mentality of the two groups. More people who do not have prior coding experience do not enjoy ETC1010, whereas in people with prior coding experience, there is a large proportion of people who enjoy this subject than those who do not.
```{r warning=FALSE}
ggplot(data = survey2) + 
  geom_bar(mapping = aes(x = Q31, fill = Q31)) + 
  facet_wrap(~ Q17) +
  labs(x = "Prior Coding Experience Facetting", y = "Number of people", fill = "Do you like ETC1010")
```


#### BarChart Plot of Struggle of ETC101 filtered by Prior Coding Experience
The filled bar chart reveals that: over 90% of the students with coding experience are not struggling with ETC1010, however the right bar clearly indicates that having no prior coding experience generally means you may struggle with this unit. The notions in the visual representation seem logical in the sense that prior knowledge generally helps in all avenues, and in the case of coding, it is the same.
```{r warning=FALSE}
# those struggling with unit and how that is reflected by the prior coding experience
ggplot(data = survey2) + 
  geom_bar(mapping = aes(x = Q30, fill = Q17), position = "fill") + 
  labs(y = "Percentage", x = "Are you struggling with ETC1010", fill = "Prior Coding Experience") 
```


#### BarChart Plot of Cohort's Pets
The bar chart below showcases the types of pets that the cohort has. The highest option was otherc (F). This is a very arbitary graph, because option other (F) in the questions could mean two different things. Firstly, it could selected if the respondent does not have a pet, as there was no "no pet" option. Secondly it could mean that other than: dogs, cats, birds, reptiles and rabbits the respondent has anothe pet. For those respodents that chose more than one field, they were placed into the multiple option. 

Clearly two results can be stated: either the cohort cares for unique pets, or most of the cohort does not have pets. The team believes the latter to be true. This is because there are many internationl students in our unit, and generally they would not have time or space to care for a pet.

A pie chart was not applied, because too many options would lead to the data being unreadable.
```{r warning=FALSE}
# the pets that we have
ggplot(data = survey2) + geom_bar(mapping = aes(x = fct_infreq(Q16), fill = Q16)) +
  labs(y = "Number of pets", x = "Pets")
```
