---
title: "ETC1010_Assignment 2"
author: "Raymond & Co."
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r library}
library(tidyverse)
library(stringr)
```

```{r function}
###function 1: Convert fraction to 1 decimal number
f1<-function(x){
  x<-round(
    eval(
      parse(text=x)
      )
    ,1)
  return(x)
}
##function 2:extract the numeric and round to 1 decimal
f2<-function(x){
  x<-round(
    extract_numeric(x),
    1
  )
  return(x)
}
```

```{r load_data}
survey <- read_csv("survey.csv")
head(survey)
summary(survey)
```

```{r drop_id_64}
survey%>%filter(id == 64)
survey<-survey%>%filter(id != 64)
```

```{r id}
survey$id
summary(survey$id)
```

```{r MCQ}
###factor
survey$Q1<-as.factor(survey$Q1)
survey$Q7<-as.factor(survey$Q7)
survey$Q8<-as.factor(survey$Q8)
survey$Q11<-as.factor(survey$Q11)
survey$Q12<-as.factor(survey$Q12)
survey$Q14<-as.factor(survey$Q14)
survey$Q15<-as.factor(survey$Q15)
survey$Q16<-as.factor(survey$Q16)
survey$Q17<-as.factor(survey$Q17)
survey$Q19<-as.factor(survey$Q19)
survey$Q26<-as.factor(survey$Q26)
survey$Q27<-as.factor(survey$Q27)
survey$Q28<-as.factor(survey$Q28)
survey$Q29<-as.factor(survey$Q29)
survey$Q30<-as.factor(survey$Q30)
survey$Q31<-as.factor(survey$Q31)
##logical
survey$Q3<-as.logical(survey$Q3)
survey$Q20<-as.logical(survey$Q20)
survey$Q33<-as.logical(survey$Q33)
```

```{r Q1}
levels(survey$Q1)<-c("1","2","3","4","Other")
survey$Q1
```

```{r Q7}
levels(survey$Q7)<-c("two sizes too big","one size too small","neither of these")
survey$Q7
```

```{r Q8}
levels(survey$Q8)<-c("[0,3)","[3,6)","[6,9)","[9,12)","'More than 12")
survey$Q8
```

```{r Q11}
levels(survey$Q11)<-c("elective","core")
survey$Q11
```

```{r Q12}
levels(survey$Q12)<-c("single","double")
survey$Q12
```

```{r Q14}
levels(survey$Q14)<- c("0", "1", "2", "3", "4", "5", "more than 5")
survey$Q14
```

```{r Q15}
levels(survey$Q15)<-c("cucumbers","carrots","neither")
survey$Q15
```

```{r Q17}
levels(survey$Q17)<-c("yes","no")
survey$Q17
```

```{r Q19}
levels(survey$Q19)<-c("Windows","Mac","Linux","Other")
survey$Q19
```

```{r Q26}
levels(survey$Q26)<-c("female","male","prefer not to answer")
survey$Q26
```

```{r Q27}
levels(survey$Q27)<-c("fridge","pantry")
survey$Q27
```

```{r Q28}
levels(survey$Q28)<-c("0","1-5","6-10","10-20","more than 20")
survey$Q28
```

```{r Q29}
levels(survey$Q29)<-c("yes","no")
survey$Q29
```

```{r Q30}
levels(survey$Q30)<-c("no","yes")
survey$Q30
```

```{r Q31}
levels(survey$Q31)<-c("yes","no")
survey$Q31
```

```{r factor_summary}
k2<-survey%>%select(Q1,Q3,Q7,Q8,Q11,Q12,Q14,Q15,Q16,Q17,Q19,Q20,Q26,Q27,Q28,Q29,Q30,Q31,Q33)
summary(k2)
```

```{r Numeric_NA}
k1<-survey
k1$Q2<-as.numeric(k1$Q2)
k1$Q5<-as.numeric(k1$Q5)
k1$Q10<-as.numeric(k1$Q10)
k1$Q13<-as.numeric(k1$Q13)
k1$Q21<-as.numeric(k1$Q21)
k1$Q22<-as.numeric(k1$Q22)
k1$Q23<-as.numeric(k1$Q23)
k1$Q24<-as.numeric(k1$Q24)
k1<-k1%>%select(Q2,Q5,Q10,Q13,Q21,Q22,Q23,Q24)
summary(k1)
```

```{r Q2}
p2<-survey%>%select(id,Q2)%>%
  mutate(k1_Q2=k1$Q2)%>%
  mutate(Na=is.na(k1$Q2))
p2%>%filter(Na == "TRUE")
survey$Q2[survey$Q2 == '1/2']<- f1(1/2)
survey$Q2[survey$Q2 == '4/7']<- f1(4/7)
survey$Q2<-f2(survey$Q2)
survey$Q2[is.na(survey$Q2)]<-NA
survey$Q2
```

```{r Q5}
p5<-survey%>%select(id,Q5)%>%
  mutate(k1_Q5=k1$Q5)%>%
  mutate(Na=is.na(k1$Q5))
p5%>%filter(Na == "TRUE")
survey$Q5<-f2(survey$Q5)
survey$Q5[is.na(survey$Q5)]<-NA
survey$Q5
```

```{r Q10}
p10<-survey%>%select(id,Q10)%>%
  mutate(k1_Q10=k1$Q10)%>%
  mutate(Na=is.na(k1$Q10))
p10%>%filter(Na == "TRUE")
survey$Q10<-f2(survey$Q10)
survey$Q10[is.na(survey$Q10)]<-NA
survey$Q10
```

```{r Q13}
p13<-survey%>%select(id,Q13)%>%
  mutate(k1_Q13=k1$Q13)%>%
  mutate(Na=is.na(k1$Q13))
p13%>%filter(Na == "TRUE")
survey$Q13<-f2(survey$Q13)
survey$Q13
```

```{r Q21}
p21<-survey%>%select(id,Q21)%>%
  mutate(k1_Q21=k1$Q21)%>%
  mutate(Na=is.na(k1$Q21))
p21%>%filter(Na == "TRUE")
survey$Q21[59]<-7.5
survey$Q21<-as.numeric(survey$Q21)
survey$Q21
```

```{r Q22}
p22<-survey%>%select(id,Q22)%>%
  mutate(k1_Q22=k1$Q22)%>%
  mutate(Na=is.na(k1$Q22))
p22%>%filter(Na == "TRUE")
survey$Q22<-f2(survey$Q22)
survey$Q22
```

```{r Q23}
p23<-survey%>%select(id,Q23)%>%
  mutate(k1_Q23=k1$Q23)%>%
  mutate(Na=is.na(k1$Q23))
p23%>%filter(Na == "TRUE")
survey$Q23[63]<-21
survey$Q23[59]<-21
survey$Q23<-as.numeric(survey$Q23)
survey$Q23
```

```{r Q24}
p24<-survey%>%select(id,Q24)%>%
  mutate(k1_Q24=k1$Q24)%>%
  mutate(Na=is.na(k1$Q24))
p24%>%filter(Na == "TRUE")
survey$Q24[38]<-NA
survey$Q24<-as.numeric(survey$Q24)
survey$Q24
```

```{r Num_clean}
k1_tidy<-survey%>%select(Q2,Q5,Q10,Q13,Q21,Q22,Q23,Q24)
head(k1_tidy)
summary(k1_tidy)
```

```{r character}
k3<-survey%>%select(id,Q4,Q6,Q9,Q18,Q25,Q32)
head(k3)
```

```{r Q4}
survey <- survey %>% 
  mutate(Q4 = str_to_lower(Q4)) %>%
  mutate(Q4 = sub("hip-hop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip-pop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hiphop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("&", "n", Q4)) %>%
  mutate(Q4 = sub("popular", "pop", Q4)) %>%
  mutate(Q4 = sub("music", "", Q4)) %>%
  mutate(Q4 = sub("\n", "", Q4)) %>%
  mutate(Q4 = sub("-", "", Q4)) %>%
  mutate(Q4 = trimws(Q4))
```

```{r Q4_clean}
survey$Q4
```

```{r Q6}
survey$Q6
survey$Q6[7]<-"forecasting"
for(i in 1:63){
  if(survey$Q6[i] == "Flying"){survey$Q6[i]<-"fly"}
  if(survey$Q6[i] == "Flight"){survey$Q6[i]<-"fly"}
  if(survey$Q6[i] == "Flight"){survey$Q6[i]<-"fly"}
  if(survey$Q6[i] == "flying"){survey$Q6[i]<-"fly"}
  if(survey$Q6[i] == "flying"){survey$Q6[i]<-"fly"}
  if(survey$Q6[i] == "invisible"){survey$Q6[i]<-"invisibility"}
}
survey$Q6<-str_to_lower(survey$Q6)
```

```{r Q6_clean}
survey$Q6
```

```{r Q9}
survey$Q9
survey$Q9<-str_to_lower(survey$Q9)
survey$Q9[3]<-NA
survey$Q9[6]<-"stream.tv"
survey$Q9[7]<-"no idea"
survey$Q9[54]<-"foxtel"
```

```{r Q9_clean}
survey$Q9
```

```{r Q18}
survey$Q18
survey$Q18<-str_to_lower(survey$Q18)
survey$Q18[5]<-"watch youtube"
survey$Q18[34]<-"gym, internet surfing"
survey$Q18[48]<-"Have a trip"
survey$Q18[54]<-"Watch Something, Light Exercise"
```

```{r Q18_clean}
survey$Q18
```

```{r Q25}
survey$Q25
survey$Q25<-str_to_lower(survey$Q25)
survey$Q25[55]<-"swimming"
```

```{r Q25_clean}
survey$Q25
```

```{r Q32}
survey$Q32
survey$Q32<-str_to_lower(survey$Q32)
```

```{r Q32_clean}
survey$Q32
```


### 1.For Question 16, it was a multiple choices question, we need to create the dummy variables in order to obtain the combination.

```{r Q16}
survey<-survey%>%
  mutate(Q16_dog = (ifelse(str_detect(survey$Q16,pattern = "A"),1,0)))%>%
  mutate(Q16_cat = (ifelse(str_detect(survey$Q16,pattern = "B"),1,0)))%>%
  mutate(Q16_rabbit = (ifelse(str_detect(survey$Q16,pattern = "C"),1,0)))%>%
  mutate(Q16_bird = (ifelse(str_detect(survey$Q16,pattern = "D"),1,0)))%>%
  mutate(Q16_reptile = (ifelse(str_detect(survey$Q16,pattern = "E"),1,0)))%>%
  mutate(Q16_other = (ifelse(str_detect(survey$Q16,pattern = "F"),1,0)))
d16<-survey%>%select(id,Q16,Q16_dog,Q16_cat,Q16_rabbit,Q16_bird,Q16_reptile,Q16_other)
head(d16)
d16[56,]
```


### 2
```{r finding_1}
g2<-survey%>%select(id)%>%
  mutate(degree = survey$Q12)%>%
  mutate(gender = survey$Q26)
ggplot(g2, aes(x = degree, label = gender, fill = degree))+
  geom_bar()+
  facet_grid(~gender)
```

As the graph shows, in this ETC1010 unit, female more likely doing a single degree but male more likely doing a double degree. Overall there are more males in this unit than females. There is a higher proportion of females that do a single degree as opposed to a double degree. There is a higher proportion of males who do a double degree than a single degree


### 3
```{r finding_2}
g2<-survey%>%select(id)%>%
  mutate(unit_property = survey$Q11)%>%
  mutate(struggling_with_unit = survey$Q30)
ggplot(g2, aes(x = struggling_with_unit, label = unit_property, fill = struggling_with_unit))+
  geom_bar()+
  facet_grid(~unit_property)
```

People choose this unit as elective more likely struggling with this unit, but people who choose this unit as core less likely to feel struggle.

```{r summary}
survey$Q16<-NULL
summary(survey)
```

```{r eval = FALSE}
write.csv(survey,file = "survey_tidy.csv",row.names = FALSE)
```









