---
title: "Assignment 1"
author: "Katarina and Tim"
date: "10 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Tidying Data
```{r}
library(readxl)
passengers <- read_xls("data/WebAirport_FY_1986-2017.xls", sheet=3, skip=6)
library(tidyverse)
passengers_tidy <- passengers %>%
  select(AIRPORT,Year,INBOUND,OUTBOUND,INBOUND__1,OUTBOUND__1) %>%
  unite(DOMESTIC,INBOUND,OUTBOUND,sep="_") %>%
  unite(INTERNATIONAL,INBOUND__1,OUTBOUND__1,sep="_")%>%
  gather(DOMESTIC,INTERNATIONAL,key=type_of_flight,value=Passengers) %>%
  separate(Passengers,into=c("IN","OUT"),sep="_") %>%
  gather(IN,OUT,key=bound,value=Passengers) %>%
  filter(AIRPORT != "TOTAL AUSTRALIA") %>%
  filter(Passengers != "NA")%>%
  mutate(Passengers = as.numeric(Passengers))
passengers_tidy

```

##Barchart of flight type facetted by year
```{r}
ggplot(passengers_tidy,aes(x=type_of_flight,y=Passengers,fill=type_of_flight))+
  geom_bar(stat="identity")+
  scale_y_continuous(labels=scales::comma)+
  facet_wrap(~Year)+
  ggtitle("Yearly Number of \n Domestic v International Passengers")+xlab("Type of Flight")+ylab("Number of Passengers")+
  theme(plot.title = element_text(size=14,hjust = 0.5),
        axis.text.x=element_blank(), axis.ticks.x=element_blank(),axis.text.y = element_text(size=5))

```

##Scatterplot of Domestic Inbound vs Outbound Passengers Facetted by Year
```{r}
library(ggplot2)
ggplot(passengers,aes(x=OUTBOUND,y=INBOUND)) +
  geom_point(color="blue")+
  scale_y_continuous(labels=scales::comma)+
  scale_x_continuous(labels=scales::comma)+
  facet_wrap(~Year)+
  ggtitle("Domestic Inbound v \n Outbound Passengers Per Year")+xlab("Outbound")+ylab("Inbound")+
  theme(plot.title = element_text(size=14,hjust = 0.5),axis.text.y = element_text(size=5), axis.text.x = element_text(size = 7, angle = 90))
```

##Scatterplot of International Inbound vs Outbound Passengers Facetted by Year
```{r}
library(ggplot2)
ggplot(passengers,aes(x=OUTBOUND__1,y=INBOUND__1)) +
  geom_point(color="red")+
  scale_y_continuous(labels=scales::comma)+
  scale_x_continuous(labels=scales::comma)+
  facet_wrap(~Year)+
  ggtitle("International Inbound v \n Outbound Passengers Per Year")+xlab("Outbound")+ylab("Inbound")+
  theme(plot.title = element_text(size=14,hjust = 0.5),axis.text.y = element_text(size=5), axis.text.x = element_text(size = 7, angle = 90))
```

The three airports which appear as outliers on the above scatterplot are airports Sydney, Melbourne and Brisbane.

##Summary
From these 3 plots we can make several inferences. From the faceted bar graph we can clearly see that domestic flights are significantly more popular than international flights: approximately 4 times more popular averaging across the years. We can also note that the number of flights (both international and domestic) continues to rise as time passes with approximately 30,000 domestic flights and 7000 international flights between '85 and '86 now being approximately 120,000 domestic flights and 35,000 international flights between '16 and '17. From the first and second scatterplot we make the unsurprising discovery that as the number of inbound flights increase so do the number of outbound flights (they are extremely strongly positively correlated).Although it is somewhat difficult to tell because of the size of the plots, it appears that the number of inbound and outbound flights for both international and domestic graphs are roughly the same (close to perfect linear correlation). From these scatterplots, we can also see the datapoints begin to extend following a linear pattern, to higher x and y values as each year passes. This not only indicates that flights in and out of these airports (data points) have increased over time but also maybe that certain airports have expanded their capacity and become more popular in recent times while for some other airports the traffic may not have increased as much explaining why some datapoints don't appear to have moved from '85-'86 compared to '16-'17. 