---
title: "Advanced Analysis of crime data"
subtitle: "Victorian Crime Sales"
author: "Amelia Forcat & YOUR NAME & YOUR TEAM MEMBERS NAMES & YOUR TEAM NAME"
date: "`r format(Sys.Date(), '%A, %B %e %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions to Students

OK, so you have submitted your report, and it went down pretty well!

This assignment is designed to simulate a real-life scenario where you have done some data analysis, and then you are asked to take it further by combining another dataset with the one you have used. 

Linking datasets together is a skill very important in data analysis. It allows you to combine different pieces of information together to make new insights that are otherwise not possible. It can also be challenging, as many of these datasets were not initially designed to be combined!

In this assignment you will be given two datasets to explore, and join. You will also be asked to find a third dataset, which you will need to import, tidy, and join to the data.

The first dataset is the dataset you have already used, the dataset from the Crime Statistics Agency Victoria:

> https://www.crimestatistics.vic.gov.au/download-data "Data tables - Spotlight: Burglary/Break and Enter Offences Recorded in Victoria visualisation - year ending December 2018 (XLSX, 4.4 MB)".

As mentioned in the first assignment, there are actually a few tables in the crime excel sheet - these will be used together in this assignment, along with an additional dataset. This additional dataset is from the Australian Bureau of Statistics, and contains information on the ages of people in local government areas.

You will be required to join together these datasets to answer questions from your manager. 

There will be some helper code provided by your colleague, Amelia (who has returned from her trip  to New Zealand). Amelia will be providing less assistance with this project.

The type of output expected from this assignment requires you to write some sentences or a paragraph or so about what you did, and how. Your manager will likely take figures and text directly out of the document you provide and put it into a report or presentation.

We have written example text at the start, to give an example of the type of writing we are looking for.

Like in the previous assignment, questions that are work marks are indicated with `**` at the start and end of the question, as well as a number of marks in parenthesis. You will be asked to write more in this assignment, make sure you pay attention to what sort of writing will answer the question.

## Marking + Grades

This assignment will be worth 4% of your total grade, and will be marked out of 22 marks total.

* 3 Marks for grammar and clarity. You must write in complete sentences and do a spell check.
* 3 Marks for overall presentation of the data visualisations
* 16 marks for the questions

* Your marks will then be weighted according to peer evaluation.

* Sections that contain marks are indicated with `**`, and will have the number of marks indicated in parentheses. For example:

```
# `**` ... (XX Mark) `**`
```

Where `...` is the question text, and `XX` indicates the number of marks

## A Note on skills

As of week 4, you have seen most of the code used here. 

I do not expect you to know immediately what every line of code below does. But I do expect that you can search through the helpfile to learn new parts of the code, or to turn to ED, the consults, or search engines like google to help clarify misunderstandings and errors.

This is a challenge for you! 

As mentioned earlier, this assignment is designed to simulate a real life work situation - this means that there are some things where you need to "learn on the job". 

But the vast majority of the assignment will cover things that you will have seen in class, or the readings.

Remember, you can look up the help file for functions by typing `?function_name`. For example, `?mean`. Feel free to google questions you have about how to do other kinds of plots, and post on the ED if you have any questions about the assignment.

## How to complete this assignment.

To complete the assignment you will need to fill in the blanks for function names, arguments, or other names. These sections are marked with `***` or `___`. At a minimum, your assignment should be able to be "knitted" using the `knit` button for your Rmarkdown document. 

If you want to look at what the assignment looks like in progress, but you do not have valid R code in all the R code chunks, remember that you can set the chunk options to `eval = FALSE`. If you do set `eval = FALSE`, please remember to ensure that you remove this chunk option or set it to `eval = TRUE` when you submit the assignment, to ensure all your R code runs.

You will be completing this assignment in your assigned groups. 

A reminder regarding our recommendations for completing group assignments:

 * Each member of the group completes the entire assignment, as best they can.
 * Group members compare answers and combine it into one document for the final submission. 

Your assignments will be peer reviewed, and results checked for reproducibility. 
This means:

* 25% of the assignment grade will come from peer evaluation. 
* Peer evaluation is an important learning tool. 

Each student will be randomly assigned another team's submission to provide feedback on three things: 

  1. Could you reproduce the analysis? 
  2. Did you learn something new from the other team's approach?
  3. What would you suggest to improve their work?

## Due Date

This assignment is due before the start of class close of business (5pm) on Monday 2nd Septemer You will submit the assignment via ED. 

Please change the file name to include your teams name. For example, if you are team `dplyr`, your assignment file name could read: "assignment-2-2019-s2-team-dplyr.Rmd"

Please also note that you need to submit two files:

1. The .Rmd file
2. The .html file

# Treatment

Your manager read your report on the crime - they liked it! They then immediately had a few other questions and want to see more of what you can do.

Amelia has returned from her holiday, and found a nice dataset on ages for each LGA from the ABS (Australian Bureau of Statistics) website, and has read it in for you, but hadn't really cleaned it up into tidy format. 

Amelia thinks that this would be a good learning opportunity for you to describe how you tidy your data, so she can see how you are going with your data analysis skils.

```{r read-data}
library(tidyverse)
library(readxl)

# This code tells us to read in the ABS data, which is "Table 3" in the excel file called, "32350ds0003_lga_2017.xls"
demographics_raw <- read_excel(
  "data-raw/32350ds0003_lga_2017.xls",
  # This tells us the name of the excel sheet
  sheet = "Table 3",
  # this tells us how many rows to skip before reading in data
  skip = 10,
  # This tells us the names of the columns
  col_names = c(
    c(
      "S/T code",
      "S/T name",
      "LGA code",
      "LGA name",
      "0-4",
      "5–9",
      "10–14",
      "15–19",
      "20–24",
      "25–29",
      "30–34",
      "35–39",
      "40–44",
      "45–49",
      "50–54",
      "55–59",
      "60–64",
      "65–69",
      "70–74",
      "75–79",
      "80–84",
      "85 and over",
      "Total Persons"
    )
  ),
  # this tells us to rename the columns nicely
  .name_repair = janitor::make_clean_names
) %>%
  # we only want to look at the state "Victoria"
  filter(s_t_name == "Victoria") %>%
  # We need to do some fancy string manipulation here
  # This will involve removing everything contained inside brackets
  # these are called "regular expression" - we haven't learnt about them yet
  # ut they are a good thing to know aout
  mutate(lga_name = str_remove(
    string = lga_name,
    pattern = "\\([^()]*\\)"
  ) %>%
    str_remove("\\(.*?\\)") %>%
    # str_trim is a handy function that trims off excess white space
    # from a string. So " Victoria " becomes "Victoria"
    str_trim()) %>%
  # we don't need these variables
  select(-s_t_code,
         -s_t_name,
         -lga_code)
```


# **Put the data, `demographics_raw`, into Tidy Data form. Describe the steps you will need to take to read in the ABS data and tidy it. The end product should probably look something like this:** (1 marks)

```
   s_t_name lga_code lga_name   total_persons age_group age_group_count
   <chr>       <dbl> <chr>              <dbl> <chr>               <dbl>
 1 Victoria    20110 Alpine             12639 0_4                   553
 2 Victoria    20260 Ararat             11757 0_4                   602
 3 Victoria    20570 Ballarat          105438 0_4                  7038
 4 Victoria    20660 Banyule           129125 0_4                  8448
 5 Victoria    20740 Bass Coast         34390 0_4                  1879
```

|s_t_name | lga_code|lga_name   | total_persons|age_group | age_group_count|
|:--------|--------:|:----------|-------------:|:---------|---------------:|
|Victoria |    20110|Alpine     |         12639|0_4       |             553|
|Victoria |    20260|Ararat     |         11757|0_4       |             602|
|Victoria |    20570|Ballarat   |        105438|0_4       |            7038|
|Victoria |    20660|Banyule    |        129125|0_4       |            8448|
|Victoria |    20740|Bass Coast |         34390|0_4       |            1879|


```{r tidy-data}
demographics_vic <- demographics_raw %>%
  # we want to put the data into a longer format, 
  gather(
    # which variales do we want to gather?
    ___
    # we are putting information into two columns
    key = ___,
    value = ___
  ) %>%
  # this removes any characters "x"  from the age_group variable
  mutate(age_group = str_remove(age_group, "x"))

demographics_vic
```

> Your answer here...

# Choose an LGA of your choice (no criteria - just choose one!) to explore in addition to Monash. **Important - you will choose only these two LGAs to explore for (almost) every other part of the assignment**

```{r burg-demog-lgas}
demo_subset <- demographics_vic %>%
  filter(lga_name %in% c("Monash", ___"your chosen LGA goes here"___))
```

# `**` **How do the age groups of the two LGAs differ? Draw a plot (or create a table) to show how the LGAs differ and write a few sentences discussing the result.** `**` (2 Marks).

```{r age-lgas}

# IF you want to make a table, here is some code to help with that
demo_subset %>%
  # what variables do we want to group by?
  group_by(___,
           ___) %>%
  summarise(n = sum(age_group_count)) %>%
  mutate(freq = n / sum(n)) %>%
  select(-n) %>%
  spread(key = age_group,
         value = freq)

# We want to make a column (barplot) of age group and the count for each,
# filling for LGA name
ggplot(demo_subset,
       aes(x = ___,
           y = ___,
           fill = ___)) + 
  geom_col(position = "dodge")

# or try a different plot
ggplot(___,
       aes(x = ___,
           y = ___,
           fill = ___)) + 
  geom_col() + 
  facet_wrap(~lga_name,
             ncol = 1) +
  theme(legend.position = "none")
```

>  ...

# Amelia: Let's read in the burglary entry point data, only using the residential burglaries

```{r read-crime}
# Only keep residential burglaries
burglary_entry <- read_excel(
  "data-raw/Data_tables_spotlight_burglary_break_and_enter_visualisation_year_ending_December_2018_v3.xlsx",
  sheet = 3
) %>%
  # remove Non-residential or "Unknown" from the `Offence Sugroup` column
  filter(str_detect(string = `Offence Subgroup`, 
                    pattern = "Non-residential", 
                    negate = TRUE)) %>%
  filter(str_detect(string = `Offence Subgroup`,
                    pattern = "Unknown", 
                    negate = TRUE)) %>%
  # give the variables better names
  rename(
    year = `Year ending December`,
    local_gov_area = `Local Government Area`,
    offence_subgroup = `Offence Subgroup`,
    building_entry_point = `Building Point of Entry`,
    offence_count = `Offence Count`
  )
```

# `**` **Subset your data again to have the same two LGAs you previously explored. What are the two most, and two least common methods of entry into houses? Does this vary for your two chosen LGAs? Draw a plot (or make a table) and write a few sentences discussing the result.** `**` (2 Marks)

```{r entry-frequencies}
burglary_subset <- burglary_entry %>%
  filter(local_gov_area %in% c("Monash", **your chosen LGA here***))

# we want to calculate the number of crimes, grouped y each building entry point
burglary_summary <- burglary_subset %>%
  group_by(___) %>%
  count(___)

# This is an example table you could look at if you want to 
# compare exact numbers
burglary_summary %>%
  spread(key = ___,
         value = ___)

# This is a plot of the numer of building entry points used in crimes, for each local gov area.
ggplot(___,
       aes(x = reorder(___, n),
           y = ___,
           fill = ___)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs() +
  # Amelia: this makes the plots colourlind friendly - a good idea to do 
  # this for your plots
  scale_fill_brewer(palette = "Dark2")
```

> ...


# `**` **We want to look at how method of entry is related to age groups in an LGA. To do this, we need to join the crime data (entry data only) with the demographics data. Write a sentence explaining how you joined the data.** `**` (1 Mark)


```{r join-crime}
burglary_demog <- burglary_subset %>%
  left_join(demo_subset,
            by = c("local_gov_area" = "lga_name"))
```

# We need to look into the times of break ins - let's read in the burlgary time data.

```{r burglary-times-read}
# Only keep residential burglaries
burglary_time <- read_excel(
  "data-raw/Data_tables_spotlight_burglary_break_and_enter_visualisation_year_ending_December_2018_v3.xlsx",
  sheet = 7
) %>%
  # remove entries with "Non-residential" in `Offence Sugroup`
  filter(str_detect(
    string = `Offence Subgroup`,
    pattern = "Non-residential",
    negate = TRUE
  )) %>%
  # remove entries with "Unknown" in `Offence Sugroup`
  filter(str_detect(
    string = `Offence Subgroup`,
    pattern = "Unknown",
    negate = TRUE
  )) %>%
  rename(
    year = `Year ending December`,
    local_gov_area = `Local Government Area`,
    offence_subgroup = `Offence Subgroup`,
    day = `Commit From Day`,
    time = `Commit From Time`,
    offence_count = `Offence Count`
  ) %>%
  # Here we are creating factor information for each of the times
  # so that they are ordered properly and have good names
  mutate(
    time = fct_relevel(
      time,
      "12am to 5:59am",
      "6am to 11:59am",
      "12pm to 5:59pm",
      "6pm to 11:59pm"
    ),
    day = fct_relevel(
      day,
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  )
```

# `**` **What are the most common times to break into houses? Draw a plot )or create a table) to show how the LGAs differ and write a few sentences discussing the result. Write at least three sentences about this.** `**` (2 Marks)

```{r burglary-times}
# We want to select the burglary times to be the LGA that was Monash plus
# the other one we selected.
burglary_time_my_lgas <- burglary_time %>%
  filter(local_gov_area %in% c("Monash", **your chosen lga**))

# perform a summary that counts the number of offences by each time, in a day, for a local government area
burglary_time_my_lga_summary <- burglary_time_my_lgas %>%
  group_by(local_gov_area,
           ___,
           ___) %>%
  summarise(n = sum(___))

# Make a plot that looks at the time of day and the count of crimes
# facetted in a grid by each day and local government area
ggplot(___,
  aes(x = ___,
      y = ___)) +
  geom_col() +
  facet_grid(___ ~ ___)
```

# `**` **Is there a higher proportion of burglaries in LGAs of a certain age group?  Produce a plot (and/or table) and write about three sentences about this.** `**` (2 Marks)

```{r join-burglary-w-lga-data}
# join the burglary data with the lga age data to make this plot
```


# `**` **Go to the ABS website (https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1410.02013-18?OpenDocument) and find some data on income for each LGA. Read the data in, put it into tidy format (Hint: Look at how the previous ABS dataset was read into R)** `**` (3 Marks)

```{r read-abs-income}
# Hint - the ABS data has a similar structure for most of its datasets
# This means that you can re-use the code from earlier here
```

# `**` **Join this ABS income data from all LGAs to the data with burlgary times and LGA information to answer the next question. Describe how you joined the data** `**` (1 marks)

```{r join-abs-burglary}

```


# `**` **Which income areas (if any) are being targeted? If it is higher income places, maybe we can target them with some of our premium products? What do you think? Does a plot help confirm / deny this story?** `**` (2 marks)

```{r which-areas-targetted}

```

