---
title: "Assignment 1"
author: "MNS2"
date: "13 March 2018"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
```

## Tidy The Date
### Produce a data set with Following Columns
#### * Airport
#### * Year
#### * Type of Flight: Domestic or International
#### * Bound: IN or OUT

```{r}
passengers <- read_xls("data/WebAirport_FY_1986-2017.xls", sheet=3, skip=6)
passengers_tidy <- passengers %>%
  gather('INBOUND', 'OUTBOUND', 'INBOUND__1', 'OUTBOUND__1', 'INBOUND__2', 'OUTBOUND__2', key = 'BOUND', value = 'RevenuePassengers') %>%
  separate(BOUND, c('BOUND', 'type'), fill = 'right') %>%
  mutate(type = replace_na(type, 'Domestic')) %>% 
  mutate(type = sub('1', 'International', type)) %>% 
  mutate(type = sub('2', 'Total', type)) %>%
  select(AIRPORT, Year, BOUND, type, RevenuePassengers) %>% 
  complete(AIRPORT) %>% 
  filter(AIRPORT != 'TOTAL AUSTRALIA') %>%
  filter(type != 'Total')

print(passengers_tidy)
```

## Barchart: Total Passengers Facetted by Year


```{r fig.width=15, fig.height=15}
ggplot(passengers_tidy, aes(x = Year, fill = type)) +
  geom_bar(aes(weight = RevenuePassengers)) +
  labs(title = "Domestic and Internationl Flight Passenger Revenue", x = "Year", y = "Revenue Passengers") + theme(axis.text.x = element_text(angle= 90, hjust = 1), axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
```

## Spreading the Data for Scatter Plots

```{r}
passengers_tidy_spread <- passengers_tidy %>% spread(key=BOUND, value=RevenuePassengers)
passengers_tidy_spread
```

## Scatterplot: Domestic Inbound vs Outbound Passengers

```{r fig.width=15, fig.height=15}
passengers_tidy_dom <- passengers_tidy_spread %>% filter(type == "Domestic")
ggplot(data = passengers_tidy_dom) + geom_point(mapping = aes(x=INBOUND, y=OUTBOUND), alpha=0.3) + facet_wrap(~ Year) + coord_fixed(ratio = 1) + xlab('Inbound Passengers') + ylab('Outbound Passengers') + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
```

## Scatterplot: International Inbound vs Outbound Passengers

```{r fig.width=15, fig.height=15}
passengers_tidy_int <- passengers_tidy_spread %>% filter(type=="International")
ggplot(data = passengers_tidy_int, aes(x=INBOUND,y=OUTBOUND)) + geom_point(alpha = 0.3) + facet_wrap(~Year) + coord_fixed(ratio = 1) + xlab('Inbound Passengers') + ylab('Outbound Passengers') + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
```

## What is learned from each plot:

All three graphs indicate a clear increase in the number of flights to and from Australian airports between 1985 and 2017. The total number of passengers increased every year except for in 89-90 and 01-02. The effect of the pilots' dispute in 1989-1990 is clear shown by the decrease in the number of domestic passengers transported. This did not seem to have any impact on the number of international passengers. The drop in 2001-2002 was likely caused by the 9/11 terroist attacks and the collapse of Ansett Australia.

Sydney, Melbourne and Brisbane were the three busiest airports for every year. 
For both international and domestic flights at all airports the number of inbound and outbound passengers were roughly the same.
