---
title:
author: 
output: html_document
---

First load up the tidyverse

```{r library}
library(tidyverse)
```

Then read in the data, `rates.csv` and name it `rates`
```{r read-rates}
___ <- read_csv("data/___")

head(rates)
```

Now plot the data:

- on the x axis put `date`
- on the y azis put `AUD`

```{r plot-rates}
ggplot(rates, 
       aes(x = ___, 
           y = ___)) + 
  geom_line()
```

Make a subset of the data that includes the following columns:


- `date`
- `AUD`
- `GBP`
- `JPY`
- `CNY`
- `CAD`

```{r subset-rates}
rates_sub <- rates %>% 
  select(___, 
         ___, 
         ___, 
         ___, 
         ___, 
         ___)

rates_sub
```


Now put the data into long form, where:

* `key` is `currency`
* `value` is `rate`

```{r rates-long}
rates_long <- rates_sub %>%
  gather(key = ___, 
         value = ___, 
         -date)
```

Now plot the currency rates

* with `date` on the x axis
* with `rate` on the y axis
* And faceted by each currency

```{r plot-all-currencies, fig.width=6, fig.height=10}

ggplot(rates_long, 
       aes(x = ___, 
           y = ___)) + 
  geom_line() +
  facet_wrap(~currency, # facetted by currency
             ncol = 1, 
             scales = "free_y")
```

* What do you learn?

Describe the similarities and differences between the currencies. 
