# Demonstrate with koala bilby data (write this as live code)

Here is a little data set to practice gather, spread and separate on.

First we load the library `tidyverse` to load up all of our functions.

```{r library}
library(tidyverse)
```


```{r koala-bilby}
kb <- read_csv("data/koala_bilby.csv")
kb
```

---

### Exercise 1

Gather the data into long form, naming the two new variables, `label` and `count`

```{r kb-long}
kb_long <- kb %>% 
  gather(key = label, 
         value = count, 
         -ID)

kb_long
```

### Exercise 2

Separate the labels into two new variables, `animal`, `state`

```{r kb-separate}
kb_long_sep <- kb_long %>%
  separate(col = label, 
           into = c(__, ___))

kb_long_sep
```

### Exercise 3

Spread the long form data into wide form, where the columns are `state`s. 
```{r kb-wide-state}
kb_long_sep %>% 
  spread(key = ___, 
         value = count)
```

### Exercise 4

Spread the long form data into wide form, where the columns are the `animal`s. 

```{r kb-spread-animal}
kb_long_sep %>% 
  spread(key = ___, 
         value = count)
```

---
