---
title: "Rude Recliners"
author: 
output: html_document
---

This contains data from the article [41% Of Fliers Think You're Rude If You Recline Your Seat](http://fivethirtyeight.com/datalab/airplane-etiquette-recline-seat/). 

V1 is the response to question: "Is it rude to recline your seat on a plane?"

V2 is the response to question: "Do you ever recline your seat when you fly?". 

# Exercise 1: Rude Recliners 

Answer the following questions in the rmarkdown document.

First, let's read in the data

```{r read-in}
library(tidyverse)

recliners <- read_csv("data/recliners.csv")
recliners
```

## Ex1 A) What are the variables and observations in this data?

* The variables are ...
* The observations are ...

## Ex 1B) Put the data in tidy long form (using the names `V2` as the key variable, and `count` as the value). 

```{r recline-long}

recline_long <- recliners %>% gather(key = ___, 
                                     value = ___,
                                     -V1)
```

## Ex 1C) Use the `rename` function to better describe the variable names.

```{r rename-recline}
recline_long %>%
  rename(V1 = "response",
         V2 = "question")
```

