---
title:
author:
output: html_document
---

- Write down the steps that will be needed go from raw data to tidy
- Save a new copy of the data called `tb_raw`, which contains the read in data.
- Print the data

First, load `library(tidyverse)`, and read in the data

```{r tb-read}
library(tidyverse)
tb_raw <- read_csv("data/TB_notifications_2018-03-18.csv")
tb_raw
```

- select the variables `country`, `year`, and those that start with `new_sp_`. See `?select` for more information
- then filter the observations to be between 1996 and 2012.
- Save this shorterned dataset as `tb_short`.


```{r tb-short}
tb_short <-  tb_raw %>%
  select(country, # select variables country and year
         year, 
         starts_with("new_sp_")) %>%
  filter(year > 1996, 
         year < 2012) # before 2012
```

- Identify the top 20 observations using `top_n`

```{r tb-top}
tb_short %>% top_n(20) # place 20 inside top_n
```

`gather` the data into columns into columns named `stuff` and `count`

```{r tb-gather}
tb_gather <- tb_short %>% 
  gather(key = stuff, 
         value = count, 
         starts_with("new_sp_"))

tb_gather
```

- Separate the column `stuff`,  `into` three columns named "stuff1", "stuff2", and "genderage"

```{r tb-sep1}
tb_sep_step1 <- tb_gather %>%
  separate(col = stuff, 
           into = c("stuff1", "stuff2", "genderage"))

tb_sep_step1
```

- Separate `genderage` into two columns named "gender" and "age", separating on the first position.

```{r tb-sep2}
tb_sep_step2 <- tb_sep_step1 %>%
  separate(col = genderage, 
           into = c("gender", "age"), 
           sep = 1)
```

Drop the columns named stuff1 and stuff2

```{r tb-tidy}
tb_tidy <- tb_sep_step2 %>%
  select(- stuff1, 
         - stuff2) # drop stuff1 and stuff2

```
