---
title: "Assignment 1"
output: html_document
author: "The pirates"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr);
library(dplyr);
library(tidyverse);
library(readxl)
passengers <- read_xls("data/WebAirport_FY_1986-2017.xls", sheet=3, skip=6) %>% filter(!is.na(INBOUND)) #gets data
passengers

```

```{r}
#Data is first split into domestic and international
myvars <- c("AIRPORT", "Year", "INBOUND", "OUTBOUND")

passengers_domestic <- passengers[myvars]

myvars <- c("AIRPORT", "Year", "INBOUND__1", "OUTBOUND__1")
  passengers_international <- passengers[myvars]

#Add a column to each set for type of flight
passengers_domestic['Type'] = 'Domestic'
passengers_international['Type'] = 'International'

#Rename columns to desired names
passengers_domestic <- plyr::rename(passengers_domestic, c("INBOUND"="IN", "OUTBOUND" ="OUT"))
passengers_international <- plyr::rename(passengers_international, c("INBOUND__1"="IN", "OUTBOUND__1"="OUT"))

#combine sets into one
passengers_new <- rbind(passengers_domestic, passengers_international)
 
passengers_new
```
```{r, fig.width =18, fig.height =10}
#Part 2 
#Use your data to make
    #a barchart of type of flight facetted by year.
    #a scatterplot of domestic inbound vs outbound passengers facetted by year. There are three airports with much higher numbers, which are these three?
    #a scatterplot of international inbound vs outbound passengers facetted by year

####This is for a column graph, left the code here in case we cant to use it.  Looks the same as the bar graph.  Assignment seems to want the bar graph though#######
#ggplot(passengers_new, aes(x=Type,y=IN+OUT, fill= Type))+
  #geom_col() + facet_wrap(~Year, nrow=8)+ scale_y_continuous(labels=scales::comma)


ggplot(passengers_new, aes(x=Type, fill= Type))+
  geom_bar(aes(weight=IN+OUT)) + facet_wrap(~Year, nrow=8)+ scale_y_continuous(labels=scales::comma)

#####removing Total Australia from data set########
passengers_domestic2 <- passengers_domestic[!passengers_domestic$AIRPORT=='TOTAL AUSTRALIA',];

ggplot(passengers_domestic2, aes(x=IN,y=OUT, colour = AIRPORT)) + geom_point() + facet_wrap(~Year,nrow=10)+ 
  scale_y_continuous("Out(10mil)",breaks=seq(0,10000000,5000000), labels = seq(0,10,5))+ scale_x_continuous("IN(10mil)",breaks=seq(0,10000000,5000000), labels = seq(0,10,5))
  #scale_y_continuous(labels=scales::comma) + #scale_x_continuous(labels=scales::comma)
#####Three airports with much higher numbers are sydney, melbourne and brisbane######

########removing Total Australia from data set########
passengers_international2 <- passengers_international[!passengers_domestic$AIRPORT=='TOTAL AUSTRALIA',];

ggplot(passengers_international2, aes(x=IN,y=OUT,colour = AIRPORT)) + geom_point() + facet_wrap(~Year,nrow=8)+scale_y_continuous("Out(10mil)",breaks=seq(0,10000000,5000000), labels = seq(0,10,5))+ scale_x_continuous("IN(10mil)",breaks=seq(0,10000000,5000000), labels = seq(0,10,5))

#scale_y_continuous(labels=scales::comma) + #scale_x_continuous(labels=scales::comma)

```


Graph 1 plots the total number of Inbound and Outbound airline passengers against the type of flight; which is domestic or international. A steady upward trend of the total number of inbound and outbound passengers for domestic flights can be seen, with it increasing from about 60mil passengers in 1985-86 to over 200mil passengers in 2016-17.

The same can be said for the trend of the number of inbound and outbound passengers for international flights as it shows an increasing one; albeit being slightly lower in overall numbers, the chart shows it  increasing from about 20mil passengers in 1985-86 to over 60mil passengers in 2016-17.Overall, there has been an increasing trend in number of airline passengers altogether, technological advancements may have attributed to this, making it easier to fly in 2016-17 compared to 1985-86. Domestic flight totals has also been greater than international may be because international flights tend to be for summer or winter breaks as a means of having a vacation.


From graph 2, We plotted the total inbound flights against outbound flights domestically. The three airports with much higher numbers are Sydney, Melbourne and Brisbane. We plotted the total inbound flights against outbound flights domestically. Sydney, Melbourne and Brisbane have much higher totals for domestics flights, compared to other cities and this may be due to denser population numbers in these 3 areas.

Similarly, according to graph 3, we plotted the total inbound flights against outbound flights but internationally. The three airports with much higher numbers were also Sydney, Melbourne and Brisbane, though with a lower total numbers for international flights compared to domestic. Comparing the two, Local travelling required by respective jobs may attribute to domestic totals being greater than international totals.


