---
title: "Black 4"
output: html_document
---
#Tidy data

```{r}
library(tidyverse)
library(readr)
survey <- read.csv("data/survey.csv")
survey <- survey[-64,]

#1
survey$Q1 <- str_replace_all(survey$Q1,c("A"="1","B"="2","C"="3","D"="4","E"="Other"))

#2
survey$Q2 <- survey$Q2 %>% str_replace_all(c("1/2"=0.5,"three times"="","4/7"=4/7)) 
survey$Q2 <- survey$Q2 %>% str_remove_all("[HhoOuUrRs()touni]")
survey$Q2 <- as.numeric(survey$Q2)

#3
survey$Q3 <- as.logical(survey$Q3)

#4
survey$Q4 <- str_to_lower(str_sub(survey$Q4, 1, 20))
survey$Q4 <- str_replace_all(survey$Q4,c("-"=" ","alt rock"="alternative rock","rnb"="r&b","popular"="pop","rock and/or roll"="rock","pop music"="pop","country music"="country"))
survey$Q4 <- str_replace_all(survey$Q4,c("hip pop"="hip hop","hiphop"="hip hop","movie soundtracks\n"="movie soundtracks","light\n"="light"))
survey$Q4[46] <- NA
survey$Q4[22] <- NA
survey$Q4 <- as.factor(survey$Q4)

#5
survey$Q5 <- str_remove_all(survey$Q5,"[HhOouUuRrsS]")
survey$Q5 <- as.integer(survey$Q5)

#6
survey$Q6 <- str_to_lower(str_sub(survey$Q6, 1, 50)) 
survey$Q6 <- survey$Q6 %>% str_replace_all(c("fly"="flight","flying"="flight","flighting"="flight", "go back to a past time"="time travel","predict the future"="forecasting","forecast the future"="forecasting","invisibility"="invisible","can go to wherever i want in any time"="time travel","forecasting\n"="forecasting"))
survey$Q6 <- survey$Q6 %>% str_replace_all(c("time control"="control time","time-reverse"="control time","freeze time"="control time","to never have to sleep"="not needing sleep","teleporting"="teleportation","time travel in any time"="control time"))
survey$Q6[which(survey$Q6=="time")] <- "control time"
survey$Q6[which(survey$Q6=="can go to wherever i want")] <- "time travel"
survey$Q6 <- as.factor(survey$Q6)

#9
survey$Q9 <- str_to_lower(str_sub(survey$Q9, 1, 50))
survey$Q9[c(33,57,62,40,31,16,50,21,38,41,23,19,7,59,32,37,55)] <- "none"
survey$Q9[3] <- NA
survey$Q9 <- str_replace_all(survey$Q9,c("foxtel..."="foxtel","netlfix"="netflix","stream.tv\n"="stream.tv"))
survey$Q9 <- as.factor(survey$Q9)

#10
survey$Q10 <- str_remove_all(survey$Q10,"[?\n]")
survey$Q10 <- as.numeric(survey$Q10)

#13
survey$Q13 <- str_remove_all(survey$Q13,"[dollars$]")
survey$Q13 <- as.numeric(survey$Q13)

#17
survey$Q17 <- str_replace_all(survey$Q17,c("A"="YES","B"="NO"))
survey$Q17 <- as.factor(survey$Q17)

#18
survey$Q18 <- str_to_lower(str_sub(survey$Q18, 1, 50))
survey$Q18 <- as.factor(survey$Q18)

#20
survey$Q20 <- as.logical(survey$Q20)

#21
survey$Q21 <- str_remove_all(survey$Q21,"[\n]")
survey$Q21[59] <- 7.5
survey$Q21 <- as.numeric(survey$Q21)

#22
survey$Q22[59] <- 3
survey$Q22 <- as.numeric(substr(survey$Q22,1,3))

#23
survey$Q23[59] <- 21
survey$Q23[63] <- NA
survey$Q23 <- as.numeric(substr(survey$Q23,1,3))

#24
survey$Q24[38] <- NA
survey$Q24 <- as.integer(substr(survey$Q24,1,3))

#25
survey$Q25 <- str_to_lower(str_sub(survey$Q25, 1, 50))
survey$Q25[40] <- NA
survey$Q25[32] <- NA
survey$Q25 <- str_replace_all(survey$Q25,"swimming\n","swimming")
survey$Q25[which(survey$Q25=="swim")]="swimming"
survey$Q25 <- as.factor(survey$Q25)

#26
survey$Q26 <- str_replace_all(survey$Q26,c("A"="Female","B"="Male","C"="prefer not to answer"))
survey$Q26 <- as.factor(survey$Q26)

#30
survey$Q30 <- str_replace_all(survey$Q30,c("B"="YES","A"="NO"))
survey$Q30 <- as.factor(survey$Q30)

#31
survey$Q31 <- str_replace_all(survey$Q31,c("A"="Do like ETC1010","B"="Doesn't like ETC1010"))
survey$Q31 <- as.factor(survey$Q31)

#32
survey$Q32 <- str_to_lower(str_sub(survey$Q32, 1, 50))
survey$Q32 <- str_remove_all(survey$Q32,"[()]")
survey$Q32 <- as.factor(survey$Q32)

#33
survey$Q33 <- as.logical(survey$Q33)

write_csv(survey,"data/Black4_tidy up.csv")                          
```

#summary plots 

```{r}
ggplot(survey) + 
  geom_histogram(aes(x=Q2),rm.na=TRUE) +
  xlab("How much time do you spend commuting to university each week?") 
```

The histogram shows how much time students in our class spent on commuting each week. The x-axis represents the time they spent and the y-axis represents the percentage of students. From this diagram, we can see that there are around 24 students who spent on less than 2 hours commuting each week, which implied that almost one third of students in our class live relatively close to Monash clayton campus. In addition, there are more than a quarter of students who live relatively far from Monash clayton campus because they need to spend at least 6 hours on commuting each week. One thing to notice is that the value of 120 in this diagram might be an accidental mistake because it is impossible to spend 120 hours each week on commuting. It is likely a joke. In conclusion, it is convenient for most of students in our class to go to school. 

```{r}
ggplot(survey,aes(x="",y=1/63*100,fill=Q1)) +
  geom_bar(stat="identity",width = 1) +
  coord_polar("y", start=0) +
  xlab("percentage") +
  ylab("")+
  labs(fill="Year of uni")
```

The pie chart shows the proportion that number of students from our class make up for each year of unit. Different colors represent different year of unit. From this chart, we can see that most of students in our class are year-2 or year-3 students, which accounts for approximately 35% and 30% respectively. Besides, we also find that the percentage of student in ¡°other¡± year of unit are greater than the sum of the percentage of year-1 and year-4 students, which accounts for around 20%. In conclusion, it is obvious that few students will choose to learn this unit when they are in first year.

```{r}
ggplot(survey,aes(Q26,Q24)) + 
  geom_boxplot(na.rm=TRUE) +
  xlab("Gender") +
  ylab("Height") 
```

The boxplot chart shows the distribution of the relationship between gender and height. From this chart, we can see that the median of female height is around 1.63 m and the median of male height is around 1.78 m. We also find that there are some outliers from this chart. For example, the outlier of female height is 1.78 m, which is equal to the median of male height. Besides, there are also 5 outliers in the male height, which they are over 1.55 m,1.93 m 1.96 m,2.00m and 2.01m. The difference between the maximum and the minimum is 0.46 m. In conclusion, males in our class are taller than females.

```{r}
ggplot(survey,aes(x=Q26,y=Q21)) + 
  geom_boxplot(na.rm=TRUE) + 
  xlab("Gender") +
  ylab("time of sleep per day(H)")
```

The boxplot chart shows the distribution of the relationship between gender and time of sleep. It shows that the median of female¡¯s time of sleep is approximately 7.3h, which is slightly more than the median of male¡¯s time of sleep (7h). Besides, there are some extreme values in this chart. For example, the minimum of female¡¯s time of sleep is only 5 hours each day and the minimum of female¡¯s time of sleep is only 4 hours each day. In conclusion, it is hard to see whether gender and time of sleep has a relationship because the amount of the data is too small. 

```{r}
ggplot(survey,aes(Q17,fill=Q30)) + 
  geom_bar(na.rm=TRUE) + 
  xlab("Do you have prior coding experience?") + 
  labs(fill="Are you struggling with this unit?") +
  theme(legend.position="top") +
  facet_grid(.~Q31)
```

At first, our group members think "Do you like ETC1010", "Are you struggling with ETC1010" and "Do you have prior coding experience" are the questions which have the main method with this unit. So we put these three questions together. 
In this graph, we want to use the bar chart to show the possible relationship for these three question. From this graph, we can find out that most people who like ETC1010 have prior coding experience. And by the colour of the graph, we can find out that the most people who have the experience of prior coding aren't struggling with the unit ETC1010. In the other hand, the most people who don't like the ETC1010 are doesn't have prior coding experience. And all of them are struggling with this unit.
From all of the information we can gain from the graph, we can get the interesting relationship between these three question is the people  who have the prior coding experience may think this unit is not too difficult for them and they will be more like the ETC1010 then the people who didn't have any prior coding experience. In the other hand, all people who don't like this unit and feel the unit is hard are the people who didn't have any prior coding experience. So we can gain the conclusion is although ETC1010 is a primary unit for data and computing this unit still needs some prior coding experience.


```{r}
ggplot(survey,aes(Q1,Q21)) + 
  geom_boxplot(na.rm=TRUE) + 
  xlab("year of uni") +
  ylab("time of sleep per day(H)") 
```

All of us think that the more large year of university the less sleep hour they can get. Since we had drawn the box-plot graph to show the situation of our class. From the graph, we can get that the year one student, year two students and year three students have a very same median long for sleeping. But for Year 3 and Year Other we can find out the year level is higher and the more time they will sleep. That is interesting. Because before we drawn the graph we think the year level is higher they will become more hard-working and they will have less time for sleeping. But we find that the situation is not as our expectation. Maybe the university students who had learned level year 4 and level other can manage their time more easily than the level year 1,2,3.

```{r}
ggplot(survey,aes(Q26,fill=Q17)) + 
  geom_bar(na.rm=TRUE,position="fill") +
  xlab("Gender") +
  ylab("Position") +
  labs(fill="Do you have prior coding experience?") +
  theme(legend.position="top")
```

From the bar chart, we can find that between the female and the male, a little bit more male students who have the prior coding experience than the female students who have the prior experience. And all the students who prefer not to answer have prior coding experience.

```{r}
ggplot(survey,aes(Q23,Q24,colour=Q1))+
  geom_jitter(alpha=1,size=3,na.rm=TRUE)  +
  geom_smooth(se=FALSE,method="lm",na.rm=TRUE) +
  xlab("Age") +
  ylab("Height") + 
  labs(colour="Year of uni")
```

At first, we just wanted to put the Age and Height into one Scatter graph. But we had found that the most point is landed at the left of the age 25. But we consider about that the sample is the 64 students in our class. Because we are the university students so the most of us are at the age between 18 to 25.  Although this is not a huge sample, we still want to gain some interesting information about our classmate. So we put the year of university as our third data in this graph. By this action, we had found that the most year two have the age between 18 to 22.5, and the height of them are around 155 to 180. And by the regression line, we can also find that the trend of the height and age for student year 2 is the elder student also have a higher height. For the year 3 students, the age for them are between 19 to 22.5, and most of them are around 20 years old.  We can find out that the highest student is at this year level, and we can also find out that the year 3 student has the situation is the older student will have a higher height. The sample for year 1 and year 3 is very small so we cannot give some interesting conclusion. 
But we can also find out that the only three years 3 students have a very similar height. 
The final year level is year other. From this graph, we can find out that the age for the students which are year other is between 23 to 28. And the height of them also can be the older students have a higher height. And the smallest height for the year other students is higher than all another year level. 
By all of this, we can gain the conclusion although we can say the older students have the higher height, we didn't have a very large sample. So we can just say in our class the year 2,3,4 and another student have the situation is the older students have the higher height.


