
---
title: "ETC 1010 Assignment 2"
author: "5T"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 8,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE
)
```


```{r}
library(tidyverse)
library(ggplot2)
survey <- read_csv("data/survey.csv",
                    na = c("n/a", "NA")) %>%
  filter(id != 64)
#tidysurvey <- read_csv("~/Documents/Uni/ETC1010/Assignment 2/survey_tidy.csv")
```

#Recode Q1 to have answers 1,2,3,4, Other
```{r}
library(dplyr)
survey$Q1[survey$Q1=="A"] <- 1
survey$Q1[survey$Q1=="B"] <- 2
survey$Q1[survey$Q1=="C"] <- 3
survey$Q1[survey$Q1=="D"] <- 4
survey$Q1[survey$Q1=="E"] <- "Other"
ggplot(data = survey, aes(x=Q1)) +
  geom_bar()
```
The majority of the class are in second year, wth third years closely following. Interestingly, first years are the minority even though this is a first year unit. This could be because ETC1000 is a prerequisite, so many first years will be ineligible to take this unit in this semester. 

#Cleaning up Q2
```{r}
library(stringr) 
survey <- survey %>%
  #change 1/2 to 0.5
  mutate(Q2 = sub("1/2", "0.5", Q2)) %>%
  #remove hours
  mutate(Q2 = sub(" hours", "", Q2)) %>%
  #remobe \n
  mutate(Q2 = sub("\n", "", Q2)) %>%
  #remove hrs
  mutate(Q2 = sub("hrs", "", Q2)) %>%
  #remove HOUR
  mutate(Q2 = sub("HOUR", "", Q2)) %>%
  #change 4/7 to 0.57
  mutate(Q2 = sub("4/7", "0.57", Q2)) %>%
  #remove three times
  mutate(Q2 = sub("three times", "", Q2))
#this changes the value in Q2 on line 27 to "2.0"
survey$Q2[27] <- "2.0"
#this mutates all the values in Q2 to be numbers
survey <- survey %>% mutate(Q2 = as.numeric(Q2))
ggplot(survey, aes(x=Q2)) + geom_histogram()
```
The majority of students takes a short time to commute to uni with the highest being 120 hours. 

Cleaning up Q3:
```{r}
ggplot(data = survey, aes(x=Q3)) +
  geom_bar()
```
The majority of the class did not attend Ed Sheeran's concert. 

Cleaning up Q4:
```{r}
survey <- survey %>%
  #changing alt rock to Alternative Rock
  mutate(Q4 = sub("alt rock", "Alternative Rock", Q4)) %>%
  #Changing classical to Classical
  mutate(Q4 = sub("classical", "Classical", Q4)) %>%
  #change hip hop variations to hip hop
  mutate(Q4 = sub("Hip hop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("Hip Hop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip-hop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip-pop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("Hiphop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip Pop", "hip hop", Q4)) %>%
  #change jazz to Jazz
  mutate(Q4 = sub("jazz", "Jazz", Q4)) %>%
  #change pop music variation to Pop
  mutate(Q4 = sub("pop", "Pop", Q4)) %>%
  mutate(Q4 = sub("pop music", "Pop", Q4)) %>%
  mutate(Q4 = sub("Pop music", "Pop", Q4)) %>%
  #changing R&B variations to R&B
  mutate(Q4 = sub("rnb", "R&B", Q4)) %>%
  mutate(Q4 = sub("RnB", "R&B", Q4)) %>%
  mutate(Q4 = sub("RNB", "R&B", Q4)) %>%
  #changing Rock and/or Roll to Rock
  mutate(Q4 = sub("Rock and/or Roll", "Rock", Q4))
#change - to NA
survey$Q4[46] <- NA
ggplot(data = survey, aes(x=Q4, na.rm = TRUE)) +
  geom_bar() + coord_flip()
```
The top 3 most listened to genres of music are Pop, R&B and hip hop.

#Cleaning up Q5
```{r}
library(tidyverse)
library(stringr)

survey <- survey %>%
  #"Choosing not to answer" turned into NA
  # remove "\n"
  mutate(Q5 = sub("\n", "", Q5)) %>%
  # remove "hrs"
  mutate(Q5 = sub("hrs", "", Q5)) %>%
  #remove Hours
  mutate(Q5 = sub("Hours", "", Q5)) %>%
    #remove hours
  mutate(Q5 = sub("hours", "", Q5)) %>%
    #remove HOUR
  mutate(Q5 = sub("HOUR", "", Q5))

survey$Q5[27] <- "40.0"
    
survey <- survey %>% mutate(Q5 = as.numeric(Q5))

ggplot(data=survey, aes(x=Q5)) +
  geom_histogram()
```
The majority of students spend around 0 to 30 hours on the internet each week.

#Cleaning up Q6:
```{r}
survey <- survey %>%
  #cleaning up all the responses relating to flying to "flight"
  mutate(Q6 = sub("Flight", "flight", Q6)) %>%
  mutate(Q6 = sub("Flying", "flight", Q6)) %>%
  mutate(Q6 = sub("flying", "flight", Q6)) %>%
  mutate(Q6 = sub("fly", "flight", Q6)) %>%
  mutate(Q6 = sub("Fly", "flight", Q6)) %>%
  #cleaning up all the responses relationg to time travel to "time travel"
  mutate(Q6 = sub("Time travel", "time travel", Q6)) %>%
  mutate(Q6 = sub("go back to a past time", "time travel", Q6)) %>%
  mutate(Q6 = sub("can go to wherever i want in any time", "time travel", Q6)) %>%
  #cleaning up all the responses related to control time to "control time"
  mutate(Q6 = sub("Time-Reverse", "control time", Q6)) %>%
  mutate(Q6 = sub("freeze time", "control time", Q6)) %>%
  mutate(Q6 = sub("Control Time", "control time", Q6)) %>%
  mutate(Q6 = sub("time control", "control time", Q6)) %>%
  #cleaning up all the responses related to forecasting to "forecasting"
  mutate(Q6 = sub("predict the future", "forecasting", Q6)) %>%
  mutate(Q6 = sub("forecast the future", "forecasting", Q6)) %>%
  #cleaning up all the responses relating to insivibility to "invisible"
  mutate(Q6 = sub("Invisible", "invisible", Q6)) %>%
  mutate(Q6 = sub("Invisibility", "invisible", Q6)) %>%
  mutate(Q6 = sub("invisibility", "invisible", Q6)) %>%
  #cleaning up all the responses relating to teleporting to "teleportation"
  mutate(Q6 = sub("teleporting", "teleportation", Q6)) %>%
  mutate(Q6 = sub("Teleportation", "teleportation", Q6))
survey$Q6[survey$Q6=="time"] <- "control time" 
ggplot(data = survey, aes(x=Q6, na.rm = TRUE)) +
  geom_bar() + coord_flip()
```
The top 3 most wanted super powers in the class are flight, invisibility and teleportation. 

#Cleaning up Q7:
```{r}
#changing answer names
survey$Q7[survey$Q7=="A"] <- "two sizes too big"
survey$Q7[survey$Q7=="B"] <- "one size too small"
survey$Q7[survey$Q7=="C"] <- "neither of these"
ggplot(data = survey, aes(x = Q7)) +
  geom_bar()
```
The class prefers either clothes two sizes too big or clothes just right and this is split pretty evenly. 

#Cleaning up Q8:
```{r}
#changing answer names
survey$Q8[survey$Q8=="A"] <- "[0,3)"
survey$Q8[survey$Q8=="B"] <- "[3,6)"
survey$Q8[survey$Q8=="C"] <- "[6,9)"
survey$Q8[survey$Q8=="D"] <- "[9,12)"
survey$Q8[survey$Q8=="E"] <- "More than 12"
ggplot(data = survey, aes(x = Q8)) +
  geom_bar()
```
For the time spent studying, it is interesting as we can see that the majority of students spend more than 12 hours per week studying. 

#Cleaning up Q9:
```{r}
#changing all the responses related to not watching television to "none"
survey <- survey %>%
  mutate(Q9 = sub("no idea", "none", Q9)) %>%
  mutate(Q9 = sub("None", "none", Q9)) %>%
  mutate(Q9 = sub("No television", "none", Q9)) %>%
  mutate(Q9 = sub("I do not watch television", "none", Q9)) %>%
  mutate(Q9 = sub("I don't watch television", "none", Q9)) %>%
  mutate(Q9 = sub("Nothing", "none", Q9)) %>%
  mutate(Q9 = sub("Don't watch television", "none", Q9)) %>%
  mutate(Q9 = sub("I don't", "none", Q9)) %>%
  mutate(Q9 = sub("dont watch", "none", Q9)) %>%
  mutate(Q9 = sub("i don't watch tv", "none", Q9)) %>%
  #mutate(Q9 = sub("non", "none", Q9)) %>%
  mutate(Q9 = sub("dont watch tv", "none", Q9))%>%
  #change all answers related to Netflix to "Netflix"
  mutate(Q9 = sub("netflix", "Netflix", Q9))%>%
  #change all answers related to Foxtel to "Foxtel"
  mutate(Q9 = sub("Foxtel...", "Foxtel", Q9)) %>%
  #change all answers related to Stan to "Stan"
  mutate(Q9 = sub("stan", "Stan", Q9))

survey$Q9[survey$Q9=="non"] <- "none"

ggplot(data = survey, aes(x = Q9)) +
  geom_bar() + coord_flip()
```

#Q10:
```{r}
survey$Q10<-as.numeric(survey$Q10)
ggplot(survey, aes(x = Q10)) +
  geom_histogram(stat = "count")
```
The majority of students only drink between 1 and 2 litres of water. 

#Q11:
```{r}
survey$Q11[survey$Q11=="A"] <- "elective"
survey$Q11[survey$Q11=="B"] <- "core"
ggplot(data = survey, aes(x = Q11)) +
  geom_bar()
```
The majority of students are taking this unit as a core, however the difference is not too large. 

#Q12:
```{r}
survey$Q12[survey$Q12=="A"] <- "single"
survey$Q12[survey$Q12=="B"] <- "double"
ggplot(data = survey, aes(x = Q12)) +
  geom_bar()
```
The majority of students take a double degree, however the difference between these two are also not too large. 

#Q13:
```{r}
#remove dollars
survey <- survey %>%
  mutate(Q13 = sub("dollars", "", Q13))
survey$Q13[survey$Q13=="$30"] <- "30" 
survey$Q13[survey$Q13=="$15"] <- "15"
survey$Q13[survey$Q13=="$35"] <- "35"

survey$Q13<-as.numeric(survey$Q13)
ggplot(data=survey, aes(x=Q13)) +
  geom_histogram()
#survey$Q13a<-as.numeric(gsub("*\\D+\\.\\D*", "", survey$Q13))
#survey$Q13a<-as.numeric(gsub("\\D", "", survey$Q13))
```
Most students use below 50 dollars per week on food each day, however someone uses 300 dollars on average per day at uni. 

#Q14:
```{r}
#we change the responses to the actual number
survey$Q14[survey$Q14=="A"] <- "0"
survey$Q14[survey$Q14=="B"] <- "1"
survey$Q14[survey$Q14=="C"] <- "2"
survey$Q14[survey$Q14=="D"] <- "3"
survey$Q14[survey$Q14=="E"] <- "4"
survey$Q14[survey$Q14=="F"] <- "5"
survey$Q14[survey$Q14=="G"] <- "more than 5"
ggplot(data = survey, aes(x=Q14)) +
  geom_bar()
```
The majority of students do know at least one "Sam" as a friend, however there are also a lot of students who do not know a single "Sam" as a friend.

#Q15.
```{r}
#we change the responses to cucumbers/carrots/neither
survey$Q15[survey$Q15=="A"] <- "cucumbers"
survey$Q15[survey$Q15=="B"] <- "carrots"
survey$Q15[survey$Q15=="C"] <- "neither"
ggplot(data = survey, aes(x=Q15)) +
  geom_bar()
```
The majority of students prefer cucumbers. 

#Q16:
```{r}
#we change the answers to the animals
survey$Q16[survey$Q16=="A"] <- "dog"
survey$Q16[survey$Q16=="B"] <- "cat"
survey$Q16[survey$Q16=="C"] <- "rabbit"
survey$Q16[survey$Q16=="D"] <- "bird"
survey$Q16[survey$Q16=="E"] <- "reptile"
survey$Q16[survey$Q16=="F"] <- "other"
ggplot(data = survey, aes(x=Q16)) +
  geom_bar() + coord_flip()
```
THe majority of students own "other"" animals. We would assume this is so large since it includes students who own no pets. 

#Q17:
```{r}
#we change the answers to yes/no
survey$Q17[survey$Q17=="A"] <- "yes"
survey$Q17[survey$Q17=="B"] <- "no"
ggplot(data = survey, aes(x=Q17)) +
  geom_bar()
```
The majority of students do have coding experience, however the different between students who do and do not have coding experience is also quite small.

#Q18:
```{r}
survey <- survey %>%
  #music related response to "music"
  mutate(Q18 = sub("listen to music", "music", Q18)) %>%
  mutate(Q18 = sub("To listen music", "music", Q18)) %>%
  mutate(Q18 = sub("listen music", "music", Q18)) %>%
  # tv and video related responses to "screen time"
  mutate(Q18 = sub("TV", "screen time", Q18)) %>%
  mutate(Q18 = sub("watch vedio", "screen time", Q18)) %>%
  mutate(Q18 = sub("watch youtube", "screen time", Q18)) %>%
  mutate(Q18 = sub("youtube", "screen time", Q18)) %>%
  mutate(Q18 = sub("watch tv", "screen time", Q18)) %>%
  mutate(Q18 = sub("the office", "screen time", Q18)) %>%
  mutate(Q18 = sub("tv", "screen time", Q18)) %>%
  #gaming related responses to "gaming"
  mutate(Q18 = sub("Play games", "gaming", Q18)) %>%
  mutate(Q18 = sub("video game", "gaming", Q18)) %>%
  mutate(Q18 = sub("Games", "gaming", Q18)) %>%
  mutate(Q18 = sub("computer games", "gaming", Q18)) %>%
  mutate(Q18 = sub("play games", "gaming", Q18)) %>%
  mutate(Q18 = sub("Game", "gaming", Q18)) %>%
  mutate(Q18 = sub("Play video games", "gaming", Q18)) %>%
  #sport related answers to "sport"
  mutate(Q18 = sub("Basketball", "sport", Q18)) %>%
  mutate(Q18 = sub("Gym", "sport", Q18)) %>%
  mutate(Q18 = sub("Sport", "sport", Q18)) %>%
  #sleep to "sleep"
  mutate(Q18 = sub("Sleep", "sleep", Q18))
ggplot(data = survey, aes(x=Q18)) +
  geom_bar() + coord_flip()
```
The top 3 ways students do to take a break from study are gaming, sleeping and screen time. 

#Q19:
```{r}
#change responses to names
survey$Q19[survey$Q19=="A"] <- "Windows"
survey$Q19[survey$Q19=="B"] <- "Mac"
survey$Q19[survey$Q19=="C"] <- "Linux"
survey$Q19[survey$Q19=="D"] <- "Other"
ggplot(data = survey, aes(x=Q19)) +
  geom_bar()
```
THe majority of students have Windows computers.

#Q20:
```{r}
ggplot(data = survey, aes(x=Q20)) +
  geom_bar()
```
The majority of students support same-sex marriage. 

#cleaning up Q21
```{r}
library(tidyverse)
library(stringr)

survey <- survey %>%
  # change "6\n" to "6"
  mutate(Q21 = sub("6\n", "6", Q21)) %>%
  # change "7;30" to "7.5"
  mutate(Q21 = sub("7;30", "7.5", Q21))
#changing the 27th responses to 7
survey$Q21[27] <- "7.0"
#changing all responses to numbers  
survey <- survey %>% mutate(Q21 = as.numeric(Q21))

ggplot(survey, aes(x= Q21)) + 
  geom_bar()
```
The majority of students get between 6 and 8 hours of sleep per day. 

#Q22
```{r}
#remove days
survey <- survey %>%
  mutate(Q22 = sub("days", "", Q22))
#changing the responses to all numbers
survey <- survey %>% mutate(Q22 = as.numeric(Q22))
ggplot(survey, aes(x=Q14)) +
  geom_bar()
```

#Q23
```{r}
survey <- survey %>%
  #changing the year of birth to their age
  mutate(Q23 = sub("1997", "21", Q23))
#this idiot needs to be deleted
survey$Q23[63] <- NA
survey <- survey %>% mutate(Q23 = as.numeric(Q23))
ggplot(survey, aes(x=Q23)) +
  geom_histogram()
```
The majority of students range from 19-21. 

#Q24
```{r}
#changing the "I don't know" to NA
survey$Q24[38] <- NA
survey <- survey %>% mutate(Q24 = as.numeric(Q24))
ggplot(survey, aes(x=Q24)) +
  geom_histogram()
```
The majority of students range from 160cm to 180cm. 

#Q25
```{r}
survey <- survey%>%
  #we are simply changing all the responses so they are the same
  mutate(Q25 = sub("basketball", "Basketball", Q25))%>%
  mutate(Q25 = sub("swimming", "Swimming", Q25))%>%
  mutate(Q25 = sub("swim", "Swimming", Q25))%>%
  mutate(Q25 = sub("tennis", "Tennis", Q25))%>%
  mutate(Q25 = sub("badminton", "Badminton", Q25))%>%
  mutate(Q25 = sub("afl", "AFL", Q25))%>%
  mutate(Q25 = sub("football", "Football", Q25)) %>%
  mutate(Q25 = sub("soccer", "Soccer", Q25))
#Changing no answer responses to NA
survey$Q25[32] <- NA
survey$Q25[40] <- NA
ggplot(survey, aes(x=Q25)) +
  geom_bar() + coord_flip()
```
The most popular sport is basketball, followed by swimming, then tennis. 

#Q26
```{r}
#changing the responses to their genders
survey$Q26[survey$Q26=="A"] <- "Female"
survey$Q26[survey$Q26=="B"] <- "Male"
survey$Q26[survey$Q26=="C"] <- "Prefer not to answer"

ggplot(survey, aes(x= Q26)) +
  geom_bar()
```
The majority of students in this unit are male. 

#Q27
```{r}
#changing the responses to their names
survey$Q27[survey$Q27=="A"] <- "Fridge"
survey$Q27[survey$Q27=="B"] <- "Pantry"

ggplot(survey, aes(x= Q27)) +
  geom_bar()
```
The majority of students keep their tomato sauce in the fridge.

#Q28
```{r}
#changing the responses to their ranges
survey$Q28[survey$Q28 == "A"] <- "0"
survey$Q28[survey$Q28 == "B"] <- "1-5"
survey$Q28[survey$Q28 == "C"] <- "6-10"
survey$Q28[survey$Q28 == "D"] <- "10-20"
survey$Q28[survey$Q28 == "E"] <- "more than 20"

ggplot(survey, aes(x= Q28)) +
  geom_bar()
```
The majority os students read 1-5 books last year. 

#Q29
```{r}
#changing the responses to yes/no
survey$Q29[survey$Q29 == "A"] <- "Yes"
survey$Q29[survey$Q29 == "B"] <- "No"

ggplot(survey, aes(x= Q29)) +
  geom_bar()
```
The majority os students have not been to the Australian Open.

#Q30
```{r}
#changing the responses to yes/no
survey$Q30[survey$Q30 == "A"] <- "no"
survey$Q30[survey$Q30 == "B"] <- "yes"

ggplot(survey, aes(x= Q30)) +
  geom_bar()
```
The majority of students are struggling with this unit. 

#Q31
```{r}
#changing responses to yes/no
survey$Q31[survey$Q31 == "A"] <- "Yes"
survey$Q31[survey$Q31 == "B"] <- "No"

ggplot(survey, aes(x= Q31)) +
  geom_bar()
```
THe majority of students do enjoy this unit. 

#Q32
```{r}
survey <- survey %>%
  #knife to "Knife"
  mutate(Q32 = sub("knife", "Knife", Q32))%>%
  #spoon related response to "Spoon"
  mutate(Q32 = sub("Soup Spoon", "Spoon", Q32))%>%
  mutate(Q32 = sub("spoon", "Spoon", Q32))%>%
  mutate(Q32 = sub("Soup Spoon", "Spoon", Q32))%>%
  mutate(Q32 = sub("a Spoon", "Spoon", Q32))%>%
  #fork to "Fork"
  mutate(Q32 = sub("fork", "Fork", Q32))%>%
  #ladle related answers to "Ladle"
  mutate(Q32 = sub("ladle", "Ladle", Q32))%>%
  mutate(Q32 = sub("IKEA soup ladle", "Ladle", Q32))%>%
  mutate(Q32 = sub("spork", "Spork", Q32))%>%
  mutate(Q32 = sub("whisk", "Whisk", Q32))%>%
  mutate(Q32 = sub("spatula", "Spatula", Q32))%>%
  mutate(Q32 = sub("Coffee machine", "Coffee Machine", Q32))%>%
  mutate(Q32 = sub("chopstick(s)", "Chopsticks", Q32))


ggplot(survey, aes(x=Q32)) +
  geom_bar() + coord_flip()

```
The top 3 kitchen utensils that students consider themselves to be are knives, forks then spoons. 

#Q33
```{r}
#changing responses to yes/no
survey$Q33[survey$Q33 == "A"] <- "Yes"
survey$Q33[survey$Q33 == "B"] <- "No"

ggplot(survey, aes(x= Q33)) +
  geom_bar()
```
The majority of students like cheese.




#comparing the number of the year level to that of double vs single degrees
```{r}
ggplot(survey) +
  geom_bar(aes(x=Q1)) +
  facet_wrap(~Q12)
```
We can see that in ETC1010 majority of the students are in there second year doing a single degree. However there is also a high number of double degree students in their third year also completeing this unit 

#compare the number of study hours put in according to the level of year each student is in 
```{r}
ggplot(survey)+
  geom_bar(aes(x=Q1)) +
  facet_wrap(~Q8) 
```
We can see overall 3rd year students put a lot more hours into study per week. This could perhaps be due to increased workload during the 3rd year as opposed to 1st year


#Comparing "Are you strggling with this unit" (Q30) and "Do you like ETC1010" (Q31)
```{r}
library(gmodels)
CrossTable(survey$Q30, survey$Q31,prop.chisq=FALSE)

```
Most people (62%) are struggling in the unit, and there is no relationship between struggling in the unit and enjoying the unit. However the vast majority of people who are not struggling are enjoying the unit.


#Comparing "Are you strggling with this unit" (Q30) and "Do you have prior coding experience?" (Q17)
```{r}

CrossTable(survey$Q30, survey$Q17,prop.chisq=FALSE)
```
Just over 55% of the people in the unit have prior coding experience, and two thirds of these people are not struggling. The vast majority of people (96%) of people with no coding experience are struggling in the unit.

```{r}
write_csv(survey, path="data/survey_tidy.csv")