---
title: "Assignment3"
author: "Goup5"
date: "24 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Loading the data:
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
houses <- read_csv("data/Melbourne_housing_FULL.csv")
```

This was provided as part of the assignment properties. 
Warnings have been put to false so that the graphs presented can be seen clearly and the visual representations are clear.




## Collecting certain variables into a new data frame:
```{r message=FALSE, warning=FALSE}
new_houses <- houses %>%
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom)

library(visdat)
vis_miss(new_houses)
colSums(is.na(new_houses))
```
Firstly using the select method, the necessary variables are placed into a new data frame: new_houses. Next vis_miss method and colSums method are applied so shocase the summary of missing values. Values are missing in variables: Price, Distance, Bedroom2 and Bathroom, and number of missing variables is also presented. 




## Missing values summary and strategy for dealing with them:
```{r message=FALSE, warning=FALSE}
rooms_calculation <- new_houses %>%
  mutate(special = Bedroom2 + Bathroom)

rooms_changing <- rooms_calculation %>%
  mutate(
    Rooms = if_else(
      Rooms < !is.na(special), special, Rooms
    ))

new_houses1 <- rooms_changing %>%
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom)

names(new_houses1)[2] <- "TotalRooms"
names(new_houses1)[5] <- "Bedrooms"

# Price, TotalRooms, Type, Distance, Bedrooms, Bathroom

imputed_houses <- new_houses1 %>%
  filter(!is.na(Price)) %>% 
  mutate(
    Distance = if_else(
      is.na(Distance), mean(Distance, na.rm = TRUE), Distance
    ),
    Bedrooms = if_else(
      is.na(Bedrooms), as.integer(median(Bedrooms, na.rm = TRUE)), Bedrooms
    ),
    Bathroom = if_else(
      is.na(Bathroom), as.integer(median(Bathroom, na.rm = TRUE)), Bathroom
    ))

colSums(is.na(imputed_houses))
```

Firstly, a new variable: special had to be created, which is the sum of the bedroom2 and bathroom variable. Then special was compared with the Rooms variable, since some observations had less total rooms than the sum of bedroom2 and bathroom, which is illogical. 
Next all observations where Price was a missing value, was deleted from the data frame. This is because if the response variable is not present, the model would not work properly.
Lastly, missing distance values were replaced with the mean distance (since it can be a double), and the bedrooms and bathroom were replaced with median bedrooms and bathrooms respectively, as they just be whole number - you cannot have half rooms by appling means.
Lastly, again applying colSums, it can be seen that now there are no missing values in the imputed_houses data frame - so models can be developed.




## Plots to summarise relationship between price:
#### The Impact of Type of Housing on Pricing
```{r message=FALSE, warning=FALSE}
ggplot(data = imputed_houses, aes(x = Type, y = Price)) + 
  geom_boxplot(aes(fill = Type)) +
  coord_flip()
```

The trend shows that the type of house does have an effect on price. This is evident through each type of house having a different mean as well as larger quartiles and maximum and minimum values. We can see that type h has a longer tail towards the maximum ending, indicating that there are more houses priced in the higher end. However, both type t and u are similar, both of which have some.

#### The Impact of Total number of Rooms on Pricing
```{r message=FALSE, warning=FALSE}
ggplot(data = imputed_houses) +
  geom_point(mapping = aes(x = TotalRooms, y = Price))
```

There is a relative bell curve shape to this graph. As total rooms increases, so does the price. As expected, there is a majority in the centre where the amount of rooms and the price peaks, then as the total rooms increases, there is a dip. This is expected, as there is less need for more rooms. There are some houses with many rooms and close to the price of the average, however, these are outliers and cannot be expected from all houses.

#### The Impact of Distance from CBD on Pricing
```{r message=FALSE, warning=FALSE}
ggplot(data = imputed_houses) +
  geom_point(mapping = aes(x = Distance, y = Price))
```

There is a clear trend in this graph with the data showing heavy positive skewing.There is a clear majority when there is a smaller distance, with greater house prices. As the distance gets larger, the house prices go down. This makes sense as those who are from overseas or looking to move out of home and buy a new house may be inclined to try and purchase towards the city. Also, there are only small amount of houses that have 0 distance (in the heart of the cbd) this is most likely caused by not being able to build primarily in the city.

#### The Impact of number of Bedrroms on Pricing
```{r message=FALSE, warning=FALSE}
ggplot(data = imputed_houses) +
  geom_point(mapping = aes(x = Bedrooms, y = Price))
```

There is a bell-curve trend in this graph, with the data being normally distributed. As the bedrooms increase, so does the price. Howeverm likewise to the other graph, there are outliers with 15+ and 20 bedrooms. This is probably because 20 rooms is excessive, however the price of this house is cheaper than some of the houses with less bedrooms and hence better value.

#### The Impact of number of Bathrooms on Pricing
```{r message=FALSE, warning=FALSE}
ggplot(data = imputed_houses) +
  geom_point(mapping = aes(x = Bathroom, y = Price))
```

There is no clear pattern to this trend. It is mostly randomly scattered and no definitive skewness. It seems as if most of the variables are randomly mixed and the amount of bathrooms has no clear indication of house price.




## Model and justification for choices of transforamtions, variable selection and interactions:
### Model 1
```{r message=FALSE, warning=FALSE}
imputed_houses <- imputed_houses %>%
  mutate(
    l_distance = log10(Distance + 1),
    l_price = log10(Price)
  )

Model1_fita <- lm(l_price ~ l_distance+I(l_distance^2), data = imputed_houses)
tidy(Model1_fita)
glance(Model1_fita)

dis_mod1 <- augment(Model1_fita, imputed_houses)

ggplot(dis_mod1, aes(x = l_distance, y = l_price)) +
  geom_point() +
  geom_line(aes(y = .fitted), colour = "red")
```

This Model expresses the relationship between distance and price, indicating how a change in the distance of the property from the CBD. As first hypothesised, Price tends to increase by a small proportion as distance increases for a small distance, and then declines as predicted quite as such with distance away from CBD with a potential factor being a reduced demand.
Due to the high levels of variability, and majority of the data being stuck in a small area, a log model was chosen for a better representation. Geom_smooth was also utilised to indicate the quadratic nature of the model I(l_distance^2). As for the accuracy of the model, although it returned a small p value of 4.467498e-313, it had an adjusted r squared of 0.051 and, concluding that this model does not depict a clear view of the data.

### Model 2
```{r message=FALSE, warning=FALSE}
Model1_fitb <- lm(l_price ~ Type, data = imputed_houses)
tidy(Model1_fitb)
glance(Model1_fitb)

dis_mod2 <- augment(Model1_fitb, imputed_houses)

ggplot(dis_mod2, aes(x = Type, y = l_price)) + 
  geom_boxplot() +
  geom_point(aes(y = .fitted), colour = "red", size = 4)
```

This model shows how the type of property can impact the price, varying from house, townhouse and units as broad categories. The model indicates that for the house category, there is a mean price of 602,614 with majority of the houses being between 585,000 - 630,000, albeit many outliers with prices more than 660,000. The Townhouse category expressed a mean of 594,000 and the Unit category had a mean of 575,000. The adjusted r squared for this model worked out to be 0.226, meaning that this model accurately depicted 22.6% of the data in the data set. The P value for this model was quite small and the deviance and df.residual indicate the model was somewhat suitable, and the variables are helping in being able to predict the model.

### Model 3
```{r message=FALSE, warning=FALSE}
Model1_fitc <- lm(l_price ~ l_distance*Type + I(l_distance^2), data = imputed_houses)
tidy(Model1_fitc)
glance(Model1_fitc)

dis_mod3 <- augment(Model1_fitb, imputed_houses)

Final_model <- ggplot(dis_mod3, aes(x = l_distance, y = l_price)) +
  geom_point() +
  geom_line(aes(y = .fitted), colour = "red") + 
  facet_wrap(~ Type)

Final_model
```

The third model uses an interaction between Distance and type in attempt to predict the price for Melbourne housing. It split the housing into 3 categories, Houses, Townhouses and Units, and plots the price against distance from CBD. Carrying the assumption from previous models, that both Distance and Type are key factors in determining the price of the property, we are able to create a better model of the data. This is expressed through the adjusted r squared, being 0.3618, thus indicating that the model can accurately depict 36.18% of the data. Additionally, the p value, deviance and df.residual are relativley smaller which indicates an improved model and that the variables are helping to predict the price of these properties.




## Final model summary and why it is the most suitable:
```{r message=FALSE, warning=FALSE}
tidy(Model1_fitc)
glance(Model1_fitc)
Final_model
```

We decided to choose Model 3 as our final model as it best expressed that data for Melbourne housing. Additionally, this model's coefficients show that they are helping predict that data and make sense of the assumptions held. This model has the highest r squared at 36.18%, while still describing a proportion of the possible relationship. The main reason why we did not choose model1 and model2, what primarily because they are basic models and as a result don't take into account the different variables that could also impact the price. however with model3, the interaction between distance and type of property is taken into account, and shown to be useful in terms of adjusted r squared, p value and deviance statistics.