---
title: "Exercise 7A"
author: "YOUR NAME"
output: html_document
---

```{r}
library(tidyverse)
```


- Fit the models to all countries
- Pick your favourite country (not Australia), print the coefficients, and make a hand sketch of the the model fit.

# Using the gapminder data

# Gapminder

- Hans Rosling was a Swedish doctor, academic and statistician, Professor of International Health at Karolinska Institute. Sadley he passed away in 2017. 
- He developed a keen interest in health and wealth across the globe, and the relationship with other factors like agriculture, education, energy. 
- You can play with the gapminder data using animations at https://www.gapminder.org/tools/. 


<iframe width="1008" height="567" src="https://www.youtube.com/embed/jbkSRLYSojo" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

---
class: bg-main1

# R package: `gapminder`

Contains subset of the data on five year intervals from 1952 to 2007.

```{r show-gapminder}
library(gapminder)
glimpse(gapminder)
```


# "How has life expectancy changed over years, for each country?"

```{r gg-gapminder-line, fig.height = 4, echo = FALSE}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)

```


# "How has life expectancy changed over years, for each country?"

.huge[
- There generally appears to be an increase in life expectancy
- A number of countries have big dips from the 70s through 90s
- a cluster of countries starts off with low life expectancy but ends up close to the highest by the end of the period.
]


# Gapminder: Australia

.huge[
Australia was already had one of the top life expectancies in the 1950s.
]

```{r gapminder-oz}
oz <- gapminder %>% filter(country == "Australia")

oz
```


# Gapminder: Australia

```{r plot-gapminder-oz, fig.height=3, fig.width=8}
ggplot(data = oz, 
       aes(x = year, 
           y = lifeExp)) + 
  geom_line() 
```


# Gapminder: Australia

```{r lm-oz-gapminder}
oz_lm <- lm(lifeExp ~ year, data = oz)

oz_lm
```


```{r tidy-oz-gapminder}
tidy(oz_lm)
```

.huge[
$$\widehat{lifeExp} = -376.1163 - 0.2277~year$$
]


# center year

.huge[
Let us treat 1950 is the first year, so for model fitting we are going to shift year to begin in 1950, makes interpretability easier.
]

```{r center-eyar}
gap <- gapminder %>% mutate(year1950 = year - 1950)
oz <- gap %>%  filter(country == "Australia")
```



# model for centered year

```{r oz-gapminder-centered-year}
oz_lm <- lm(lifeExp ~ year1950, data = oz)

oz_lm
```



```{r tidy-oz-gapminder-centered-year}
tidy(oz_lm)
```

.huge[
$$\widehat{lifeExp} = 67.9 +  0.2277~year$$
]

```{r oz-gapminder-augment, fig.height=3, fig.width=8}
oz_aug <- augment(oz_lm, oz)

oz_aug
```


```{r oz-gap-aug, fig.height=3, fig.width=8}
ggplot(data = oz_aug, 
       aes(x = year, 
           y = .fitted)) + 
  geom_line(colour = "blue") + 
  geom_point(aes(x = year,
                 y = lifeExp))
```


```{r oz-gap-year-resid, fig.height=3, fig.width=8}
ggplot(data = oz_aug, 
             aes(x = year, 
                 y = .std.resid)) + 
  geom_hline(yintercept = 0,
             colour = "white", 
             size = 2) +
  geom_line() 

```


# Making inferences from this

.huge[
- Life expectancy has increased 2.3 years every decade, on average. 
- There was a slow period from 1960 through to 1972, probably related to mortality during the Vietnam war.
]


# Can we fit for New Zealand?

```{r lm-nz}
nz <- gap %>%  filter(country == "New Zealand")
nz_lm <- lm(lifeExp ~ year1950, data = nz)
nz_lm
```


# Can we fit for Japan?

```{r lm-japan}
japan <- gap %>%  filter(country == "Japan")
japan_lm <- lm(lifeExp ~ year1950, data = japan)
japan_lm
```


# Can we fit for italy?

```{r lm-italy}
italy <- gap %>%  filter(country == "Italy")
italy_lm <- lm(lifeExp ~ year1950, data = italy)
italy_lm
```

# Is there a better way?



# Like, what if we wanted to fit a model for ALL countries?


# Let's tinker with the data.


# nest country level data (one row = one country)

```{r nest}
by_country <- gap %>% 
  select(country, year1950, lifeExp, continent) %>%
  group_by(country, continent) %>% 
  nest()

by_country
```


# What is in `data`?

```{r show-nest}
by_country$data[[1]]
```


# It's a list!


# fit a linear model to each one?

```{r fit-many-model}
lm_afganistan <- lm(lifeExp ~ year1950, data = by_country$data[[1]])
lm_albania <- lm(lifeExp ~ year1950, data = by_country$data[[2]])
lm_algeria <- lm(lifeExp ~ year1950, data = by_country$data[[3]])
```


# But we are copying and pasting this code more than twice...is there a better way?


# A case for map???

.huge[
`map(<data object>, <function>)`
]


# A case for map???

```{r mapped-lm}
mapped_lm <- 
map(by_country$data, function(x){
  lm(lifeExp ~ year1950, data = x) 
  })

mapped_lm
```



# Map inside the data?

```{r map-country}
country_model <- by_country %>% 
  mutate(model = map(data, function(x){
    lm(lifeExp ~ year1950, data = x)
    })
    )

country_model
```


# A case for map (shorthand function)

```{r map-country-short}
country_model <- by_country %>% 
  mutate(model = map(data, ~lm(lifeExp ~ year1950, data = .)))

country_model
```


# where's the model?

```{r print-model}
country_model$model[[1]]
```


# we need to summarise this content?

```{r tidy-printed-model}
tidy(country_model$model[[1]])
```


# So should we repeat it for each one?

```{r tidy-many-printed-model}
tidy(country_model$model[[1]])
tidy(country_model$model[[2]])
tidy(country_model$model[[3]])
```


# NO NICK, THERE's A BETTER WAY.

```{r}
country_model %>%
  mutate(tidy = map(model, tidy))
```



```{r map-tidy-model}
country_coefs <- country_model %>%
  mutate(tidy = map(model, tidy)) %>%
  unnest(tidy) %>%
  select(country, continent, term, estimate)

country_coefs
```


```{r tidy-unnested-coefs}
tidy_country_coefs <- country_coefs %>%
  spread(key = term, 
         value = estimate) %>%
  rename(intercept = `(Intercept)`)

tidy_country_coefs
```


```{r oz-unnested-coef, echo=TRUE}
tidy_country_coefs %>%
  filter(country == "Australia")
```


# Your turn: Five minute challenge

- Fit the models to all countries
- Pick your favourite country (not Australia), print the coefficients, and make a hand sketch of the the model fit.


# Plot all the models

```{r augmente-many-countries, fig.height=4}
country_aug <- country_model %>% 
  mutate(augmented = map(model, augment)) %>%
  unnest(augmented)

country_aug
```


```{r plot-gapminder-data, fig.height=4}
p1 <- gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3) + ggtitle("Data")

p2 <- ggplot(country_aug) + 
  geom_line(aes(x = year1950 + 1950, 
                y = .fitted, 
                group = country), 
            alpha = 1/3) +
  xlab("year") +
  ggtitle("Fitted models")
```

```{r plot-print-gapminder, fig.height = 4}
grid.arrange(p1, p2, ncol=2)
```


# Plot all the model coefficients

```{r ggplotly-aug-coef}
p <- ggplot(tidy_country_coefs, 
            aes(x = intercept, 
                y = year1950, 
                colour = continent, 
                label = country)) +
  geom_point(alpha = 0.5, 
             size = 2) +
  scale_color_brewer(palette = "Dark2")
```

---
class: bg-main1

```{r ggplotly-aug-data}
library(plotly)
ggplotly(p)

```


# Let's summarise the information learned from the model coefficients.

- Generally the relationship is negative: this means that if a country started with a high intercept tends to have lower rate of increase.
- There is a difference across the continents: Countries in Europe and Oceania tended to start with higher life expectancy and increased; countries in Asia and America tended to start lower but have high rates of improvement; Africa tends to start lower and have a huge range in rate of change.
- Three countries had negative growth in life expectancy: Rwand, Zimbabwe, Zambia


# Model diagnostics by country

```{r glance-country}
country_glance <- country_model %>% 
  mutate(glance = map(model, glance)) %>%
  unnest(glance)

country_glance
```


# Plot the $R^2$ values as a histogram. 

```{r country-fit, fig.height = 3.5}
ggplot(country_glance, 
       aes(x = r.squared)) + 
  geom_histogram()
```


# Countries with worst fit

Examine the countries with the worst fit, countries with $R^2<0.45$, by making scatterplots of the data, with the linear model overlaid. 

```{r show-bad-fit}
badfit <- country_glance %>% filter(r.squared <= 0.45)

gap_bad <- gap %>% filter(country %in% badfit$country)

gg_bad_fit <-
ggplot(data = gap_bad, 
       aes(x = year, 
           y = lifeExp)) + 
         geom_point() +
  facet_wrap(~country) +
  scale_x_continuous(breaks = seq(1950,2000,10), 
                     labels = c("1950", "60","70", "80","90","2000")) +
  geom_smooth(method = "lm", 
              se = FALSE)
```


```{r gg-show-bad-fit, fig.height=4}
gg_bad_fit
```

Each of these countries had been moving on a nice trajectory of increasing life expectancy, and then suffered a big dip during the time period.




## Your turn: new data - fertility data

-Using the fertility data from the [gapminder web site](https://www.gapminder.org/data/) (in data/ folder)
- Conduct the same analysis as previous, but for life expectancy.
- Find unusual countries

The code below will help you read in the data and process it, but you will need to make some changes to do the full analysis.

```{r read-data}
library(readxl)
library(janitor)
fertility_raw <- read_xlsx("data-raw/indicator-undata-total-fertility.xlsx") %>%
  clean_names() %>%
  rename(country = total_fertility_rate)
```

You will need to gather the data into longer form, use the `gather` function. 

You should know how to do the rest from your experience with past assignments.

```{r tidy-fertility}
fertility <- fertility_raw %>% 
  gather(key = year, 
                value = fertility, 
                -country) %>%
  mutate(year = parse_number(year)) %>%
  filter(year > 1950) %>%
  mutate(year1950 = year - 1950)
```


plot the fertility data for each year

```{r eval=FALSE}
ggplot(fertility, 
       aes(x = year, 
           y = fertility, 
           group = country)) +
  geom_line(alpha=0.1)
```

# How has fertility changed over years? For each country?

## How has fertility changed in Australia?

## Center the year variable, using 1950 as the centerpoint

## Fit a model for Australia

## Explore model fit for Australia

## How has fertility increased every year / decade in Australia?

## How well does the model fit?

## Fit a linear model for all countries

## Plot the model fit against the data in two plots

## plot the model coefficients with plotly

## Summarise what you have learned from the model coefficients

## Find those countries with the worst fit

