---
title: "Mapping flight movements"
author: "YOUR NAME"
output: html_document
---

```{r setup}
library(tidyverse)
```

# Aim: Show flight movement on the map

# First: load in the plane and airline data

```{r load-plane}
plane_N4YRAA <- read_csv("data/plane_N4YRAA.csv")
glimpse(plane_N4YRAA)
```

# Load Airline travel, airport location

```{r read-airports}
airport_raw <- read_csv("data/airports.csv")
airport_raw %>%
  select(AIRPORT, 
         LATITUDE, 
         LONGITUDE, 
         AIRPORT_STATE_NAME) %>%
  glimpse()

```

The Key is the airport three letter code, 
    - called ORIGIN or DEST in `plane_N4YRAA` table
    - called AIRPORT in the `airport` table

One table, plane_N4YRAA, has less airports than the other
    - We only want to keep the rows of airport table, for those that appear in the plane_N4YRAA table

```{r tidy-airport}
airport <- airport_raw %>%
  select(AIRPORT, 
         LATITUDE, 
         LONGITUDE, 
         AIRPORT_IS_LATEST, 
         DISPLAY_AIRPORT_NAME) %>%
  filter(AIRPORT_IS_LATEST == 1) %>%
  select(-AIRPORT_IS_LATEST)

airport
```


```{r tidy-flight}
N4YRAA_latlon <- left_join(plane_N4YRAA,
                           airport,
                           by = c("ORIGIN" = "AIRPORT")
) %>%
  rename(
    "ORIGIN_LATITUDE" = "LATITUDE",
    "ORIGIN_LONGITUDE" = "LONGITUDE"
  )

N4YRAA_latlon %>%
  select(
    ORIGIN,
    ORIGIN_LATITUDE,
    ORIGIN_LONGITUDE,
    DISPLAY_AIRPORT_NAME
  )
```

The variables ORIGIN_LATITUDE, ORIGIN_LONGITUDE, DISPLAY_AIRPORT_NAME are added to corresponding row in the plane_N4YRAA table.


# Add destination locations

- Added spatial coordinates (lat, lon) for the origin airport
- Same needs to be done for the destination airport
- Then the airports can be drawn over a map

Now we combine the current data with some new data to add the destinations.

```{r}
N4YRAA_latlon <- left_join(
  N4YRAA_latlon,
  airport,
  by = c("DEST" = "AIRPORT")
) %>%
  rename(
    "DEST_LATITUDE" = "LATITUDE",
    "DEST_LONGITUDE" = "LONGITUDE"
  )

N4YRAA_latlon <- N4YRAA_latlon %>% arrange(FL_DATE, DEP_TIME)
```

# Map it

NOTE: THIS CODE DOES NOT NEED TO RUN. It is for illustrative purposes only, to show you how to get google maps info.

```{r eval=FALSE}
# Need to use a registered google account to get map
register_google("XXX")
map <- get_map(c(lon = -92.20562, lat = 36.20259), zoom = 5)
write_rds(map, "data/map.rds")
```


```{r}
library(ggmap)
library(lubridate)
library(ggthemes)

map <- read_rds("data/map.rds")
# load("data/map.rda")

N4YRAA_latlon_2017_05_06 <- N4YRAA_latlon %>%
  filter(FL_DATE == ymd("2017-05-06"))

ggmap(map) +
  geom_segment(data = N4YRAA_latlon_2017_05_06,
    aes(x = ORIGIN_LONGITUDE,
        xend = DEST_LONGITUDE,
        y = ORIGIN_LATITUDE,
        yend = DEST_LATITUDE),
    color = "navyblue") +
  geom_point(data = N4YRAA_latlon_2017_05_06,
    aes(x = ORIGIN_LONGITUDE,
        y = ORIGIN_LATITUDE),
    color = "orange",
    alpha = 0.3,
    size = 3) +
  geom_point(data = N4YRAA_latlon_2017_05_06,
    aes(x = DEST_LONGITUDE,
        y = DEST_LATITUDE),
    color = "red",
    alpha = 0.3,
    size = 1) +
  theme_map()
```


```{r}
ggmap(map) +
  geom_segment(data = N4YRAA_latlon,
               aes(x = ORIGIN_LONGITUDE,
                   xend = DEST_LONGITUDE,
                   y = ORIGIN_LATITUDE,
                   yend = DEST_LATITUDE),
               color = "navyblue") +
  geom_point(data = N4YRAA_latlon_2017_05_06,
             aes(x = ORIGIN_LONGITUDE,
                 y = ORIGIN_LATITUDE),
             color = "orange",
             alpha = 0.3,
             size = 3) +
  geom_point(data = N4YRAA_latlon_2017_05_06,
             aes(x = DEST_LONGITUDE,
                 y = DEST_LATITUDE),
             color = "red",
             alpha = 0.3,
             size = 1) +
  theme_map() + 
  facet_wrap(~FL_DATE)
```

# Your Turn: Exercise

- Go to rstudio.cloud
- Open `nycflights.Rmd` and complete exercises
