---
title: "Lab exercise: Exploring Women's AFL"
author: "YOUR NAME"
output: html_document
---

```{r setup}
library(tidyverse)
```

# Read in the data

We are looking at data for Women's AFL.

There are two tables:

- players.csv
- teams.csv 

The data was sourced from this website: https://github.com/ropenscilabs/ozwomensport/tree/master/AFLW/data

Read them into R:

```{r read-data}
# to clean up the names of the data
library(janitor)
players <- read_csv("data/players.csv") %>% clean_names() 
teams <- read_csv("data/teams.csv") %>% clean_names()
```

# Have a look at the data with `View()` or `glimpse()`.

# Using the teams dataset, make a barchart of the total kicks by team.  What do you learn?

```{r barchart-total-kicks}
ggplot(teams,
       aes(x = club,
           y = kicks_tot)) + 
  geom_col()
```

# Take that same barplot and reorder the club by `kicks_tot`, what changes?

```{r barchart-total-kicks-reorder}
ggplot(teams,
       aes(x = reorder(club, kicks_tot),
           y = kicks_tot)) + 
  geom_col()
```


## Now, make the plot again, but sort from highest to lowest, and flip the plot with `coord_flip()`. How did this help make it easier to see what was going on?


```{r barchart-total-kicks-flip}
ggplot(teams,
       aes(x = reorder(club, kicks_tot),
           y = kicks_tot)) + 
  geom_col() + 
  coord_flip()
```

# To compare teams, should you use average or total numbers? Explain your thinking.

> ...

# Make a scatterplot of average kicks by average handballs. Use an aspect ratio of 1. What do you learn?

```{r gg-kicks-handballs}
ggplot(teams,
       aes(x = kicks_avg,
           y = handballs_avg)) + 
  geom_point() + 
  theme(aspect.ratio = 1)
```

##  Use the `plotly` package to make the plot interactive (with the `ggplotly` function). Report the name of the team(s) that has(ve) the highest on both variables.

```{r plotly}
library(plotly)
ggplot(teams,
       aes(x = kicks_avg,
           y = handballs_avg,
           # this label option gets passed to ggplotly
           label = club)) + 
  geom_point() + 
  theme(aspect.ratio = 1)

# ggplotly uses some programming magic and uses the last ggplot object you made
# to make it interactive.
ggplotly()
```


# Make a scatterplot of average frees for (frees_for_avg) by against (frees_agst_avg), for each season. Use an aspect ratio of 1. What do you learn? 

```{r gg-avg-frees}
ggplot(teams,
       aes(x = frees_for_avg,
           y = frees_agst_avg)) + 
  geom_point() +
  theme(aspect.ratio = 1)
```


## Make the plot interactive and report the name of the team that has the most frees against and fewest for.

```{r gg-team-frees-for-against}
# use the similar ggplotly code from before
ggplot(teams,
       aes(x = frees_for_avg,
           y = frees_agst_avg,
           text = club)) + 
  geom_point() + 
  theme(aspect.ratio = 1)

ggplotly()
```


# Using the players data, make a side-by-side boxplot of the average kicks by team. What do you learn? 

```{r gg-players-club-avg-kicks}
ggplot(players,
       aes(x = club,
           y = kicks_avg)) + 
  geom_boxplot()
```

# Which team had the player with the highest average kicks? Do some teams appear to have "star" players, and others have more even skills in this area?

# Using the players data, make a scatterplot of average kicks against handballs, for each season. Overlay a linear model. 

```{r gg-players}
ggplot(players,
       aes(x = kicks_avg,
           y = handballs_avg)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

## Add interaction with plotly. Which players are more likely to kick vs hand ball? Also handball rather than kick?

```{r gg-players-kicks-handballs}
ggplot(players,
       aes(x = kicks_avg,
           y = handballs_avg,
           label = player)) + 
  geom_point() +
  geom_smooth(method = "lm")

ggplotly()
```

