---
title: "ETC 1010 Assignment 3"
author: "Group 2"
output: html_document
---


```{r, eval=FALSE}
## try http:// if https:// URLs are not supported
source("https://bioconductor.org/biocLite.R")
biocLite("impute")
```

```{r}
library(dplyr)
library(VIM)
library(impute)
library(broom)
library(mice)
library(modelr)
library(tidyverse)
library(naniar)
library(visdat)
```

## Step 1
Read the data and select the required information

```{r}
houses <- read_csv("Melbourne_housing_FULL.csv") %>%
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom)
```

## Step 2
Summary the situation of missing value

```{r}
vis_dat(houses, palette = "cb_safe")
vis_miss(houses, sort_miss = TRUE)+theme(aspect.ratio=1)
```
According to the chart we foud out that the missing value account for 11.5% of the total data. Onlt three variables (Bathroom, Bedroom2, Price) have missing values.
```{r}
miss_summary(houses)
```
Although the missings are small in number (11.5%), there are a huge proportion of missing variables (66.7%). Hence we cannot simply drop the missing values. Imputing the missings is necessary.

## Step 3
Dealing with missing values and data mistakes

```{r}
rooms_calsulation <-  houses %>%
  mutate(Trooms = Bedroom2 + Bathroom)

rooms_changing <- rooms_calsulation %>%
  mutate(Rooms = if_else(Rooms < !is.na(Trooms),Trooms, Rooms))

new_houses1 <- rooms_changing %>%
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom)
```
We found that the number of total rooms is not equal to the sum of bathrooms and bedrooms in some cases. So we fixed this problem by replacing the existing data with the sum of number of bathrooms and bedrooms.


```{r}
names(new_houses1)[2] <- "TotalRooms"
names(new_houses1)[5] <- "Bedrooms"

imputed_houses <- new_houses1 %>%
  filter(!is.na(Price)) %>%
  mutate(
    Distance = if_else(
      is.na(Distance), mean(Distance, na.rm = TRUE), Distance),
    Bedrooms = if_else(
      is.na(Bedrooms), as.integer(median(Bedrooms, na.rm = TRUE)), Bedrooms    ),
    Bathroom = if_else(
      is.na(Bathroom), as.integer(median(Bathroom, na.rm = TRUE)), Bathroom    ),
    TotalRooms = if_else(is.na(TotalRooms),Bathroom+Bedrooms,TotalRooms),
    Type = sub("u", "0", Type),
    Type = sub("t", "1", Type),
    Type = sub("h", "2", Type)  )
  colSums(is.na(imputed_houses))
```
As we will create a regression for predicting Price useing this dataset, so we drop the missings in Price and impute the missing values in Distance, Bedrooms, Bathroom and TotalRooms. Additionally, Type is changed from a catorgorical vairbale to a numberical variable for regression line prediction. In the final code, we use "colSums" function to make sure there are no missing values in the dataset.

## Step 4
Anaylsing the data by plots

```{r}
imputed_houses$Price <- as.numeric(imputed_houses$Price)
imputed_houses$TotalRooms <- as.numeric(imputed_houses$TotalRooms)
imputed_houses$Type <- as.integer(imputed_houses$Type)
```

```{r}

ggplot(data = imputed_houses,aes(x=TotalRooms,y=Price))+
  geom_smooth()
ggplot(data = imputed_houses,aes(x=TotalRooms,y=log(Price)))+
  geom_smooth()
```
This plot shows that the number of total rooms strongly effect the price. We also tested the correlation between total rooms and logarithmic price in order to create our regression. 
```{r}
ggplot(imputed_houses)+
  geom_smooth(aes(x = Distance,y = log(Price)), se = FALSE,color = "orange")+
  facet_wrap(~Type)
ggplot(imputed_houses)+
  geom_smooth(aes(x = Distance,y = Price), se = FALSE, color = "orange")+
  facet_wrap(~Type)
```
This graph indicates that the distance has a negative impact on the price. We also tested the correlation between distance and logarithmic price in order to create our final regression. 

```{r}
ggplot(imputed_houses,aes(x=Bedrooms,y=Price))+
  geom_smooth(color="red")
ggplot(imputed_houses,aes(x=Bedrooms,y=log(Price)))+
  geom_smooth(color="red")
```
According to this graph, we can that the number of bedrooms can impact the final price of houses.  We also tested the correlation between the number of bedrooms and logarithmic price in order to create our final regression. 

```{r}
ggplot(imputed_houses,aes(x=Bathroom,y=Price))+
  geom_smooth(color="purple")
ggplot(imputed_houses,aes(x=Bathroom,y=log(Price)))+
  geom_smooth(color="purple")
```
According to this graph, we can that the number of bathrooms can effect the final price of houses. Additionally, We detected the correlation between the number of bedrooms and logarithmic price in order to create our final regression. 
```{r}
tbg1 <- filter(imputed_houses, Bathroom <= 2 & TotalRooms <= 8)
ggplot(tbg1, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg2 <- filter(imputed_houses, Bathroom == 3 & TotalRooms <= 8)
ggplot(tbg2, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg3 <- filter(imputed_houses, Bathroom == 4 & TotalRooms <= 8)
ggplot(tbg3, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg4 <- filter(imputed_houses, Bathroom == 5 & TotalRooms <= 8)
ggplot(tbg4, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg5 <- filter(imputed_houses, Bathroom == 6 & TotalRooms <= 8)
ggplot(tbg5, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg6 <- filter(imputed_houses, Bathroom == 7 & TotalRooms <= 8)
ggplot(tbg6, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg7 <- filter(imputed_houses, Bathroom == 8 & TotalRooms <= 8)
ggplot(tbg7, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
ggplot(imputed_houses)+
  geom_hex(aes(x = Distance,y = log(Price)))+
  facet_wrap(~Type)
ggplot(imputed_houses)+
  geom_hex(aes(x = Distance,y = Price))+
  facet_wrap(~Type)


```
```{r}
imputed_houses <- mutate(imputed_houses, logprice = log(Price))
tbg1 <- filter(imputed_houses, Bathroom <= 2 & TotalRooms <= 8)
ggplot(tbg1, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg2 <- filter(imputed_houses, Bathroom == 3 & TotalRooms <= 8)
ggplot(tbg2, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg3 <- filter(imputed_houses, Bathroom == 4 & TotalRooms <= 8)
ggplot(tbg3, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg4 <- filter(imputed_houses, Bathroom == 5 & TotalRooms <= 8)
ggplot(tbg4, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg5 <- filter(imputed_houses, Bathroom == 6 & TotalRooms <= 8)
ggplot(tbg5, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg6 <- filter(imputed_houses, Bathroom == 7 & TotalRooms <= 8)
ggplot(tbg6, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg7 <- filter(imputed_houses, Bathroom == 8 & TotalRooms <= 8)
ggplot(tbg7, mapping = aes(x = Distance, y = logprice, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
ggplot(imputed_houses)+
  geom_hex(aes(x = Distance,y = log(Price)))+
  facet_wrap(~Type)
ggplot(imputed_houses)+
  geom_hex(aes(x = Distance,y = Price))+
  facet_wrap(~Type)
imputed_houses <- imputed_houses %>%
  mutate(Type = sub("u", "0", Type),
         Type = sub("t", "1", Type),
         Type = sub("h", "2", Type))

```



## Step 5 
Establish the regression

```{r}
sim1 <- lm(Price ~ TotalRooms, data=imputed_houses)
sim2 <- lm(Price ~ Distance, data=imputed_houses)
sim3 <- lm(Price ~ Bedrooms, data=imputed_houses)
sim4 <- lm(Price ~ Bathroom, data=imputed_houses)
sim5 <- lm(Price ~ Type, data=imputed_houses)
glance(sim1)
glance(sim2)
glance(sim3)
glance(sim4)
glance(sim5)
```
The first model type we tried is single factor model, we will see the relation ship between price and different factors, seems like in single factor model, Distance model(sim2) has the lowest R^2, which mean distance is the worst variables to fit the pridicting price.

```{r}
sim11 <- lm(Price ~ TotalRooms+Distance, data=imputed_houses)
sim12 <- lm(Price ~ TotalRooms*Distance, data=imputed_houses)
sim13 <- lm(Price ~ TotalRooms+Bedrooms, data=imputed_houses)
sim14 <- lm(Price ~ TotalRooms*Bedrooms, data=imputed_houses)
sim15 <- lm(Price ~ TotalRooms+Bathroom, data=imputed_houses)
sim16 <- lm(Price ~ TotalRooms*Bathroom, data=imputed_houses)
sim17 <- lm(Price ~ TotalRooms+Type, data=imputed_houses)
sim18 <- lm(Price ~ TotalRooms*Type, data=imputed_houses)

glance(sim11)
glance(sim12)
glance(sim13)
glance(sim14)
glance(sim15)
glance(sim16)
glance(sim17)
glance(sim18)
```
By adding another factor, the adjusted r^2 of model is significantly improved, and for total rooms and distance, the model with interaction has higher adjusted r^2, therefore total rooms will be times by distance in our final model
```{r}
sim19 <- lm(Price ~ Distance+Bedrooms, data=imputed_houses)
sim20 <- lm(Price ~ Distance*Bedrooms, data=imputed_houses)
sim21 <- lm(Price ~ Distance+Bathroom, data=imputed_houses)
sim22 <- lm(Price ~ Distance*Bathroom, data=imputed_houses)
sim23 <- lm(Price ~ Distance+Type, data=imputed_houses)
sim24 <- lm(Price ~ Distance*Type, data=imputed_houses)

glance(sim19)
glance(sim20)
glance(sim21)
glance(sim22)
glance(sim23)
glance(sim24)
```
In theses 6 models, models with interaction shown a higher adjusted r^2 compared with models without interaction. Therefore, in final model, distance will have interaction with other factors
```{r}
sim25 <- lm(Price ~ Bedrooms+Bathroom, data=imputed_houses)
sim26 <- lm(Price ~ Bedrooms*Bathroom, data=imputed_houses)
sim27 <- lm(Price ~ Bedrooms+Type, data=imputed_houses)
sim28 <- lm(Price ~ Bedrooms*Type, data=imputed_houses)
sim29 <- lm(Price ~ Bathroom+Type, data=imputed_houses)
sim30 <- lm(Price ~ Bathroom*Type, data=imputed_houses)

glance(sim25)
glance(sim26)
glance(sim27)
glance(sim28)
glance(sim29)
glance(sim30)
```
In these 6 models, there is not significant difference between model with interaction and model without interaction, therefore there will not be interaction between these factors in the final model
```{r}
sim31 <- lm(Price ~ Distance*TotalRooms+Distance*Bedrooms+Distance*Bathroom+Distance*Type, data=imputed_houses)
sim32 <- lm(Price ~ Distance*TotalRooms*Bedrooms*Bathroom*Type, data=imputed_houses)
sim33 <- lm(Price ~ Distance*TotalRooms*(Bedrooms+Bathroom)*Type, data=imputed_houses)
sim34 <- lm(log(Price) ~ Distance*TotalRooms*Bedrooms*Bathroom*Type, data=imputed_houses)

glance(sim31)
glance(sim32)
glance(sim33)
glance(sim34)
```
Based on the results above, the sim34 has the largest adjusted R-squared which is 0.539. Sim34 is the best regression of price.

## Step 6
Summary the final model
```{r}
summary(sim34)
```

In "step 4" we have proved that all 5 variable have significant effect on price, so we do not need to drop any variables in our final model. Furthermore, in "step 5" we tried lots of formula to find out the best equation of the regression. Finally, the sim34 comes out to be the best model which has the largest adjusted R-squared.

## Step 7
Predicting the Price

```{r}
imputed_houses <- add_predictions(imputed_houses, sim34)
tbg21 <- filter(imputed_houses, Bathroom <= 2 & TotalRooms <= 8)
ggplot(tbg21, mapping = aes(x = Distance, color = TotalRooms))+
  geom_line(aes(y = pred), colour = "red")+
  geom_point(aes(y = logprice))+
  facet_grid(Bathroom ~ Type)

imputed_houses <- add_residuals(imputed_houses, sim34)

ggplot(imputed_houses, aes(resid))+
  geom_freqpoly(binwidth = 0.5)

ggplot(imputed_houses, aes(logprice, resid))+
  geom_ref_line(h = 0)+
  geom_point()

```
These plots are from data that filled in the pridicted price.

```{r}
rooms_calsulation <-  houses %>%
  mutate(special = Bedroom2 + Bathroom)

rooms_changing <- rooms_calsulation %>%
  mutate(TotalRooms = if_else(Rooms < !is.na(special),special, Rooms))

houses3 <-  rooms_changing %>%
  select(Price, TotalRooms, Type, Distance, Bedroom2, Bathroom)


houses4 <- houses3 %>%
  mutate(
    Distance = if_else(
      is.na(Distance), mean(Distance, na.rm = TRUE), Distance),
    Bedrooms = if_else(
      is.na(Bedroom2), as.integer(median(Bedroom2, na.rm = TRUE)), Bedroom2    ),
    Bathroom = if_else(
      is.na(Bathroom), as.integer(median(Bathroom, na.rm = TRUE)), Bathroom
    ))
  
  md.pattern(houses4)
```
We run the code again to make sure there is no missings in the prdicted results.

```{r}
houses4 <- houses4 %>%
  mutate(Type = sub("0", "unit_houses", Type),
         Type = sub("1", "town_houses", Type),
         Type = sub("2", "houses", Type))
tbg21 <- filter(houses4, Bathroom <= 2 & TotalRooms <= 8)
ggplot(tbg21, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg22 <- filter(houses4, Bathroom == 3 & TotalRooms <= 8)
ggplot(tbg22, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg23 <- filter(houses4, Bathroom == 4 & TotalRooms <= 8)
ggplot(tbg23, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg24 <- filter(houses4, Bathroom == 5 & TotalRooms <= 8)
ggplot(tbg24, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg25 <- filter(houses4, Bathroom == 6 & TotalRooms <= 8)
ggplot(tbg25, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg26 <- filter(houses4, Bathroom == 7 & TotalRooms <= 8)
ggplot(tbg26, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg27 <- filter(houses4, Bathroom == 8 & TotalRooms <= 8)
ggplot(tbg27, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg28 <- filter(houses4, Bathroom >= 8 & TotalRooms <= 8)
ggplot(tbg28, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
tbg29 <- filter(houses4, TotalRooms >= 8)
ggplot(tbg29, mapping = aes(x = Distance, y = Price, color = TotalRooms))+
  geom_point()+
  geom_smooth(se = FALSE)+
  facet_grid(Bathroom ~ Type)
ggplot(houses4)+
  geom_hex(aes(x = Distance,y = log(Price)))+
  facet_wrap(~Type)
ggplot(houses4)+
  geom_hex(aes(x = Distance,y = Price))+
  facet_wrap(~Type)

```
We make plots for the pridicted data.