---
title: 'Summarising missing values'
---

```{r setup}
library(naniar)
library(dplyr)
```

# Summarizing missingness

Now that you understand the behavior of missing values in R, and how to count them, let's scale up our summaries for cases (rows) and variables, using `miss_var_summary()` and `miss_case_summary()`, and also explore how they can be applied for groups in a dataframe, using the `group_by` function from `dplyr`.

Calculate summaries of missingness in the `airquality` dataset for variables using the `miss_var_summary()` function.

## Hint

* Pass `airquality` into the `miss_var_summary()` function.

```{r}
# Summarise missingness in each variable of the `airquality` dataset
miss_var_summary(___)
```

Calculate summaries of missingness in the `airquality` dataset for the cases using the `miss_case_summary()` function.

## Hint

* Pass `airquality` into the `miss_case_summary()` function.

```{r}
# Summarise missingness in each case of the `airquality` dataset
___(___)
```

Using the `airquality` dataset, use `group_by()` to create summaries for each variable and case, by each `Month`.

## Hint
* Pass `Month` into the `group_by()` function.
* Then use the `miss_var_summary()` and `miss_case_summary()` function.

```{r}
# Return the summary of missingness in each variable, grouped by Month, in the `airquality` dataset
airquality %>% group_by(___) %>% ___

# Return the summary of missingness in each case, grouped by Month, in the `airquality` dataset
airquality %>% group_by(___) %>% ___
```

# Tabulating Missingness

The summaries of missingness we just calculated give us the number and percentage of missing observations for the cases and variables. 

Another way to summarise missingness is by tabulating the number of times that there are 0, 1, 2, 3, missings in a variable, or in a case. 

In this exercise we are going to tabulate the number of missings in each case and variable using `miss_var_table()` and `miss_case_table()`, and also combine these summaries with the the `group_by` operator from `dplyr`. to explore the summaries over a grouping variable in the dataset.

## Instructions
For the `airquality` dataset:

- Tabulate missingness for each variable using `miss_var_table()`.
- Tabulate missingness for every case using `miss_case_table()`.
- Combine previous tabulations with `dplyr`'s `group_by()` function to create tabulations for each variable and case, by each `Month`.

### Hint

Don't forget to use `miss_case_table(airquality)`.

Also, don't forget to group by `Month`, using `group_by(Month)` like this: `airquality %>% group_by(Month) %>% miss_case_table()`.

```{r}
# Tabulate missingness in each variable and case of the `airquality` dataset
___(airquality)
___(___)

# Tabulate the missingness in each variable, grouped by Month, in the `airquality` dataset
airquality %>% group_by(___) %>% miss_var_table()

# Tabulate of missingness in each case, grouped by Month, in the `airquality` dataset
airquality %>% ___ %>% miss_case_table()
```
