---
title: "Exploring Housing prices in Melbourne"
author: "YOUR NAME"
---

```{r setup}
library(tidyverse)
library(naniar)
library(simputation)
```

# Lab exercise

We are going to examine Melbourne house prices, using a data set available at https://www.kaggle.com/anthonypino/melbourne-housing-market). It was constructed by scraping the auction reports over several years, and adding more information on property details by scraping the web pages on the properties at domain.com.


```{r prepare-data}
houses_raw <- read_csv("data/Melbourne_housing_FULL.csv") %>%
  select(Price, 
         Rooms, 
         Type, 
         Distance, 
         Bedroom2, 
         Bathroom)

```

1. Make an overview plot (e.g., vis_dat, vis_miss) of the full data. Which variables have missings? 

```{r use-vis-miss}
vis_miss(houses_raw)
```

# 2. Make a missing values summary of all the variables in the data. What proportion of observations are missing on Price?

```{r use-miss-var-summary}
miss_var_summary(houses_raw)
```


# 3. Remove the observations that have missing values on Price, because this is the response variable that we want to ultimately predict. We can't build a stable model for price if we don't know the price. 

```{r filter-na-price}
houses <- houses_raw %>% filter(!is.na(Price))
```


# 4. Make the tidy missing data for the houses data

```{r houses-shadow}
houses_shadow <- bind_shadow(houses)
```

## 4A Plot Bathroom vs Bedroom2 coloured by missingness on Bedroom2. Why don't any missing values show up?

```{r gg-houses-shadow}
ggplot(houses_shadow,
       aes(x = Bathroom,
           y = Bedroom2,
           colour = Bedroom2_NA)) + 
  geom_point()
```

# 5. Impute the missing values for Bedroom2 and Bathroom, by using mean imputation. 

```{r houses-imp}
houses_imp_mean <- houses_shadow %>%
  mutate(Bedroom2 = impute_mean(Bedroom2),
         Bathroom = impute_mean(Bathroom)) %>%
  # identify where there were missings
  add_label_shadow()
```

## 5A Make a plot of the two variables, with the imputed values coloured. Describe the pattern that you see.

```{r gg-show-missings}
ggplot(houses_imp_mean,
       aes(x = Bedroom2,
           y = Bathroom,
           colour = any_missing)) + 
  geom_point()
```


# 6. Impute the missing values for Bedroom2 and Bathroom, using a linear model on the variable Rooms. 

```{r}
library(simputation)
houses_imp_lm <- houses_shadow %>%
  impute_lm(formula = Bedroom2 ~ Rooms) %>%
  impute_lm(formula = Bathroom ~ Rooms) %>%
  add_label_shadow()
```


## 6A Make a plot of the two variables, with the imputed values coloured. Is this better or worse than the mean value imputed values? Explain your thinking. (The code below can help you do this type of imputation, but **you will need to make the start of the code yourself, to read in the data, and do the initial exercises**. )

```{r echo=TRUE, eval=FALSE}
ggplot(houses_imp_lm, 
       aes(x = Bedroom2, 
           y = Bathroom,
           colour = any_missing)) +
  geom_point(alpha=0.2) + 
  scale_colour_brewer("Miss bedroom", 
                      palette="Dark2")
```


