---
title: 'Introduction to missing values'
---

```{r setup}
library(naniar)
```


# Using and finding missing values

When working with missing data, there are a couple of commands that you should be familiar with - firstly, you should be able to identify if there are **any** missing values, and where these **are**.

Using the `any_na()` and `are_na()` tools, identify which values are missing.

- Create a vector, `x`, which has the values `NA`, `NaN`, `Inf`, `"."`, and `"missing"`.
- Pass a vector `x` to `any_na()` and `are_na()` to find and explore which ones are missing.

### Hints

* Using `any_na(x)` will return `TRUE` if there are missing values, and `FALSE` if there are none.
* Using `are_na(x)` will return `TRUE` or `FALSE` for each value in heights - if a value is missing, it will be `TRUE`, else, `FALSE`.

```{r}
# Create x, a vector, with values NA, NaN, Inf, ".", and "missing"
x <- c(___, ___, ___, ___, ___)

# Use any_na() and are_na() on to explore the missings
any_na(___)
are_na(___)
```

## How many missing values are there?

One of the first things that you will want to check with a new dataset is if there are any missing missing values, and how many there are.

You could use `are_na()` to and count up the missing values, but **the most efficient way** to count missings is to use the `n_miss()` function. This will tell you the **total number of missing values** in the data.

You can then find the percent of missing values in the data with the `pct_miss` function. This will tell you the **percentage of missing values** in the data.

You can also find the complement to these - how many complete values there are - using `n_complete` and `pct_complete`.

## instructions

Using the example dataframe of heights and weights `dat_hw`:

- Use `n_miss()` on the dataframe `dat_hw` to count the total number of missing values the dataframe.
- Use `n_miss()` on the variable `dat_hw$weight` to count the total number of missing values it.
- Similarly, use `prop_miss()`, `n_complete()`, and `prop_complete()` to get the proportion of missings, and the number and proportion of complete values for the dataframe and the variables.

### Hints:

* Use `n_miss()` on data like so: `n_miss(dat_hw)` or `n_miss(dat_hw$weight)`.

```{r setup-new-data}
# need to make height and weight related to each other.
dat_hw <- data.frame(weight = rnorm(100, mean = 85, sd = 15),
                     height = rnorm(100, mean = 1.7, sd = 1))

# add missingness
dat_hw[sample(x = 1:100, size = 15), 1] <- NA
dat_hw[sample(x = 1:100, size = 15), 2] <- NA
```

#### Exercise

```{r exercise-n-miss}
# Use n_miss() to count the total number of missing values in dat_hw
n_miss(___)

# Use n_miss() on dat_hw$weight to count the total number of missing values
n_miss(___$___)

# Use n_complete() on dat_hw to count the total number of complete values
n_complete(___)

# Use n_complete() on dat_hw$weight to count the total number of complete values
___(___$___)

# Use prop_miss() and prop_complete() on dat_hw to count the total number of missing values in each of the variables
prop_miss(____)
prop_complete(___)
```

# Working with missing values

R stores missing values as `NA`, which have some special behavior. Now that you can define missing data and understand how R stores missing values, can  you **predict what will happen when we operate with some missing values**?

What is the output of the following four commands in R? Try them out in the code console to test them before you submit your answer.

```r
1 + NA
NA + NA
NA | TRUE
NA | FALSE
```


Remember that for the most part any calculation using `NA` returns `NA` - this is contrasted with a few cases:

* `NA | TRUE` returns TRUE - this is one of the rare exceptions when working with NA values in R, it is a good one to be aware of.
* `NA | FALSE` returns FALSE - this is one of the rare exceptions when working with NA values in R, it is a good one to be aware of.

