---
title: "Visualising missing values"
---

```{r setup}
library(naniar)
```

## Your first missing data visualisations

It can be difficult to get a handle on where the missing values are in your data, and here is where visualisation can really help.

The function `vis_miss()` creates an overview visualisation of the missingness in the data. It also has options to cluster rows based on missingness, using `cluster = TRUE`; as well as options for sorting the columns, from most missing to least missing (`sort_miss = TRUE`).

# Task

Using the `riskfactors` dataset from `naniar`:

- Use `vis_miss()` to visualise the missingness in the data.
- Use `vis_miss()` with `cluster = TRUE` to explore some clusters of missingness.
- Use `vis_miss()` and sort the missings with `sort_miss` to arrange the columns by missingness.

## Hint
Use the options `cluster = TRUE` and `sort_miss = TRUE` inside of `vis_miss` to cluster and sort missing values

```{r}
# Visualise all of the missingness in the `riskfactors`  dataset
vis_miss(___)

# Visualise and cluster all of the missingness in the `riskfactors` dataset
vis_miss(___, ___ = TRUE)

# visualise and sort the columns by missingness in the `riskfactors` dataset
vis_miss(___, ___ = TRUE)
```

# Visualising missing cases and variables

To get a clear picture of the missingness across variables and cases, use `gg_miss_var()` and `gg_miss_case()`. These are the visual counterpart to `miss_var_summary()` and `miss_case_summary()`.

These can be split up into multiple plots with one for each category by choosing a variable to facet by.

## Task
Using the `riskfactors` dataset:

- Visualise the number of missings in cases using `gg_miss_case()`.
- Explore the number of missings in cases using `gg_miss_case()` and facet by the variable `education`.
- Visualise the number of missings in variables using `gg_miss_var()`.
- Explore the number of missings in variables using `gg_miss_var()` and facet by the variable `education`.

What do you notice in the visualisations of the whole data compared to the faceting?

### Hint

Use `gg_miss_case(riskfactors)` to get started, and `gg_miss_case(riskfactors, facet = education)` to facet by education.

```{r}
# Visualise the number of missings in cases using `gg_miss_case()`
gg_miss_case(___)

# Explore the number of missings in cases using `gg_miss_case()` and facet by the variable `education`
gg_miss_case(___, facet = ___)

# Visualise the number of missings in variables using `gg_miss_var()`
gg_miss_var(___)

# Explore the number of missings in variables using `gg_miss_var()` and facet by the variable `education`
___(___, facet = ___)
```

# Visualising missingness patterns

Let's practice a few different ways to visualise patterns of missingness using:

* `gg_miss_upset()` to give an overall pattern of missingness.
* `gg_miss_fct()` for a dataset that has a factor of interest: marriage.
* and `gg_miss_span()` to explore the missingness in a time series dataset.

What do you notice with the missingness and the facetting in the data?

## Task

- Explore missingness pattern of the `airquality` dataset with `gg_miss_upset()`.
- Explore how the missingness changes in the `riskfactors` dataset across the `marital` variable using `gg_miss_fct()`.

Using the `pedestrian` dataset:

- Explore how the missingness of the `hourly_counts` changes over a span of 3000 (you can also try different spans from 2000-5000).
- Explore the impact of `month` by facetting by `month`, and explore how missingness changes for a span of 1000.

### Hint

You will use the arguments `var = hourly_counts`, `span_every = 1000`, and `facet = month` inside of the `gg_miss_span()` function.

```{r}
# Using the airquality dataset, explore the missingness pattern using gg_miss_upset()
gg_miss_upset(airquality)

# With the riskfactors dataset, explore how the missingness changes across the marital variable using gg_miss_fct()
gg_miss_fct(x = riskfactors, fct = marital)

# Using the pedestrian dataset, explore how the missingness of hourly_counts changes over a span of 3000 
gg_miss_span(pedestrian, var = ___, span_every = ___)

# Using the pedestrian dataset, explore the impact of month by facetting by month
# and explore how missingness changes for a span of 1000
____(___, var = ___ , span_every = ___, facet = ___)
```
