---
title: ""
author: ""
output: html_document
---

```{r setup}
library(tidyverse)
```


# Continuing on from the lecture

```{r message=FALSE}
library(rvest)

page <- read_html("http://www.imdb.com/chart/top")

titles <- page %>%
  html_nodes(".titleColumn a") %>%
  html_text()

years <- page %>%
  html_nodes(".secondaryInfo") %>%
  html_text() %>%
  str_replace("\\(", "") %>% # remove (
  str_replace("\\)", "") %>% # remove )
  as.numeric()

scores <- page %>%
  html_nodes("#main strong") %>%
  html_text() %>%
  as.numeric()
  
imdb_top_250 <- tibble(
  title = titles, 
  year = years, 
  score = scores
  )

imdb_top_250
```


# Clean up / enhance

May or may not be a lot of work depending on how messy the data are

- See if you like what you got:

```{r}
glimpse(imdb_top_250)
```


- Add a variable for rank

```{r}
imdb_top_250 <- imdb_top_250 %>%
  mutate(
    rank = 1:nrow(imdb_top_250)
  )

imdb_top_250
```


# Answer the question: Which 1995 movies made the list?

```{r}
# perhaps use filter?
imdb_top_250 
```


# How would you go about answering this question: Which years have the most movies on the list?

```{r}
# perhaps group by year, then do some summaries on the data?
imdb_top_250 %>% 
  group_by(year) 
# other things here
```

# Visualize

How would you go about creating this visualization: Visualize the average yearly score for movies that made it on the top 250 list over time.

```{r echo=FALSE}
# Find the average score for each year
# perhaps group by year and take the average score?
imdb_yearly_avg_score <- imdb_top_250  %>%
  group_by(year) %>%
  summarise()
# do some more data analysis here 

# Then plot this with the year on x axis, and average score on the y axis
ggplot(aes()) +
    geom_point() +
    geom_smooth(method = "lm")
```


# Try exploring another IMDB table

- top TV shows: https://www.imdb.com/chart/toptv
- boxoffice weekend: https://www.imdb.com/chart/boxoffice
- worst films: https://www.imdb.com/chart/bottom
- Or other genres - like Action films: https://www.imdb.com/search/title/?genres=action

Choose one of these and get it into a similar format to the `imdb_top_250` data

- try answering questions like:
  - Which years have the most 
  - Which films made the least amount of money from the box office?
  - What is the oldest, worst film?
- Have some fun with it - define your own questions and explore the data yourself!
  

```{r}

```


# Extension Example: Crossrates 

An example we have seen is the cross rates data available at [https://openexchangerates.org/](https://openexchangerates.org/). To access this data you need to:

1. Get a free plan from https://openexchangerates.org/signup/free
2. Tell this function your API key -- `Sys.setenv("OER_KEY" = "your-key-here")`

Then you can access the data using a command like:

```
u <- sprintf(
    "https://openexchangerates.org/api/historical/%s.json?app_id=%s",
    day, Sys.getenv("OER_KEY")
  )
res <- jsonlite::fromJSON(u)
```

There's a nice help page here: https://gist.github.com/njtierney/0fe2daa11292e046f83fec7d8e9a8164

## Pull the crossrates for the month of August 2018

```{r}

```

## Make a plot to compare the crossrates for AUD, EUR, GBP, NZD, JPY. Make a descriptive summary of what you learn.

# Cricket data

Using the information from this vignette here: https://docs.ropensci.org/cricketdata/

## Use the cricketdata package to extract the T20 batting data of male players for your two favorite countries.

```{r}
library(cricketdata)

```


## Between 2010 and 2015, which country has the better record? Explain the reason for your answer, and your analysis choices.

```{r}

```
