---
title: "ETC1010: Data Modelling and Computing"
author: "Professor Di Cook, EBS, Monash U."
output: 
  learnr::tutorial:
    css: "css/logo.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      collapse = TRUE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center",
                      cache = FALSE)
tutorial_html_dependency()
```

# Wrangling your data

## Overview

- filter
- select
- mutate
- group_by
- summarise
- arrange
- count, tally

### *FOLLOW ALONG WITH THE NOTES* 

- Open your ETC1010 `Rproj`
- Download the `Rmd` file from the class web site
- You should already have the data sets that you need, from last class

## Side-note: Package system

We have been loading the `tidyverse` package. Its actually a suite of packages, and you can learn more about the individual packages at https://www.tidyverse.org. You could load each individually.

Because so many people contribute packages to R, it is a blessing and a curse. The best techniques are available, but there can be conflicts between function names. When you load tidyverse it prints a great summary of conflicts that it knows about, between its functions and others.

```{r message=TRUE}
library(tidyverse)
```

For example, there is a `filter` function in the `stats` package that comes with the R distribution. This can cause confusion when you want to use the filter function in `dplyr` (part of tidyverse). To be sure the function you use is the one you want to use, you can prefix it with the package name, `dplyr::filter()`.


