---
title: "Assignment 2"
output: html_document
author: Data Lads
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr);
library(dplyr);
library(tidyverse);
library(readxl)
library(forcats)
library(plotly)
library(stringr)
# First step is the bring in the data that needs to be tidied
survey <- read_csv("data/survey.csv", na = c("n/a", "NA")) %>%
  filter(id != 64)

#Q1
survey <- survey %>%
  mutate(Q1 = sub("A", "1", Q1)) %>%
  mutate(Q1 = sub("B", "2", Q1)) %>%
  mutate(Q1 = sub("C", "3", Q1)) %>%
  mutate(Q1 = sub("D", "4", Q1)) %>%
  mutate(Q1 = sub("E", "other", Q1))

#Q2
survey <- survey %>%
  mutate(Q2 = sub("1/2", "0.5", Q2)) %>%
  mutate(Q2 = sub(" hours", "", Q2)) %>%
  mutate(Q2 = sub("\n", "", Q2)) %>%
  mutate(Q2 = sub("hrs", "", Q2)) %>%
  mutate(Q2 = sub("HOUR", "", Q2)) %>%
  mutate(Q2 = sub("hours", "", Q2)) %>%
  mutate(Q2 = sub("4/7", "0.57", Q2)) %>%
  mutate(Q2 = sub("three times", "", Q2)) %>%
  mutate(Q2 = sub("(to uni)","", Q2))
survey$Q2[10] <- NA 
survey$Q2[27] <- "2.0" 
survey <- survey %>%
  mutate(Q2 = as.double(Q2)) %>%
  mutate(Q3 = as.logical(Q3))

#Q4
survey <- survey %>%
  mutate(Q4 = sub("&","n", Q4)) %>%
  mutate(Q4 = sub("Hiphop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip-Pop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip-hop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("hip-pop", "hip hop", Q4)) %>%
  mutate(Q4 = sub("music", "", Q4)) %>%
  mutate(Q4 = sub("alt", "alternative", Q4)) %>%
  mutate(Q4 = sub("\n", "", Q4)) %>%
  mutate(Q4 = sub("popular", "pop", Q4))
survey$Q4[22] <- NA
survey$Q4[29] <- "rock"
survey$Q4[46] <- NA
survey$Q4[51] <- "lo fi"
survey$Q4 <- tolower(survey$Q4)

#Q5
survey <- survey %>%
  mutate(Q5 = sub("hrs", "", Q5)) %>%
  mutate(Q5 = sub("hours", "", Q5)) %>%
  mutate(Q5 = sub("Hours\n", "", Q5)) %>%
  mutate(Q5 = sub("Choosing not to answer", "", Q5)) %>%
  mutate(Q5 = sub("\n", "", Q5)) %>%
  mutate(Q5 = sub("HOUR", "", Q5)) %>%
  mutate(Q5 = as.integer(Q5))
survey$Q5[40] <- NA
survey$Q5[46] <- NA

#Q6
survey$Q6 <- tolower(survey$Q6)
survey$Q6[1] <- "speed"
survey$Q6[12] <- "time travel"
survey$Q6[14] <- "forecasting"
survey$Q6[28] <- "control time"
survey$Q6[30] <- "not needing sleep"
survey$Q6[34] <- "forecasting"
survey$Q6[36] <- "control time"
survey$Q6[42] <- "control time"
survey$Q6[55] <- "control time"
survey$Q6[57] <- "time travel"
survey <- survey %>%
  mutate(Q6 = sub("\n","", Q6)) %>%
  mutate(Q6 = sub("teleporting", "teleportation", Q6)) %>%
  mutate(Q6 = sub("flying", "flight", Q6)) %>%
  mutate(Q6 = sub("fly", "flight", Q6)) %>%
  mutate(Q6 = sub("invisible", "invisibility", Q6))

#Q9
survey$Q9 <- tolower(survey$Q9)
survey <- survey %>%
  mutate(Q9 = sub("\n", "", Q9)) %>%
  mutate(Q9 = sub("no idea\n", "none", Q9)) %>%
  mutate(Q9 = sub("No television", "none", Q9)) %>%
  mutate(Q9 = sub("i do not watch television", "none", Q9)) %>%
  mutate(Q9 = sub("i don't watch television.", "none", Q9)) %>%
  mutate(Q9 = sub("nothing", "none", Q9)) %>%
  mutate(Q9 = sub("don't watch television", "none", Q9)) %>%
  mutate(Q9 = sub("i don't", "none", Q9)) %>%
  mutate(Q9 = sub("dont watch", "none", Q9)) %>%
  mutate(Q9 = sub("i don't watch tv", "none", Q9)) %>%
  mutate(Q9 = sub("non", "none", Q9)) %>%
  mutate(Q9 = sub("dont watch tv", "none", Q9)) %>%
  mutate(Q9 = sub("nonee.", "none", Q9))

#Q10
survey <- survey %>%
  mutate(Q10 = sub("\n", "", Q10)) %>%
  mutate(Q10 = as.double(Q10))
survey$Q10[40] <- NA

#Q13
survey <- survey %>%
  mutate(Q13 = sub("dollars", "", Q13)) %>%
  mutate(Q13 = sub("\n", "", Q13))
survey$Q13[25] <- "30"
survey$Q13[26] <- "5"
survey$Q13[40] <- "15"
survey$Q13[48] <- "30"
survey$Q13[58] <- "35"
survey <- survey %>%
  mutate(Q13 = as.double(Q13))

#Q18
survey$Q18 <- tolower(survey$Q18)
survey <- survey %>%
  mutate(Q18 = sub("\n", "", Q18))

#Q20
survey <- survey %>%
  mutate(Q20 = as.logical(Q20))

#Q21
survey <- survey %>%
  mutate(Q21 = sub("\n", "", Q21)) %>%
  mutate(Q21 = sub(";", ".", Q21)) %>%
  mutate(Q21 = as.double(Q21))

#Q22
survey <- survey %>%
  mutate(Q22 = sub("days", "", Q22)) %>%
  mutate(Q22 = as.integer(Q22))

#Q23
survey <- survey %>%
  mutate(Q23 = sub("XXI", "21", Q23)) %>%
  mutate(Q23 = sub("1997", "21", Q23)) %>%
  mutate(Q23 = as.integer(Q23))

#Q24
survey <- survey %>%
  mutate(Q24 = sub("I don't know", NA, Q24)) %>%
  mutate(Q24 = as.integer(Q24))

#Q25
survey$Q25 <- tolower(survey$Q25)
survey$Q25[45] <- "swimming"
survey <- survey %>%
  mutate(Q25 = sub("yes", NA, Q25)) %>%
  mutate(Q25 = sub("choosing not to answer", NA, Q25)) %>%
  mutate(Q25 = sub("\n", "", Q25))

#Q32
survey$Q32 <- tolower(survey$Q32)
survey$Q32[50] <- "chopsticks"

#Q33
survey <- survey %>%
  mutate(Q33 = as.logical(Q33))

survey
write.csv(survey, "data/survey_tidy1.csv")



```
# Tidying the data
The first step in tidying the data was importing it into R studio as a csv file. After this was done, we were able to
manipulate the data as we saw fit in order to produce a data set that we could extrapolate and view in meaningful ways.
Essentially, we started at Q1 and working through the data consecutively to Q33. As many of the questions were short
answer, there were many answers that had the same outcome but were written differently. Therefore, we used the “mutate”
function in order to manipulate the data. By mutating the data we were able to make the similar answers such as “Hip
hop” and “Hip-hop” read as “hip hop”. Additionally, Q9 produced multiple answers from students stating that they did not
watch TV in multiple different ways. In this case, it was easier to change these answer to “none” by using the notation
“survey$Q9[10] <- “none””.

```{r}
# Bar chart of amount of people in which year
p1 <- survey %>%
  mutate(Q26 = sub("A", "female", Q26)) %>%
  mutate(Q26 = sub("B", "male", Q26)) %>%
  mutate(Q26 = sub("C", "prefer not to say", Q26))
colnames(p1)[colnames(p1)=="Q26"] <- "Gender"
p2 <- ggplot(p1, aes(x = Q1, fill = Gender)) + 
  geom_bar() + 
  xlab("Year") + 
  theme(legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid", colour ="darkblue"))
ggplotly(p2)

```
#Plot 1
In the first plot, we have created a bar chart, which shows the number of people in their respective years of
university, whilst showing the genders that fall in each year. From the plot, it can be seen that the majority of
students are in their second or third year of university. Additionally, the graph provides evidence that there are more
males than females that completed the survey.

```{r}
# Box plot of height of people depending on gender
p1 <- survey %>%
  mutate(Q26 = sub("A", "female", Q26)) %>%
  mutate(Q26 = sub("B", "male", Q26)) %>%
  mutate(Q26 = sub("C", "prefer not to say", Q26))
colnames(p1)[colnames(p1)=="Q26"] <- "Gender"
p3 <- ggplot(p1, aes(x = Gender, y = Q24, fill = Gender)) + 
  geom_boxplot() + 
  xlab("Gender") + 
  ylab("Height") +
  theme(legend.background = element_rect(fill="lightblue", size=0.5, linetype="solid", colour ="darkblue"))
ggplotly(p3)
```
#Plot 2
The second plot is a box plot that shows the classes heights as a function of gender. The plot clearly shows that
average height of males is greater than that of females. Additionally when looking at both plot 1 and 2, it could be
seen that there is evidence to support the hypothesis that the person who chose not to disclose their gender is male.
From plot 1, the person of unknown gender is currently studying in a year, higher than that of fourth year. In this time
period there is a greater chance that those in this other bracket are male. Furthermore, looking at plot 2 the height of
the person with unknown gender is far greater than that of the average of females and just greater than that of males.
Both these pieces of information could be used to support the hypothesis that the unknown gender is in fact male.
