---
title: "Exploring French Fries"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output: html_document
---

First we load the tidyverse

```{r setup}
library(tidyverse)
```

Then, we read in the french fries data using `read_csv()`:

```{r read-fries}
french_fries <- read_csv("data/french_fries.csv")
```

Now we get the data into long form:

```{r}
fries_long <- french_fries %>% 
  gather(key = type, 
         value = rating, 
         -time, 
         -treatment, 
         -subject, 
         -rep) %>%
  mutate(type = as.factor(type))
```


Filter the french fries data to have:

- only week 1

```{r}
fries_long %>% filter(time == 1)
```

- weeks 1-4 only

```{r}
fries_long %>% filter(time %in% c("1", "2", "3", "4"))
```

- oil type 1 (oil type is called treatment)

```{r}
fries_long %>% filter(treatment == 1)
```

- oil types 1 and 3 but not 2

```{r}
fries_long %>% filter(treatment != 2)
```

# End of first example - wait until the next section.

# Using `select()`

- choose time, treatment and rep

```{r}
fries_long %>% select(time, treatment, rep)
```

- choose subject through to rating

```{r}
fries_long %>% select(subject:rating)
```

- drop subject


```{r}
fries_long %>% select(-subject)
```

# Wait until next example


# Learn to use `mutate()`

- For the french fries data compute a new variable called lrating by taking a log of the rating

```{r}
fries_long %>%
  mutate(lrating=log10(rating))
```

# Wait until next example

# group by and summarise

- Compute the average rating by subject

```{r}
fries_long %>% 
  group_by(subject) %>%
  summarise(rating = mean(rating, na.rm=TRUE))
```

- Compute the average rancid rating per week
```{r}
fries_long %>% 
  filter(type == "rancid") %>%
  group_by(time) %>%
  summarise(rating = mean(rating, na.rm=TRUE))
```

# `count()`

- count the number of subjects

```{r fries-count-subject}
fries_long %>% count(subject, sort=TRUE)
```

- cound the number of types

```{r fries-count-type}
fries_long %>% count(type)
```

# Using the skills we have learned to problem solve:

## Are ratings similar?

```{r fries-similar-ratings}
fries_long %>% 
  group_by(type) %>%
  summarise(m = mean(rating, 
                     na.rm = TRUE), 
            sd = sd(rating, 
                    na.rm = TRUE)) %>%
  arrange(-m)
```


The scales of the ratings are quite different. Mostly the chips are rated highly on potato'y, but low on grassy. 


```{r plot-fries-type-rating}
ggplot(fries_long,
       aes(x = type, 
           y = rating)) +
  geom_boxplot()
```

## Are reps like each other?

```{r fries-spread}
fries_spread <- fries_long %>% 
  spread(key = rep, 
         value = rating)
  
fries_spread
```


```{r fries-spread-summarise}
summarise(fries_spread,
          r = cor(`1`, `2`, use = "complete.obs"))
```


```{r fries-spread-plot, out.width= "80%"}
  ggplot(fries_spread,
         aes(x = `1`, 
             y = `2`)) + 
  geom_point() + 
  labs(title = "Data is poor quality: the replicates do not look like each other!")
```


## Replicates by rating type

```{r fries-group-summarise-cor}
fries_spread %>%
  group_by(type) %>%
  summarise(r = cor(x = `1`, 
                    y = `2`, 
                    use = "complete.obs"))
```


```{r plot-fries-facet-type, out.width = "90%"}
ggplot(fries_spread, aes(x=`1`, y=`2`)) + 
  geom_point() + facet_wrap(~type, ncol = 5)
```

Potato'y and buttery have better replication than the other scales, but there is still a lot of variation from rep 1 to 2.

