---
title: "Assignment 3"
author: "Group DQ , Dewei Qiu"
date: "2018年4月21日"
output: html_document
---

```{r}
library(tidyverse)
houses <- read_csv("data/Melbourne_housing_FULL.csv")
```

```{r}
library(visdat)
library(naniar)


houses <- houses %>% 
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom) 
vis_dat(houses)
summary(houses)
miss_summary(houses)
```

#missing value
```{r}
s_miss <- miss_summary(houses)
s_miss$miss_df_prop
s_miss$miss_var_prop
s_miss$miss_case_table
#by group
s_miss_group <- houses %>% 
  group_by(Suburb) %>% miss_summary()
s_miss_group$miss_case_table
vis_miss(houses, sort_miss=TRUE) + theme(aspect.ratio=1)
Houses_true<- houses %>% na.exclude(houses)
```
#Overall statistics, 11.5% of possible values are missing. That’s alot missing. Bathroom, Bedroom2 and Price, have a large number of missings, which more than 20%. But Distance, Rooms and Type, have no missings. Have more than 20% missing which is too much for us, we will check if the model ignore this three variables will perform good or not.

```{r}
tao_shadow <- bind_shadow(houses)
tao_shadow
ggplot(data = tao_shadow,
       aes(x = wind_ew, y=wind_ns, colour=air_temp_c_NA)) +
       scale_colour_brewer(palette="Dark2") +
       geom_point(alpha=0.7) + theme(aspect.ratio=1) 
```


```{r}
p1 <- ggplot(Houses_true, aes(x=Rooms, y=Price)) + geom_point() + 
  geom_smooth(method="lm", se=FALSE)
p2 <- ggplot(Houses_true, aes(x=Type, y=Price)) + geom_point() + 
  geom_smooth(method="lm", se=FALSE)
p3 <- ggplot(Houses_true, aes(x=Distance, y=Price)) + geom_point() + geom_smooth(method="lm", se=FALSE)
p4 <- ggplot(Houses_true, aes(x=Bedroom2, y=Price)) + geom_point() + geom_smooth(method="lm", se=FALSE)
p5 <- ggplot(Houses_true, aes(x= Bathroom, y=Price)) + geom_point() + geom_smooth(method="lm", se=FALSE)
p1
p2
p3
p4
p5
```

Price: Price in dollars
Rooms: Number of rooms
Type: br - bedroom(s); h - house,cottage,villa, semi,terrace; u - unit, duplex; t - townhouse; dev site - development site; o res - other residential.
Distance: Distance from CBD
Bedroom2 : Scraped # of Bedrooms (from different source)
Bathroom: Number of Bathrooms


##First we check if we use only Distance, Rooms and Type three variables

```{r}
library(broom)

mod1 <- lm(Price ~  Type + Distance, data=Houses_true)
tidy(mod1)
summary(mod1)
```

```{r}
library(broom)

mod2 <- lm(Price ~ Type * Distance, data=Houses_true)
summary(mod2)
```
#Both model with/without interact term have good P-value, but adjusted R^2 for them only 0.2336 and 0.2504 which too low. They only explained 25% of the data value


##THan we try to compare the non interaction mod with all variables
```{r}
mod3 <- lm(Price ~  Rooms + Type + Distance + Bedroom2 + Bathroom, data=Houses_true)
summary(mod3)
```
#Form result, this adjusted R^2 increase to 0.418, and we try to reject the bedrrom 2 as it greater than 0.05 P-value.

```{r}
mod4 <- lm(Price ~  Rooms + Type + Distance + Bathroom, data=Houses_true)
summary(mod4)
```

#However when we drop the varibale "bedroom2", the R^2 value and adjusted R^2 are same as before. we keep the "bedroom2" first and more to interaction model


##check the correlation between each variables
```{r}
library(GGally)
ggpairs(houses)
```
#The "Bedroom2" and "Room" has 0.947 correlation which very high interate each other
#The "Rooms" and "Bathroom" has 0.612 correlation which higher than 0.5.  
#The "Bedroom2" and "Bathroom" has 0.615 correlation also higher than 0.5, we will check them either has or not interacte.  

##show them graph
```{r}
library(viridis)

p_1 <- ggplot(Houses_true, aes(x=Bedroom2, y=Rooms, colour=y1)) + 
  geom_point(size=5, alpha=0.5) + 
  scale_colour_viridis() + 
  theme_bw() + theme(aspect.ratio=1, legend.position="none") +
  ggtitle("Bedroom2 and room")

p_2 <- ggplot(Houses_true, aes(x=Bedroom2, y=Bathroom, colour=y2)) + 
  geom_point(size=5, alpha=0.5) + 
  scale_colour_viridis() + 
  theme_bw() + theme(aspect.ratio=1, legend.position="none") +
  ggtitle("Bedroom2 and bathroom")

p_3 <- ggplot(Houses_true, aes(x=Bedroom2, y=Bathroom, colour=y3)) + 
  geom_point(size=5, alpha=0.5) + 
  scale_colour_viridis() + 
  theme_bw() + theme(aspect.ratio=1, legend.position="none") +
  ggtitle("Bedroom2 and Bathroom")

grid.arrange(p_1, p_2, p_3, ncol=3) 
```


##only two interact variable
```{r}
mod5 <- lm(Price ~  Type + Distance + Bathroom + Rooms * Bedroom2, data=Houses_true)
summary(mod5)
```
adjusted R^2=0.419 

##three interact variable
```{r}
mod6 <- lm(Price ~  Type + Distance + Rooms * Bathroom * Bedroom2, data=Houses_true)
summary(mod6)
```
adjusted R^2=0.4248

##four interact variable
```{r}
mod7 <- lm(Price ~  Type + Distance * Rooms * Bathroom * Bedroom2 , data=Houses_true)
summary(mod7)
```
adjusted R^2=0.4506

```{r}
mod8 <- lm(Price ~  Type * Distance + Bathroom  + Rooms * Bedroom2 , data=Houses_true)
summary(mod8)
```

