---
title: "ETC1010_Assignment_3"
author: "Raymond & Co"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r library}
library(tidyverse)
library(visdat)
library(naniar)
library(broom)
library(labelled)
library(gridExtra)
```

```{r load_data}
data <- read_csv("Melbourne_housing_FULL.csv")
house <- data%>%select(Price,Rooms,Type,Distance,Bedroom2,Bathroom)
```

## Handle missing value

```{r miss_var_type}
vis_dat(house, palette = "cb_safe")
```

```{r miss_perc}
vis_miss(house, sort_miss=TRUE) + theme(aspect.ratio=1)
```

```{r}
s_miss <- miss_summary(house)
s_miss$miss_df_prop
s_miss$miss_var_summary
```

### 1.Drop the observations missin in "Price".
```{r drop_price}
house_clean <- house%>%
  filter(!is.na(Price))
```

### 2.The observation missing value in "Distance" also missing in "Bedroom2" and "Bathroom", therefore, we drop it.
```{r drop_na_distance}
house_clean%>%filter(is.na(Distance))
house_clean <- house_clean%>%filter(!is.na(Distance))
```

```{r miss_summary}
s_miss_2 <- miss_summary(house_clean)
s_miss_2$miss_case_table
s_miss_2$miss_var_summary
```

It implies all the observations missing in "Bedroom2" also missing in "Bathroom" after cleaning "Price" and "Distance". Only 6 observations missing in "Bathroom" didn't miss in "Bedroom2".

```{r shadow_matrix}
house_shadow <- bind_shadow(house_clean)
house_shadow
```

```{r relationship_other_var}
ggplot(data = house_shadow, aes(x = Distance, y=Rooms, colour=Bedroom2_NA)) +
  scale_colour_brewer(palette="Dark2") +
  facet_wrap(~Type)+
  geom_point(alpha=0.7) + theme(aspect.ratio=1) 
```

```{r}
s_miss_group <- house_clean %>% 
  group_by(Type) %>% miss_summary()
s_miss_group$miss_case_table
```

```{r Distance_vs_Bathroom}
ggplot(house_clean,
       aes(x = Distance,
           y = Bathroom)) +
  scale_colour_brewer(palette="Dark2") +
  geom_miss_point()+
  facet_wrap(~Type) +
  theme(aspect.ratio=1)
```

```{r Rooms_vs_Bathroom}
ggplot(house_clean,
       aes(x = Rooms,
           y = Bathroom)) +
  scale_colour_brewer(palette="Dark2") +
  geom_miss_point()+
  facet_wrap(~Type) +
  theme(aspect.ratio=1)
```

```{r eval=FALSE}
library(impute)
house_impute_h <- house_shadow %>%
  arrange(Type, Rooms, Bathroom) %>%
  filter(Type=="h") %>%
  select(Rooms, Bathroom) 
house_impute_h <- impute.knn(as.matrix(house_impute_h), 10)

house_impute_t <- house_shadow %>%
  arrange(Type, Rooms, Bathroom) %>%
  filter(Type=="t") %>%
  select(Rooms, Bathroom) 
house_impute_t <- impute.knn(as.matrix(house_impute_t), 10)

house_impute_u <- house_shadow %>%
  arrange(Type, Rooms, Bathroom) %>%
  filter(Type=="u") %>%
  select(Rooms, Bathroom) 
house_impute_u <- impute.knn(as.matrix(house_impute_u), 10)

house_impute <- rbind(house_impute_h, house_impute_t$data, house_impute_u)
house_shadow_2 <- house_shadow %>%
  arrange(Type, Rooms, Bathroom)

house_shadow_2 <- house_shadow_2%>%
  mutate(Rooms = house_impute$Rooms,
         Bathroom = house_impute$Bathroom)
```

```{r eval=FALSE}
ggplot(house_shadow_2,
       aes(x = Rooms,
           y = Bathroom, 
           colour=Bathroom_NA)) +
  geom_point(alpha=0.7) + 
  facet_wrap(~Type) + 
  scale_colour_brewer(palette="Dark2") +
  theme(aspect.ratio=1)
```

```{r}
Bath_h <- house_shadow %>% 
  filter(Type=="h") %>%
  mutate(Bathroom = ifelse(is.na(Bathroom), 
                             mean(Bathroom, na.rm=TRUE),
                             Bathroom))

Bath_t <- house_shadow %>% 
  filter(Type=="t") %>%
  mutate(Bathroom = ifelse(is.na(Bathroom), 
                             mean(Bathroom, na.rm=TRUE),
                             Bathroom))

Bath_u <- house_shadow %>% 
  filter(Type=="u") %>%
  mutate(Bathroom = ifelse(is.na(Bathroom), 
                             mean(Bathroom, na.rm=TRUE),
                             Bathroom))

house_shadow_3 <- rbind(Bath_h, Bath_t, Bath_u)
```

```{r Rooms_vs_Bathroom_3}
ggplot(house_shadow_3,
       aes(x = Rooms,
           y = Bathroom, 
           colour=Bathroom_NA)) +
  geom_point(alpha=0.7) + 
  facet_wrap(~Type) + 
  scale_colour_brewer(palette="Dark2") +
  theme(aspect.ratio=1)
```

```{r Distance_vs_Bathroom_3}
ggplot(house_shadow_3,
       aes(x = Distance,
           y = Bathroom, 
           colour=Bathroom_NA)) +
  geom_point(alpha=0.7) + 
  facet_wrap(~Type) + 
  scale_colour_brewer(palette="Dark2") +
  theme(aspect.ratio=1)
```

### 3.Fill missing "Bathroom" by the mean of each type.

```{r fill_Bathrooom_type_mean}
Bath_clean_h <- house_clean %>% 
  filter(Type=="h") %>%
  mutate(Bathroom = ifelse(is.na(Bathroom), 
                             mean(Bathroom, na.rm=TRUE),
                             Bathroom))

Bath_clean_t <- house_clean %>% 
  filter(Type=="t") %>%
  mutate(Bathroom = ifelse(is.na(Bathroom), 
                             mean(Bathroom, na.rm=TRUE),
                             Bathroom))
Bath_clean_u <- house_clean %>% 
  filter(Type=="u") %>%
  mutate(Bathroom = ifelse(is.na(Bathroom), 
                             mean(Bathroom, na.rm=TRUE),
                             Bathroom))

house_clean <- rbind(Bath_clean_h, Bath_clean_t, Bath_clean_u)
```

```{r Rooms_vs_Bedroom2}
ggplot(house_clean,
       aes(x = Rooms,
           y = Bedroom2)) +
  scale_colour_brewer(palette="Dark2") +
  geom_miss_point()+
  facet_wrap(~Type) +
  theme(aspect.ratio=1)
```

```{r Distance_vs_Bedroom2}
ggplot(house_clean,
       aes(x = Distance,
           y = Bedroom2)) +
  scale_colour_brewer(palette="Dark2") +
  geom_miss_point()+
  facet_wrap(~Type) +
  theme(aspect.ratio=1)
```

```{r}
Bed_h <- house_shadow %>% 
  filter(Type=="h") %>%
  mutate(Bedroom2 = ifelse(is.na(Bedroom2), 
                             mean(Bedroom2, na.rm=TRUE),
                             Bedroom2))

Bed_t <- house_shadow %>% 
  filter(Type=="t") %>%
  mutate(Bedroom2 = ifelse(is.na(Bedroom2), 
                             mean(Bedroom2, na.rm=TRUE),
                             Bedroom2))

Bed_u <- house_shadow %>% 
  filter(Type=="u") %>%
  mutate(Bedroom2 = ifelse(is.na(Bedroom2), 
                             mean(Bedroom2, na.rm=TRUE),
                             Bedroom2))

house_shadow_4 <- rbind(Bed_h, Bed_t, Bed_u)
```

```{r Rooms_vs_Bedroom2_4}
ggplot(house_shadow_4,
       aes(x = Rooms,
           y = Bedroom2, 
           colour=Bedroom2_NA)) +
  geom_point(alpha=0.7) + 
  facet_wrap(~Type) + 
  scale_colour_brewer(palette="Dark2") +
  theme(aspect.ratio=1)
```

```{r Distance_vs_Bedroom2_4}
ggplot(house_shadow_4,
       aes(x = Distance,
           y = Bedroom2, 
           colour=Bedroom2_NA)) +
  geom_point(alpha=0.7) + 
  facet_wrap(~Type) + 
  scale_colour_brewer(palette="Dark2") +
  theme(aspect.ratio=1)
```

### 4.Fill missing "Bedroom2" by the mean of each type.

```{r fill_Bedroom2_type_mean}
Bed_clean_h <- house_clean %>% 
  filter(Type=="h") %>%
  mutate(Bedroom2 = ifelse(is.na(Bedroom2), 
                             mean(Bedroom2, na.rm=TRUE),
                             Bedroom2))

Bed_clean_t <- house_clean %>% 
  filter(Type=="t") %>%
  mutate(Bedroom2 = ifelse(is.na(Bedroom2), 
                             mean(Bedroom2, na.rm=TRUE),
                             Bedroom2))
Bed_clean_u <- house_clean %>% 
  filter(Type=="u") %>%
  mutate(Bedroom2 = ifelse(is.na(Bedroom2), 
                             mean(Bedroom2, na.rm=TRUE),
                             Bedroom2))

house_clean <- rbind(Bed_clean_h, Bed_clean_t, Bed_clean_u)
```

```{r}
house_clean$Type <- as.factor(house_clean$Type)
```

```{r}
summary(house_clean)
str(house_clean)
```

##Relating Price to number of rooms:
```{r}
ggplot(house_clean, mapping = aes(x=Rooms, y= Price)) + geom_point()
```

##Relating Price to type:
```{r}
ggplot(house_clean, mapping = aes(x=Type, y= Price)) + geom_boxplot()

```
This plot suggests that houses have a larger price range than townhouses and units. The price range of townhouses is the smallest compared to the other types. Also, houses have a higher median price than townhouses, which have a higher median price than units.


##Relating Price to Distance:
```{r}
ggplot(house_clean, mapping = aes(x=Distance, y= Price)) + geom_point()
```
This plot is positively skewed. This means that there is a more dense popluation of houses with smaller distances from the CBD (between 1km and 19km) than larger distances. In general, the further away the house is from the CBD, the less expensive. There are a few outliers. For example, there is a very expensive house that is only 10km form the CBD. This suggests that there must be another factor or factors affecting the price, such as size of the house or amount of land.

##Relating Price to Bedrooms:
```{r}
ggplot(house_clean, mapping = aes(x=Bedroom2, y= Price)) + geom_point()
```
Most of the properties have between 3-5 bedrooms. Houses with less than 3 bedrooms and more than 5 bedrooms are less expensive than those with 3-5 bedrooms

##Relating Price to Bathroom:
```{r}
ggplot(house_clean, mapping = aes(x=Bathroom, y= Price)) + geom_point()
```
This plot is slightly positively skewed. This means that there is a higher number of houses with a lower number of bathrooms (1-4) than with a higher number of bathrooms (more than 6). 

## Interactions between Type other variables

### 1.Interactions between Rooms and Type

```{r Rooms_Type}
mod1_data <- house_clean%>%select(Price,Rooms,Type)
mod1_1 <- lm(Price~Rooms+Type, data = mod1_data)
mod1_2 <- lm(Price~Rooms*Type, data = mod1_data)
tidy(mod1_1)
tidy(mod1_2)
```

```{r}
a_mod1_1 <- augment(mod1_1, mod1_data)
ggplot(mod1_data, aes(x = Rooms, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod1_1, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 1")
```

```{r}
a_mod1_2 <- augment(mod1_2, mod1_data)
ggplot(mod1_data, aes(x = Rooms, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod1_2, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 2")
```

```{r}
p1 <- ggplot(a_mod1_1, aes(x=Rooms, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 1")
p2 <- ggplot(a_mod1_2, aes(x=Rooms, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 2")
grid.arrange(p1, p2, ncol=2)
```

```{r}
rbind(glance(mod1_1),glance(mod1_2))
```

The model 1 and 2 build base on "Rooms" and "Type". R.squared and adj.r.squared slightly increases in model2  but AIC and BIC increases in Model1. Therefore, the model1 without interaction is better.

### 2.Interactions between Distance and Type

```{r Distance_Type}
mod2_data <- house_clean%>%select(Price,Distance,Type)
mod2_1 <- lm(Price~Distance+Type, data = mod2_data)
mod2_2 <- lm(Price~Distance*Type, data = mod2_data)
tidy(mod2_1)
tidy(mod2_2)
```

```{r}
a_mod2_1 <- augment(mod2_1, mod2_data)
ggplot(mod2_data, aes(x = Distance, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod2_1, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 3")
```

```{r}
a_mod2_2 <- augment(mod2_2, mod2_data)
ggplot(mod2_data, aes(x = Distance, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod2_2, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 4")
```

```{r}
p3 <- ggplot(a_mod2_1, aes(x=Distance, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 3")
p4 <- ggplot(a_mod2_2, aes(x=Distance, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 4")
grid.arrange(p3, p4, ncol=2)
```

```{r}
rbind(glance(mod2_1),glance(mod2_2))
```

The model 3 and 4 build base on "Distance" and "Type". R.squared and adj.r.squared increases in model4, also AIC and BIC decrease in Model4. Therefore, the model4 with interaction is better.

### 3.Interactions between Bedroom2 and Type

```{r Bedroom2_Type}
mod3_data <- house_clean%>%select(Price,Bedroom2,Type)
mod3_1 <- lm(Price~Bedroom2+Type, data = mod3_data)
mod3_2 <- lm(Price~Bedroom2*Type, data = mod3_data)
tidy(mod3_1)
tidy(mod3_2)
```

```{r}
a_mod3_1 <- augment(mod3_1, mod3_data)
ggplot(mod3_data, aes(x = Bedroom2, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod3_1, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 5")
```

```{r}
a_mod3_2 <- augment(mod3_2, mod3_data)
ggplot(mod3_data, aes(x = Bedroom2, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod3_2, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 6")
```

```{r}
p5 <- ggplot(a_mod3_1, aes(x=Bedroom2, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 5")
p6 <- ggplot(a_mod3_2, aes(x=Bedroom2, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 6")
grid.arrange(p5, p6, ncol=2)
```

```{r}
rbind(glance(mod3_1),glance(mod3_2))
```

The model 5 and 6 build base on "Bedroom2" and "Type". R.squared and adj.r.squared slightly increases in model6 but AIC and BIC increases in Model6. Therefore, the model5 without interaction is better.

### 4.Interactions between Bathroom and Type

```{r Bathroom_Type}
mod4_data <- house_clean%>%select(Price,Bathroom,Type)
mod4_1 <- lm(Price~Bathroom+Type, data = mod4_data)
mod4_2 <- lm(Price~Bathroom*Type, data = mod4_data)
tidy(mod4_1)
tidy(mod4_2)
```

```{r}
a_mod4_1 <- augment(mod4_1, mod4_data)
ggplot(mod4_data, aes(x = Bathroom, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod4_1, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 7")
```

```{r}
a_mod4_2 <- augment(mod4_2, mod4_data)
ggplot(mod4_data, aes(x = Bathroom, y = Price, color = Type))+
  geom_point()+
  facet_wrap(~Type)+
  geom_line(data=a_mod4_2, aes(y=.fitted), color="Black")+
  theme(aspect.ratio=1)+ 
  ggtitle("Model 8")
```

```{r}
p7 <- ggplot(a_mod4_1, aes(x=Bathroom, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 7")
p8 <- ggplot(a_mod4_2, aes(x=Bathroom, y=.fitted,
                            colour=Type)) +
  geom_line()+
  ggtitle("Model 8")
grid.arrange(p7, p8, ncol=2)
```

```{r}
rbind(glance(mod4_1),glance(mod4_2))
```

The model 7 and 8 build base on "Bthroom" and "Type". R.squared and adj.r.squared slightly increases in model8. And AIC decreases but BIC increases in model8. Therefore, model8 with interaction is better.

## comb, interaction bwtween type and other variables

```{r}
mod9 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type, data = house_clean)
tidy(mod9)
```

```{r}
glance(mod9)
```

## Add interactions between Rooms and other vaiables to improve model

### 1. Add interaction between Bathroom and Rooms
```{r}
mod10 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bathroom*Rooms, data = house_clean)
tidy(mod10)
```

```{r}
rbind(glance(mod9),glance(mod10))
```

Drop mod10 base on BIC, drop"Bathroom*Rooms", use mod 9 to continue.

### 2. Add interaction between Bedroom2 and Rooms

```{r}
mod11 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms, data = house_clean)
tidy(mod11)
```

```{r}
rbind(glance(mod9),glance(mod11))
```

Improvement in mod11, use mod11 to continue.

### 3. Add interaction between Distance and Rooms

```{r}
mod12 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms, data = house_clean)
tidy(mod12)
```

```{r}
rbind(glance(mod11),glance(mod12))
```

Improvement in mod12, use mod12 to continue.

## Add interactions between Distance and other vaiables to improve model

### 1. Add interaction between Bathroom and Distance

```{r}
mod13 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance, data = house_clean)
tidy(mod13)
```

```{r}
rbind(glance(mod12),glance(mod13))
```

Improvement in mod13, use mod13 to continue.

### 2. Add interaction between Bathroom and Distance

```{r}
mod14 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance+Bedroom2*Distance, data = house_clean)
tidy(mod14)
```

```{r}
rbind(glance(mod13),glance(mod14))
```

Drop mod14 base on BIC, drop "Bedroom2*Distance', use mod13 to continue.

## Add interactions between Bedroom2 and Bathroom to improve model

```{r}
mod15 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance+Bedroom2*Bathroom, data = house_clean)
tidy(mod15)
```

```{r}
rbind(glance(mod13),glance(mod15))
```

Improvement in mod15, use mod15 to continue.

## Add 3-variables interactions to improve model

## 1.Type,Distance and Bathroom

```{r}
mod16 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance+Bedroom2*Bathroom+Type*Distance*Bathroom, data = house_clean)
tidy(mod16)
```

```{r}
rbind(glance(mod15),glance(mod16))
```

Improvement base on r.squared, adj.r.squared ,deviance and AIC, just slightly increase in BIC, Overall mod 16 is better, use mod 16 to continue

## 2.Type,Distance and Rooms

```{r}
mod17 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance+Bedroom2*Bathroom+Type*Distance*Bathroom+Type*Distance*Rooms, data = house_clean)
tidy(mod17)
```

```{r}
rbind(glance(mod16),glance(mod17))
```

Improvement base on r.squared, adj.r.squared ,deviance and AIC, just slightly increase in BIC, Overall mod 17 is better, use mod 17 to continue

## 2.Type,Distance and Bedroom2

```{r}
mod18 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance+Bedroom2*Bathroom+Type*Distance*Bathroom+Type*Distance*Rooms+Type*Distance*Bedroom2, data = house_clean)
tidy(mod18)
```

```{r}
rbind(glance(mod17),glance(mod18))
```

Not significant improvement in mod 18, drop"Type*Distance*Bedroom2", use mod 17 to continue.

### 4-variables interactions to improve model

### 1.Type*Distance*Rooms*Bathroom

```{r}
mod19 <- lm(Price~Type+Rooms+Bedroom2+Distance*Type+Bathroom*Type+Bedroom2*Rooms+Distance*Rooms+Bathroom*Distance+Bedroom2*Bathroom+Type*Distance*Bathroom+Type*Distance*Rooms+Type*Distance*Rooms*Bathroom, data = house_clean)
tidy(mod19)
```

```{r}
rbind(glance(mod17),glance(mod19))
```

Improvement in mod19 base on r.squared, adj.r.squared ,deviance, AIC and BIC.

## 5_variables interactions

```{r}
mod20 <- lm(Price~Type*Distance*Rooms*Bathroom*Bedroom2, data = house_clean)
tidy(mod20)
```

```{r}
rbind(glance(mod19),glance(mod20))
```

Basae on r.squared, adj.r.squared ,deviance, AIC and BIC, mod20 is the best model, but a few predictiors in the mod20 model is not significant.




