---
title: "ETC 1010 Assignment 1"
author: "Raymond & Co."
output: html_document
---

# Exercise

```{r echo=FALSE,warning=FALSE,message=FALSE}
library(readxl)
library(tidyverse)
library(plotly)
```

###Q1.Extract the "Airport Passengers" table into R with this code.

```{r Q1}
passengers <- read_xls("WebAirport_FY_1986-2017.xls", sheet=3, skip=6)
head(passengers)
```

###Q2 .Tidy the data, to produce a data set with these columns
####-airport: all of the airports. (You can remove the TOTAL, because this can be easily recalculated as needed.)
####-year 
####-type_of_flight: DOMESTIC, INTERNATIONAL
####-bound: IN or OUT
    
```{r Q2}
passengers_tidy<-passengers%>%
  select(AIRPORT,Year)%>%
  mutate(DOMESTIC_INBOUND = passengers$INBOUND, DOMESTIC_OUTBOUND = passengers$OUTBOUND, INTERNATIONAL_INBOUND = passengers$INBOUND__1, INTERNATIONAL_OUTBOUND = passengers$OUTBOUND__1)%>%
  gather(variable, count, -AIRPORT, -Year)%>%
  separate(variable, c("type_of_flight","bound"))
passengers_tidy<-passengers_tidy%>%
  filter(AIRPORT!="TOTAL AUSTRALIA")
head(passengers_tidy)
```

###Q3.Use your data to make 
####-a barchart of type of flight facetted by year.
####-a scatterplot of domestic inbound vs outbound passengers facetted by year. There are three airports with much higher numbers, which are these three?
####-a scatterplot of international inbound vs outbound passengers facetted by year

```{r Q3.1}
p1 <- ggplot(passengers_tidy, aes(x=Year, fill=type_of_flight, label=paste(count))) +
  geom_bar(aes(weight=count)) + 
  theme(axis.text.x = element_text(angle = 90),
        axis.title = element_text(colour = "blue",size = 12),
        axis.text=element_text(size=8)) +
  xlab("Year") + 
  ylab("Pessengers (ten of millions)")+
  scale_y_continuous(labels = c("0"="0","50000000"="5","100000000"="10","150000000"="15"))+
  scale_x_discrete(labels = c("1985-86"="85-86","1986-87"="86-87","1987-88"="87-88","1988-89"="88-89","1989-90"="89-90","1990-91"="90-91","1991-92"="91-92","1992-93"="92-93","1993-94"="93-94","1994-95"="94-95","1995-96"="95-96","1996-97"="96-97","1997-98"="97-98","1998-99"="98-99","1999-00"="99-00","2000-01"="00-01","2001-02"="01-02","2002-03"="02-03","2003-04"="03-04","2004-05"="04-05","2005-06"="05-06","2006-07"="06-07","2007-08"="07-08","2008-09"="08-09","2009-10"="09-10","2010-11"="10-11","2011-12"="11-12","2012-13"="12-13","2013-14"="13-14","2014-15"="14-15","2015-16"="15-16","2016-17"="16-17"))
ggplotly(p1)
``` 

```{r Q3.2}
DOMESTIC<-passengers%>%select(AIRPORT,Year,INBOUND,OUTBOUND)
DOMESTIC<-passengers%>%filter(AIRPORT!="TOTAL AUSTRALIA")
p2<-ggplot(DOMESTIC, aes( x = INBOUND, y = OUTBOUND, label = AIRPORT ))+
  geom_point()+
  facet_wrap(~Year, ncol =8)+
  ggtitle("Inbound Demostic Pessengers vs Outbound")+
  theme(axis.title = element_text(colour = "blue",size = 12),
        axis.text=element_text(size=8))+
  xlab("Inbound (millions)")+
  ylab("Outbound (millions)")+
  scale_y_continuous(breaks = c(0,5000000,10000000), labels = c("0","5","10"))+
  scale_x_continuous(breaks = c(0,5000000,10000000), labels = c("0","5","10"))
ggplotly(p2)
```

The three airport with higest passengers are SYDNEY, MELBOURNE and BRISBANE.

```{r Q3.3}
INTERNATIONAL<-passengers%>%select(AIRPORT,Year,INBOUND__1,OUTBOUND__1)
INTERNATIONAL<-passengers%>%filter(AIRPORT!="TOTAL AUSTRALIA")
p3<-ggplot(INTERNATIONAL, aes( x = INBOUND__1, y = OUTBOUND__1, label = AIRPORT))+
  geom_point()+
  facet_wrap(~Year, ncol =8)+
  ggtitle("Inbound International Pessengers vs Outbound")+
  xlab("Inbound (millions)")+
  ylab("Outbound (millions)")+
  theme(axis.title = element_text(colour = "blue",size = 12),
        axis.text=element_text(size=8))+
  scale_y_continuous(labels = c("0"="0","2000000"="2","4000000"="4","6000000"="6","8000000"="8"))+
  scale_x_continuous(labels = c("0"="0","2000000"="2","4000000"="4","6000000"="6","8000000"="8"))
ggplotly(p3)
```

###Q4.Write a paragraph explaining what you learn from each of the plots

From graph 1 we find that there are domestic passengers versue international passengers and the number of both domestic and international passengers were increasing between 1985 to 2017. 
From graph 2 and 3,both of them show that the amount of inbound passengers versus the outbound are similar in each airport. There is a strong, positive linear correlation between the domestic inbound and outbound flights. This also applies for the international inbound and outbound flights. For example, an airport with higher inbound passengers would have a higher number in outbound passengers. The reason might be the passengers normally purchase the inbound flight and outbound flight in the same airport for convenience. Also both of the top 3 airport with higerst passengers is SYDNEY, MELBOURNE and BRISBANE. But the difference is that SYDNEY and MELBOURNE airport has the similar number of passengers in domestic but SYDNEY get much more international passengers during 1985~2017.



