---
title: "Automating scraping with functions"
author: "YOUR NAME"
output: html_document
---

```{r setup}
library(tidyverse)
library(polite)
library(rvest)
```

# Scrape the list of most popular TV shows on IMDB: http://www.imdb.com/chart/tvmeter

```{r}
tv_url <- "http://www.imdb.com/chart/tvmeter"
tv_data <- bow(tv_url) %>% scrape()

library(janitor)
tv_tables <-tv_data %>%
  html_table()

# extract the first list
tv_list <- tv_tables[[1]] %>%
  clean_names() %>%
  select(-x,
         -x_2,
         -your_rating) %>%
  # extract out year
  separate(rank_title,
           into = c("title", "year"),
           sep = "\n") %>%
  # extract number from year
  mutate(year = parse_number(year)) 
```

## get urls for each show

```{r}
tv_html <- bow("http://www.imdb.com/chart/tvmeter") %>% scrape()

urls <- tv_html %>%
  html_nodes(".titleColumn a") %>%
  html_attr("href") %>%
  paste("http://www.imdb.com/", ., "sep" = "")

urls
```

## Examine each of the *first three* (or however many you can get through) tv show subpage to also obtain genre and runtime.


## Put the scraped show information into one function

```{r download-data}
scrape_show_info <- function(url){
  
  tv_data <- bow(url) %>% scrape()
  
  title <- tv_data %>%
    html_node("#title-overview-widget h1") %>%
    html_text() %>%
    str_trim()
  
  runtime <- tv_data %>%
    html_node("time") %>%
    html_text() %>%
    str_replace("\\n", "") %>%
    str_trim()
  
  genres <- tv_data %>%
    html_nodes(".txt-block~ .canwrap a") %>%
    html_text() %>%
    str_trim() %>%
    paste(collapse = ", ")
  
  tibble(title = title, runtime = runtime, genres = genres)

  }
```

## decompose the steps of `scrape_show_info` into smaller steps, to make individual functions that scrape title, runtime, and genres

```{r}
scrape_title <- function(url){
  
  
}

scrape_runtime <- function(url){
  
  
}

scrape_genre <- function(url){
  
  
}

scrape_show_info <- function(url){
 
  title <- scrape_title(url)
  runtime <- scrape_runtime(url)
  genre <- scrape_genre(url)
  
  tibble(title = title, 
         runtime = runtime, 
         genre = genre)
   
}

```


## Time permitting, also try to get the following:

  - First five plot keywords
  - Country
  - Language
  
Add this information to the data frame created with `scrape_show_info()`

```{r}
scrape_show_info <- function(url){
 
  title <- scrape_title(url)
  runtime <- scrape_runtime(url)
  genre <- scrape_genre(url)
  
  # extra show features
  # plot
  # country
  # language
  
  tibble(title = title, 
         runtime = runtime, 
         # and extra objects here
         genre = genre)
   
}
```

# Practice making functions

Using the dataset, `mtcars` as an example, try the following:

## Calculate the difference between the minimum and maximum value

```{r}
my_range <- function(x){
  
}

```


## calculate the range of every column in a dataset

```{r}
map(mtcars, ___)
```


## create your own mean function

## calculate the mean of every column in a dataset

