---
title: "ETC1010: Data Modelling and Computing"
author: "Professor Di Cook, EBS, Monash U."
output: 
  learnr::tutorial:
    css: "css/logo.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(emo) 
knitr::opts_chunk$set(echo = TRUE,   
                      message = FALSE,
                      warning = FALSE,
                      collapse = TRUE,
                      fig.height = 4,
                      fig.width = 8,
                      fig.align = "center",
                      cache = FALSE)
tutorial_html_dependency()
```

# Basic data visualisation

## Course web site

This is a link to the course web site, in case you need to go back and forth between tutorial and web materials: http://dmac.dicook.org

## Overview

- Going from tidy data to a data plot, using a grammar
- Mapping of variables from the data to graphical elements
- Using different geoms

### *FOLLOW ALONG WITH THE NOTES* 

- Open your ETC1010 `Rproj`
- Download the `Rmd` file from the class web site

### Turn in your individual assignment

Go to ED, the assessments tab. Upload your `Rmd`, and `html`, files. 

## Example: Tuberculosis data

This is current tuberculosis data taken from [WHO](http://www.who.int/tb/country/data/download/en/), the case notifications table.

We put the data into tidy form, and select just the counts for Australia

```{r}
library(tidyverse)
tb <- read_csv("data/TB_notifications_2018-03-18.csv") %>% 
  select(country, iso3, year, new_sp_m04:new_sp_fu) %>%
  gather(stuff, count, new_sp_m04:new_sp_fu) %>%
  separate(stuff, c("stuff1", "stuff2", "genderage")) %>%
  select(-stuff1, -stuff2) %>%
  mutate(gender=substr(genderage, 1, 1), 
         age=substr(genderage, 2, length(genderage))) %>%
  select(-genderage)

tb_au <- tb %>% 
  filter(country == "Australia") %>%
  filter(!(age %in% c("04", "014", "514", "u"))) %>%
  filter(year > 1996, year < 2013)
```

## 100% charts

```{r echo=TRUE, fig.width=8, fig.height=2}
ggplot(tb_au, aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(~ age) +
  scale_fill_brewer(palette="Dark2")
```

100% charts, is what excel names these beasts. What do we learn?

### Code structure

- Basic plot function is `ggplot`
- First argument provided is the name of the data, `tb_au`
- Variable mapping: year is mapped to x, count is mapped to y, gender is mapped to colour, and age is used to subset the data and make separate plots
- The bar geom is used, `geom_bar`
- We have already counted how many TB incidences are in each combination of categories, so `stat = "identity"` says no need to compute the count
- We are mostly interested in proportions between gender, over years, separately by age. The `position = "fill"` option in `geom_bar` sets the heights of the bars to be all at 100%. It ignores counts, and emphasizes the proportion of males and females.

### What do we learn?

- Focus is on **proportion** in each category. 
- Across (almost) all ages, and years, the proportion of males having TB is higher than females
- These proportions tend to be higher in the older age groups, for all years.


## Bar charts

```{r echo=TRUE, fig.width=8, fig.height=2}
ggplot(tb_au, aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(~ age) +
  scale_fill_brewer(palette="Dark2")
```

### What is different in the code description?

`, position = "fill"` was removed

### What do we learn?

- Focus is on **counts** in each category. 
- Different across ages, and years, counts tend to be lower in middle age (45-64)
- 1999 saw a bit of an outbreak, in most age groups, with numbers doubling or tripling other years.
- Incidence has been increasing among younger age groups in recent years.


## Side-by-side barcharts

```{r echo=TRUE, fig.width=8, fig.height=2}
ggplot(tb_au, aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity", position="dodge") +
  facet_grid(~ age) +
  scale_fill_brewer(palette="Dark2")
```

### What is different in the code description?

`, position="dodge"` is used in `geom_bar`

### What do we learn?

- Focus is on counts by gender, predominantly male incidence.
- Incidence among males relative to females is from middle ag on. There is similar incidence between males and females in younger age groups.

## Separate bar charts

```{r echo=TRUE, fig.width=8, fig.height=3}
ggplot(tb_au, aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(gender ~ age) +
  scale_fill_brewer(palette="Dark2")
```

### What is different in the code description?

`facet_grid(gender ~ age) +` faceted by gender as well as age, note `facet_grid` vs `facet_wrap`

### What do we learn?

- Its easier to focus separately on males and females.
- The 1999 outbreak mostly affected males.
- The growing incidence in the 25-34 age group is still affecting females but seems to be have stablised for males.

## Pie charts?

```{r echo=TRUE, fig.width=8, fig.height=3}
ggplot(tb_au, aes(x = year, y = count, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_grid(gender ~ age) +
  scale_fill_brewer(palette="Dark2") + 
  coord_polar() +
  theme(axis.text = element_blank())
```

Nope! That's a rose chart. Bar charts in polar coordinates produce rose charts. 

### What is different in the code description?

`coord_polar() +` plot is made in polar coordinates, rather than the default Cartesian coordinates

### What do we learn?

- Emphasizes the middle years as low incidence. 

## Rainbow charts?

```{r echo=TRUE, fig.width=8, fig.height=3}
ggplot(tb_au, aes(x = 1, y = count, fill = factor(year))) +
  geom_bar(stat = "identity", position="fill") +
  facet_grid(gender ~ age) 
```

A single stacked bar, in each facet. Year is mapped to colour. 

### What is the code doing?

- Notice how the mappings are different. A single number is mapped to x, that makes a single stacked bar chart.
- year is now mapped to colour (that's what gives us the rainbow charts!)

### What do we learn?

- Pretty chart but not easy to interpret. 

## Pie charts

```{r echo=TRUE, fig.width=8, fig.height=3}
ggplot(tb_au, aes(x = 1, y = count, fill = factor(year))) +
  geom_bar(stat = "identity", position="fill") +
  facet_grid(gender ~ age) +
  coord_polar(theta="y") +
  theme(axis.text = element_blank())
```

### What is different in the code description?

`coord_polar(theta="y")` is using the y variable to do the angles for the polar coordinates to give a pie chart.

### What do we learn?

- Pretty chart but not easy to interpret, or make comparisons across age groups. 

## Lab exercise

1. Pull a copy of the Melbourne pedestrian sensor data for July, 2018, using the `rwalkr` package, filter to keep two sensors "Melbourne Central", "Flinders Street Station Underpass". You can use the code below:

```{r eval=FALSE}
library(rwalkr)
library(lubridate)
ped <- walk_melb(from=dmy("01072018"),
                 to=dmy("31072018"))
ped <- ped %>% 
  filter(Sensor %in% c("Melbourne Central", "Flinders Street Station Underpass"))
save(ped, file="data/ped_July18.rda")
```

2. Make a side-by-side boxplot of pedestrian counts by sensor. Describe what you learn.

```{r eval=FALSE}
ggplot(ped, aes(x=Sensor, y=Count)) + geom_boxplot()
```

3. Subset to Monday of the first week of Monash classes. Make line plots of Count by Time, separately for each sensor. Describe the differences in the pedestrian patterns at both locations.

```{r eval=FALSE}
ped %>%
  filter(Date == dmy("23072018")) %>%
ggplot(aes(x=Time, y=Count)) + 
  geom_line() +
  facet_wrap(~Sensor, ncol=1)
```

Take the lab quiz!

(And use the rest of the lab time to coordinate with your group on the first assignment.)
