---
title: "Automating scraping with functions"
author: "YOUR NAME"
output: html_document
---

```{r setup}
library(tidyverse)
library(polite)
library(rvest)
```

# Scrape the list of most popular TV shows on IMDB: http://www.imdb.com/chart/tvmeter

```{r}
url <- "http://www.imdb.com/chart/tvmeter"
check_url <- bow(url)

tv_data <- scrape(check_url)

scrape_show_info <- function(x){
  title <- x %>%
    html_node("#title-overview-widget h1") %>%
    html_text() %>%
    str_trim()
  runtime <- x %>%
    html_node("time") %>%
    html_text() %>%
    str_replace("\\n", "") %>%
    str_trim()
  genres <- x %>%
    html_nodes(".txt-block~ .canwrap a") %>%
    html_text() %>%
    str_trim() %>%
    paste(collapse = ", ")
  tibble(title = title, runtime = runtime, genres = genres)
}


st %>%
  html_nodes("#title-overview-widget .subtext") %>%
  html_text() %>%
  str_trim() %>%
  str_remove_all("\\n") %>%
  str_split("\\|") %>%
  map(str_trim)

```


```{r}
tv_html <- bow("http://www.imdb.com/chart/tvmeter") %>% scrape()

library(glue)

web_stubs <- tv_html %>%
  html_nodes(".titleColumn a") %>%
  html_attr("href") %>%
  paste("http://www.imdb.com/", ., sep = " ")
```

## Examine each of the *first three* (or however many you can get through) tv show subpage to also obtain genre and runtime.


## Time permitting, also try to get the following:

  - How many episodes so far
  - First five plot keywords
  - Country
  - Language
  
Add this information to the data frame you created in step 1.


# Other fun functions

## Calculate the difference between the minimum and maximum value

```{r}
my_range <- function(x){}

```

## 
