---
title: "Assignment 3"
author: "NMS2"
date: "20 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo = FALSE, include=FALSE}
library(tidyverse)
library(naniar)
library(broom)
library(visdat)
library(labelled)
library(gridExtra)
houses <- read_csv("data/Melbourne_housing_FULL.csv")
```


## Missing value summary
```{r}
houses2 <- houses %>%
  select(Price, Rooms, Type, Distance, Bedroom2, Bathroom)
vis_dat(houses2)
```

```{r}
miss_summary(houses2)
vis_miss(houses2,sort_miss = TRUE)
```

## Missing Values Strategy
####It can be seen from these summaries that there are only missing values for three variables: the price of a property (Price), the number of bathrooms it has (Bathroom) and the number of bedrooms it has (Bedroom2). Each of these three variables has just over 20% of values missing, which is a significant amount. Given the model will have Price as the response variable, observations with a missing value for Price should be excluded (which R automatically does). For the potential explanatory variables Bathroom and Bedroom2, one approach to 'fill in' the missing values is to use the mean for these numerical variables rounded to the nearest positive integer. However, this would cause there to be a lot of observations with the same value for these variables. A better option might be to find the rounded average of each of these variables for every suburb and use those values for every missing value for observations in those suburbs. If there is a suburb with all missing values, the rounded average of the entire variable could be used. However, this approach could potentially be complicated and time-consuming, especially if it has to be done for each suburb individually. The summaries of missing values above indicate that at least the majority of properties with a missing value for Bathroom also have a missing value for Bedroom2 (possibly all). For these observations, this is missing values for 1/3 of the variables we are considering for some role in the model, a significant proportion. As this is the case, it might just be best to leave out these observations (automatically done by R), even though there are many of them.

```{r}
houses2 <- houses2 %>% filter(!is.na(Price))
```

## plots summarising relationship between price
```{r}
p1<- ggplot(houses2, aes(x=Bedroom2, y=Price)) + 
  geom_point(alpha=0.1) + 
  scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="bottom") + 
  xlab('Bedrooms') + 
  ylab("Price")
p2 <- ggplot(houses2, aes(x=Rooms, y=Price)) + 
  geom_point(alpha=0.1) + 
  scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="bottom") + 
  xlab('Rooms') + 
  ylab("Price")
p3 <- ggplot(houses2, aes(x=Type, y=Price)) + 
  geom_point(alpha=0.1) +
  scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="bottom") + 
  xlab('Type') + 
  ylab("Price")
p4 <- ggplot(houses2, aes(x=Bathroom, y=Price)) + 
  geom_point(alpha=0.1) + 
  scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="bottom") + 
  xlab('Bathrooms') + 
  ylab("Price") 
p5 <- ggplot(houses2, aes(x=Distance, y=Price)) + 
  geom_point(alpha=0.1) + 
  scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="bottom") + 
  xlab('Distance') + 
  ylab("Price")
grid.arrange(p1, p2, p3, p4, p5, ncol=2)
```

## Summary Plot Explanations
####The Price-Distance scatterplot suggests that properties with high prices are most likely to be within 15 km of the Melbourne CBD compared to further away. The range most property prices are in generally increases as the number of bedrooms and bathrooms (and total rooms) increase. However, while there are more highly-priced properties when there is a higher number of rooms, there are still many at lower prices. This is likely due in part to higher distances from the CBD helping lower prices. As the number of bedrooms, bathrooms and rooms increases even further, property prices mostly are at lower levels than before, however there is not as much data for these properties. There is a range of prices for all property types (houses, townhouses and units), however it is more common for houses to be among the highest half of prices than the other types according to the Price-Type scatterplot. Because there are a lot of data points, there is a lot of black in the scatterplots, which makes it more difficult to definitively draw conclusions from these graphs.


## build model:
### Model 1
```{r}
houses3 <- houses2 %>% filter(Rooms < 10)

summarise(houses2 %>% group_by(Rooms), ave = mean(Price), count = n())


mod1 <- lm(Price~Rooms+Type+Distance+Bedroom2+Bathroom, data=houses2)
tidy(mod1)
glance(mod1)
confint_tidy(mod1)
mod1_fit <-augment(mod1,houses2)
p1<- ggplot(mod1_fit, aes(x=.fitted, y=.std.resid)) + 
  geom_point()

ggplot(mod1_fit) + geom_point(aes(x = Rooms, y = Price)) + geom_smooth(aes(x = Rooms, y =.fitted))
```

### Model 2
```{r}
mod2 <- lm(Price~Rooms+I(Rooms^2)+Type+Distance+I(Distance^2)+Bathroom, data=houses2)
tidy(mod2)
glance(mod2)
confint_tidy(mod2)
mod2_fit <-augment(mod2,houses2)
p2<- ggplot(mod2_fit, aes(x=.fitted, y=.std.resid)) + 
  geom_point()

ggplot(mod2_fit) + geom_point(aes(x = Rooms, y = Price)) + geom_smooth(aes(x = Rooms, y =.fitted))
ggplot(mod2_fit) + geom_point(aes(x = Distance, y = Price)) + geom_smooth(aes(x = Distance, y =.fitted))

grid.arrange(p1, p2, ncol=2)
```

### Model 3
```{r}
mod3 <- lm(Price~Rooms*I(Rooms^2)*Type*Distance*I(Distance^2)*Bathroom, data=houses2)
tidy(mod3)
glance(mod3)
confint_tidy(mod3)
mod3_fit <-augment(mod3,houses2)
p3<- ggplot(mod3_fit, aes(x=.fitted, y=.std.resid)) + 
  geom_point()

ggplot(mod3_fit) + geom_point(aes(x = Rooms, y = Price)) + geom_smooth(aes(x = Rooms, y =.fitted))
ggplot(mod3_fit) + geom_point(aes(x = Distance, y = Price)) + geom_smooth(aes(x = Distance, y =.fitted))

grid.arrange(p1, p2, p3, ncol=2)
```

## Explain your choices of transformations, variable selection and interactions

#### From the summary plots above, the relationship betweeen price and distance shows a pattern. Price increases as distance increases up to some point, after which the price starts to decline. This relationship justifies the use of a quadratic variable for disatnce in the model. A similar observation can be made with number of bedrooms, there is an optimun number of bedrooms for a house. From this we decided that Model 3 was the best model to acuratly help us predict houseing prices for the properties around Melbourne. 



